### 2023-07-28 manuscript/manuscript.qmd
Manuscript build working

---
author: 
  - name: "Kjell Jorner"
    orcid: 0000-0002-4191-6790
    email: "kjell.jorner@chem.ethz.ch"
    corresponding: true
    affiliations: 
      - ref: eth
      - ref: chalmers
      - ref: toronto-chem
      - ref: toronto-cs
  - name: "Robert Pollice"
    orcid: 0000-0001-8836-6266
    affiliations: 
      - ref: rug
      - ref: toronto-chem
      - ref: toronto-cs  
  - name: "Cyrille Lavigne"
    orcid: 0000-0003-2778-1866
    affiliations: 
      - ref: toronto-chem
      - ref: toronto-cs    
  - name: "Alán Aspuru Guzik"
    orcid: 0000-0002-8277-4434
    affiliations:
      - ref: toronto-chem
      - ref: toronto-cs
      - ref: toronto-chem-eng
      - ref: toronto-mat-eng
      - ref: vector
      - ref: ac

affiliations:
  - id: eth
    name: "ETH Zürich"
    department: "Institute of Chemical and Bioengineering, Department of Chemistry and Applied Biosciences"
    address: "Vladimir-Prelog-Weg 1"
    postal-code: "CH-8093"
    city: "Zürich"
    country: "Switzerland"
  - id: chalmers
    name: "Chalmers University of Technology"
    department: "Department of Chemistry and Chemical Engineering"
    address: "Kemigården 4"
    postal-code: "SE-41258"
    city: "Gothenburg"
    country: "Sweden"
  - id: toronto-chem
    name: "University of Toronto"
    department: "Chemical Physics Theory Group, Department of Chemistry"
    address: "80 St. George St."
    postal-code: "M5S 3H6"
    city: "Toronto"
    state: "Ontario"
    country: "Canada"
  - id: toronto-cs
    name: "University of Toronto"
    department: "Department of Computer Science"
    address: "40 St. George St."
    postal-code: "M5S 2E4"
    city: "Toronto"
    state: "Ontario"
    country: "Canada"
  - id: rug
    name: "University of Groningen"
    department: "Stratingh Institute for Chemistry"
    address: "Nijenborgh 4"
    postal-code: "9747 AG"
    city: "Groningen"
    country: "The Netherlands"
  - id: toronto-chem-eng
    name: "University of Toronto"
    department: "Department of Chemical Engineering & Applied Chemistry"
    address: "200 College St."
    postal-code: "M5S 3E5"
    city: "Toronto"
    state: "Ontario"
    country: "Canada"
  - id: toronto-mat-eng
    name: "University of Toronto"
    department: "Department of Materials Science & Engineering"
    address: "184 College St."
    postal-code: "M5S 3E4"
    city: "Toronto"
    state: "Ontario"
    country: "Canada"    
  - id: vector
    name: "Vector Institute for Artificial Intelligence"
    address: "661 University Ave. Suite 710"
    postal-code: "M5G 1M1"
    city: "Toronto"
    state: "Ontario"
    country: "Canada" 
  - id: cifar
    name: "Lebovic Fellow"
    department: "Canadian Institute for Advanced Research (CIFAR)"
    postal-code: "M5G 1M1"
    city: "Toronto"
    state: "Ontario"
    country: "Canada" 
  - id: ac
    name: "Acceleration Consortium"
    department: "University of Toronto"
    city: "Toronto"
    state: "Ontario"
    country: "Canada"

title: "Ultrafast computational screening of molecules with inverted singlet-triplet energy gaps using semi-empirical quantum chemistry"
bibliography: [references.bib] 
link-citations: true
date: last-modified
format:
  elsevier-pdf:
    cite-method: citeproc
    journal:
      name: "Journal of Physical Chemistry A"
      model: "3p"    
    number-sections: false
    keep-tex: false            
    include-in-header:
        - file: macros.tex    
csl: "acs.csl"
---

<!-- Definition of parameters for tables -->
```{python}
# | tags: [parameters]
# | echo: false

tab_local_azaphenalenes = ""
tab_local_azulenes = ""
tab_global_all_gaps = ""
tab_global_scaffolds = ""
tab_garner_gaps = ""
```

# Introduction

Organic light-emitting diodes (OLEDs) is a technology for generating light from electricity using organic molecules.[@yangRecentAdvancesOrganic2017] The first generation of OLEDs were limited to an efficiency of max 25% due to spin statistics. As the transition T~1~ $\rightarrow$ S~0~, from the lowest excited triplet state to the singlet ground state, is quantum-mechanically forbidden, the OLED molecule can formally only emit from its first excited state of singlet multiplicity, S~1~. The second generation of OLEDs exploited ways of increasing the rate of this spin-forbidden phosphoresence. The third generation of OLEDs are based on Thermally Activated Delayed Fluorescence (TADF), where the S~1~ state is partially populated from a near-lying T~1~ state. For this to happpen, the gap $\DeltaE{ST}$ between the states needs to sufficiently small so that the thermal activation competes with non-radiative decay processes from T~1~. However, as the T~1~ population is still substantial, problems with stability and lower than ideal intensity due to non-radiative decay persist. The logical step for the next generation of OLEDs is to emit directly from an S~1~ state that lies below the T~1~ state, potentially converting all of the electrically generated excitons into photons, _i.e._, 100% efficiency.

Molecules that fulfill this criterion have been branded with the acronym INVErted Singlet Triplet energy gap (INVEST) and are exceedingly rare.[@liOrganicMoleculesInverted2022] They are said to violate Hund's rule of maximum multiplicity, which describes that the electronic state with the highest spin (in this case the triplet) should be lowest in energy. While examples of molecules violating Hund's rule in the excited state have been known since the 1970s and 1980s,[@kollmarViolationHundRule1978; @leupinLowlyingElectronicallyExcited1980; @kosekiViolationHundMultiplicity1985] it is only recently that they have received considerable attention for application in OLEDs.[@desilvaInvertedSingletTriplet2019;@polliceOrganicMoleculesInverted2021;@aizawaDelayedFluorescenceInverted2022] As they are rare, recent efforts have used high-throughput virtual screening (HTVS) with computational chemistry to identify new compounds with inverted gaps, focusing on expanding the local chemical space around specific scaffolds[@polliceOrganicMoleculesInverted2021;@aizawaDelayedFluorescenceInverted2022] or scanning larger (combinatorially or experimentally derived) datasets to identify novel scaffolds.[@terenceblaskovitsSymmetryInducedSinglet2023;@omarIdentificationUnknownInverted2023;@garnerDoublebondDelocalizationNonalternant2023] These HTVS efforts have been successful at identifying several new INVEST candidates, some of which have also been synthesized and tested experimentally.[@aizawaDelayedFluorescenceInverted2022]

While this early success of HTVS is highly encouraging, it is hampered by the cost of the computational methods. Excitation energies are routinely calculated by time-dependent density functional theory (TD-DFT), which has become a workhorse of computational photophysics over the last decades.[@casidaProgressTimeDependentDensityFunctional2012] Unfortunately, it has been shown that TD-DFT fails to capture the inverted gap of INVEST compounds, as the method only considers single electronic excitations.[@desilvaInvertedSingletTriplet2019] Inclusion of at least double excitations are necessary to reproduce the inverted gaps, corresponding to methods such as double hybrid TD-DFT or excited state coupled-cluster methods such as second-order approximate coupled cluster singles and doubles (CC2) or equation of motion coupled cluster with single and double excitations (EOM-CCSD). In line with the results of benchmarking studies,[@ricciSingletTripletExcited2021;@sancho-garciaViolationHundRule2022] recent HTVS studies have used methods such as the complete active space self-consistent field (CASSCF) and CC2 for preliminary screening, while confirming inverted gaps with more expensive methods such as multistate complete active space second-order perturbation theory (MS-CASPT2) and EOM-CCSD. These methods are not only expensive, but in the case of the CAS methods, also require the choice of an orbital active space, something which is not routinely automatized.

Motivated by the need for faster methods for HTVS of INVEST compounds, we wondered if it would be possible to perform much simpler calculations as a pre-screening step for the more expensive methods. Based on prior work in the literature from the 1970s and 1980s, we find that the inverted gaps can be recovered already with the semi-empirical Pariser-Parr-Pople method using configuaration interaction singles (CIS) and adding key double excitations. The PPP method considers only the $\pi$ electrons in a minimal valence basis and approximates all integrals from experimental data and a few empirical parameters, making it computationally extremly cheap. At the same time, it retains the conceptual clarity of Hückel theory and allows for straightforward interpretation of the inverted gap in terms of the well-known concept of dynamic spin polarization (DSP). We show that the method is capable of finding hits both in the local chemical space around known scaffolds, and of identifying hits in more diverse datasets. While there are clear limitations to the method, we believe that it will open the doors for ultra high-throughput virtual screening (UHTVS) of compound libraries orders of magnitude larger than today. The method is also perfectly applicable as a scoring function in inverse design algorithms.

# Theory

While the recent literature has focused on applying high-level *ab initio* and DFT methods to study INVEST compounds, we believe that some conceptual clarity has been lost in process. Instead, we apply the simplest possible electronic structure method that can capture the physics of the problem. The Pariser-Parr-Pople (PPP) method is an extension of the semi-empirical Hückel molecular orbital theory for $\pi$-electron systems that includes electron correlation.[@pariserSemiEmpiricalTheory1953a;@pariserSemiEmpiricalTheory1953; @popleElectronInteractionUnsaturated1953] It originated in the 1950s, with new parametrizations being developed mainly in the 1960s and 1970s, and was used in the dye industry at least until the 1980s before the advent of more accurate ab initio methods.[@griffithsPracticalAspectsColour1982] The PPP method retains the conceptual clarity of Hückel theory, while at the same time including the electron correlation that is necessary to capture the inverted gap.

## Pariser-Parr-Pople theory

Within Pople's formulation of PPP theory, Roothan's equations are solved self-consistently.[@popleElectronInteractionUnsaturated1953]. Under the Zero Differential Overlap (ZDO) approximation,<!-- REF: pariser, parr, pople, maybe Dral paper --> overlap integrals are neglected for orbitals on different centers, leading to a simplification of the Roothan equations as $\mathbf{S}$ becomes the unit matrix $\mathbf{I}$

$$
\mathbf{FC} = \mathbf{SCE} = \mathbf{ICE} = \mathbf{CE}
$$ {#eq-roothan}

where $\mathbf{F}$ is the Fock matrix, $\mathbf{C}$ contains the molecular orbital coefficients and $\mathbf{E}$ is the diagonal matrix of the orbital energy eigenvalues. Given $\mathbf{F}$, the corresponding $\mathbf{C}$ and $\mathbf{E}$ can then be determined either self-consistently (following Pople[@popleElectronInteractionUnsaturated1953]) or via configuration interaction (following Pariser and Parr[@pariserSemiEmpiricalTheory1953a; @pariserSemiEmpiricalTheory1953]), most often starting from a guess $\mathbf{C}$ from the corresponding Hückel model. The main effort is then to construct $\mathbf{F}$, which has the following matrix elements[@beveridgeParametrizationSemiempiricalPi1971]
$$
F_{rr} = H_{rr} + \sum_{t} P_{tt} \gamma_{rt} - \frac{1}{2} P_{rr} \gamma_{rr} 
$$ {#eq-f-rr}

$$
F_{rs} = H_{rs} - \frac{1}{2} P_{rs} \gamma_{rs}
$$ {#eq-f-rs}

where $P_{rs}$ are elements of the density matrix $\mathbf{P}$ and $\gamma_{rr}$ and $\gamma_{rs}$ are parameters called the *one-center and two-center electron repulsion integrals*, respectively. The *core resonance integral* matrix elements of $\mathbf{H}$ are given by
$$
H_{rr} = \alpha_r = U_r - \sum_{s \neq r} Z_s \gamma_{rs}
$$ {#eq-h-rr}
$$
H_{rs} = \beta_{rs}
$$ {#eq-h-rs}

where $\alpha_{r}$ is the *one-center core resonance integral*, $U_r$ is a parameter of model called the *atomic valence state potential*, $Z_s$ is the effective nuclear charge and $\beta_{rs}$ is the *two-center resonance integral*, another parameter of the model.

## Parametrization

The four parameters are thus $U_{r}$, $\gamma_{rr}$, $\gamma_{rs}$ and $\beta_{rs}$. They are derived either from experiment, first principles, or a combination of both. The Pariser-Parr approximation leads to
$$
U_{r} = - \mathrm{IP}_v
$$ {#eq-u-r}

where $\mathrm{IP}_v$ is the *valence-state ionization potential* of the orbital and atom in questions (*e.g.* a 2*p* orbital of an *sp*^2^-hybridized C atom).[@mullikenNewElectroaffinityScale1934] The $\mathrm{IP}_v$ values (and the corresponding electron affinities $\mathrm{EA}_v$) are tabulated by Hinze and Jaffé as determined from experimental data.[@hinzeElectronegativity1962; @hinzeElectronegativityOrbitalElectronegativity1962] $\gamma_{rr}$ also enjoys an almost universally adopted approximation due to Pariser and Parr:
$$
\gamma_{rr} = \mathrm{IP}_v - \mathrm{EA}_v
$$ {#eq-gamma-rr}

For $\gamma_{rs}$ and $\beta_{rs}$, there is much less consensus. Formulas for $\gamma_{rs}$ have been suggested by, among others, Mataga and Nishimoto[@matagaElectronicStructureSpectra1957] and Ohno[@ohnoRemarksPariserParrPopleMethod1964], but here we follow the approach by Beveridge and Hinze[@beveridgeParametrizationSemiempiricalPi1971]:
$$
\gamma_{rs} = \frac{1}{a_{rs} \exp{\left( -r_{rs}^2/2a_{rs}^2\right)} + r_{rs}}
$$ {#eq-gamma-rs}

where $r_{rs}$ is the distance between the two atoms $r$ and $s$, and
$$
a_{rs} = \frac{2}{\gamma_{rr} + \gamma_{ss}}
$$ {#eq-a}

For $\beta_{rs}$, expressions have been given by, *e.g.*, Linderberg[@linderbergConsistencyRequirementPariserParrPople1967] and Jug[@jugOperatorEquationsApproximate1972], based on first principles and by Dewar[@chungGroundStatesConjugated1965] based on experimental data. Here we again follow Beveridge and Hinze,[@beveridgeParametrizationSemiempiricalPi1971] who derived, following Ohno[@ohnoRemarksPariserParrPopleMethod1964]:

$$
\beta_{rs} = \frac{1}{2} \left( Z_r + Z_s \right) S_{rs} \left( \gamma_{rs} - \frac{2C}{r_{rs}} \right)
$$ {#eq-beta-rs}

where $C = 0.545$ is a parameter that was fit in the original publicaiton to reproduce experimental excitation energies.[@beveridgeParametrizationSemiempiricalPi1971] The overlap integral $S_{rs}$ is determined from the overlap of Slater *p* orbitals with exponents

$$
\zeta_r = \frac{1280}{501} \gamma_{rr}
$$

according to recursion formulas given by Mulliken.[@mullikenFormulasNumericalTables1949] Thus, the four parameters $U_{r}$, $\gamma_{rr}$, $\gamma_{rs}$ and $\beta_{rs}$ are expressed completely in terms of the corresponding valence state ionization potentials $\mathrm{IP}_v$ and electron affinities $\mathrm{EA}_v$ as well as the completely empirical parameter $C$.

The distance dependence of $\beta_{rs}$ and $\gamma_{rs}$ allows for treatment of bond-length alternation beyond idealized geometries. In addition, we have followed the literature[@murrellSemiempiricalSelfconsistentfieldMolecular1972] and added an angle-dependence to $\beta_{rs}$ according to

$$
\beta_{rs}^\prime = \beta_{rs} \cos{\theta}
$$

where $\theta$ is the twist angle between the two *p* orbitals. We have determined $\theta$ as the average of all possible dihedrals involving the two atoms $r$ and $s$.

## Excited state calculations

The energy difference of the vertically excited S~1~ and T~1~ states can be determined by Roothan's expression at the SCF level[@roothaanNewDevelopmentsMolecular1951]
$$
\DeltaEST{SCF} = \E{S} - \E{T} = 2 \left( xy \vert yx \right) = 2 \left( x \lvert K_y \rvert x \right) = 2K 
$$ {#eq-exchange}

where $K$ is the exchange integral between the two orbitals involved in the single excitation (normally the HOMO and the LUMO). As defined here, a positive $\DeltaE{ST}$ corresponds to the usual situation with the triplet being lower in energy than the singlet, while a positive $\DeltaE{ST}$ corresponds to an inverted gap. The exchange integral depends strongly on the spatial overlap between the the two orbitals, which in the ZDO approximation can be computed as:[@peachExcitationEnergiesDensity2008]

$$
O_{xy} = \sum_{r} |c_{r,x}||c_{r,y}|
$$

where $|c_{r,i}|$ is the absolute value of the coefficient of orbital $i$ on center $r$. When the overlap is zero, the exchange interaction vanishes and $\DeltaEST{SCF}$ is zero. Unfortunately, so is the oscillator strength, $f$, between the two states,[@klessingerExcitedStatesPhotochemistry1995] which is non-optimal for OLED materials that should emit light. Consequently, there is a trade-off between having a small HOMO-LUMO overlap to reduce the exchange interaction, while still maintaining a sufficient oscillator strength.

A more precise expression for $\DeltaE{ST}$ can be obtained by configuration interaction singles (CIS).[@beveridgeParametrizationSemiempiricalPi1971] While CIS adds some correlation for the excited states, it is necessary to add additional excitations beyonds singles to capture the inverted $\DeltaE{ST}$. Fortunately, the most important excitations have already been identified in the literature long ago. Singlet-triplet inversion occurs also for the ground state, where several violations of Hund’s rule are well known, for example for bond-equalized cyclobutadiene at the *D*~4h~ geometry and pentalene at the *D*~2h~ geometry. Borden and Davidson explained this effect in terms of dynamic spin polarization (DSP),[@bordenTheoreticalStudiesDiradicals1981] in which the electrons of a pair of disjoint non-bonding molecular orbitals (NBMOs, [@fig-dsp]a) experience stronger electron correlation effects in the singlet state as compared to the triplet state ([@fig-dsp]b). As outlined by Kollmar and Staemmler,[@kollmarViolationHundRule1978] as well as Malrieu and co-workers,[@benamorSpinPolarizationElectronic2020] an alternative view of the same phenomenon is through admixture of singly-excited configurations into the electronic wavefunction of the ground state ([@fig-dsp]c). The DSP effect for the S~1~ excited state can correspondingly be described by admixture of doubly excited configurations,[@kollmarViolationHundRule1978] as they are singly excited with respect to the S~1~ state, and doubly excited with respect to the S~0~ ground state ([@fig-dsp]d).

As first shown by Kollmar and Staemmler,[@kollmarViolationHundRule1978] and recently re-emphasized by Pernal and co-workers,[@drwalRoleSpinPolarization2023] the $\DeltaE{ST}$ can be approximated by a combination of two terms, the first one being the exchange interaction $2K$ (given by @eq-exchange) and a correction term $\DeltaE[DSP]{}$ due to the DSP, which can be approximated with perturbation theory:[@kollmarViolationHundRule1978]

![Test to write something here](../data/figures/dynamic_spin_polarization.pdf){#fig-dsp}

$$
\DeltaE[DSP]{SCF} =  - \sum_{i}\sum_{k}
    \frac{1}{2} \left( 
    \frac{3 \Integral{-}}{\EState[1]{S} - \EState{S}}
    - \frac{\Integral{-}}{\EState[1]{T} - \EState{T}}
    - \frac{2 \Integral{+}}{\EState[2]{T} - \EState{T}}
    \right)
$$

Here, $K_x$ and $K_y$ are exchange operators, $E(\phi_{\mathrm{S}})$ and $E(\phi_{\mathrm{T}})$ are the energies of the singlet and triplet excited determinants, while $E (\phi_{\mathrm{S}}^{1})$, $E (\phi_{\mathrm{T}}^{1})$ and  $E(\phi_{\mathrm{T}}^{2})$ are the energies of doubly excited determinants. They can be written in terms of the correspond orbital energies, $\epsilon_i$, provided the same orbitals are used for both the singlet and triplet states.[@benamorSpinPolarizationElectronic2020]

$$
\DeltaE[DSP]{SCF} =  - \sum_{i}\sum_{k}
    \frac{1}{2} \left( 
    \frac{3 \Integral{-}}{\epsilon_k - \epsilon_i}
    - \frac{\Integral{-}}{\epsilon_k - \epsilon_i}
    - \frac{2 \Integral{+}}{\epsilon_k - \epsilon_i}
    \right)
$$

The simplest approximation for $\DeltaE{ST}$ taking DSP into account is then

$$
\DeltaEST[DSP]{SCF} = 2K + \DeltaE[DSP]{SCF}
$$

Alternatively, $\DeltaE[DSP]{}$ can be calculated with respect to the CIS states,[@head-gordonDoublesCorrectionElectronic1994]

$$
\begin{split}
\DeltaE[DSP]{CIS} =  - \sum_{i}\sum_{k}
    \frac{1}{2} \left( 
    \frac{3 \Integral{-}}{\epsilon_k + \epsilon_y - \epsilon_y - \epsilon_i - \DeltaE{S,CIS}}
    - \frac{\Integral{-}}{\epsilon_k + \epsilon_y - \epsilon_y - \epsilon_i - \DeltaE{T,CIS}} \right. \\
    \left. - \frac{2 \Integral{+}}{\epsilon_k + \epsilon_y - \epsilon_y - \epsilon_i - \DeltaE{T,CIS}}
    \right)
\end{split}
$$

and added as a correction to the energy gap at the CIS level to yield a possibly more accurate value

$$
\DeltaEST[DSP]{CIS} = \DeltaEST{CIS} + \DeltaE[DSP]{CIS}
$$

# Methods

The PPP method was implemented in the Python package [coulson]{.smallcaps}[@jornerCoulson] which is freely available on GitHub with a permissive MIT license. Further details on [coulson]{.smallcaps} will be reported elsewhere. The PPP wavefunction was converged with the self-consistent field method using the naive variational principle. To speed up the calculations, we derived a spline interpolation of the overlap integrals,[@mullikenFormulasNumericalTables1949] which will be reported separately. All calculations have been performed on a MacBook Air laptop computer with an M2 processor. We further integrated [coulson]{.smallcaps} with [PySCF]{.smallcaps}[@sunRecentDevelopmentsPySCF2020] to provide alternative algorithms for the SCF convergence and CIS and found the results consistent within numerical accuracy. Oscillator strengths were calculated with the dipole length approximation.[@jorgensenTimedependentHartreeFockCalculations1970;@klessingerExcitedStatesPhotochemistry1995;@damourGroundExcitedStateDipole2023] Some compounds exhibit negative triplet exitation energies and in some case even negative singlet excitation energies, which indicates a restricted-unrestricted instability of the ground state wavefunction. We have here taken the pragmatic approach to ignore these issues, while we will highlight below some examples where it occurs and if it has any effect on the overall results. This approach can be partly justified as the reference data also does not include any stability analysis. The results should anyway be indicative of the gap between the lowest singlet and triplet states with open-shell character, regardless of if they are lower in energy than the closed-shell singlet.

# Results and discussion

## Qualitative understanding of the inverted gap

We first demonstrate that our method is capable of capturing the inverted $\DeltaE{ST}$ for some model compounds. As shown by Toyota and co-workers, pentalene at the ideal  $D_\mathrm{2h}$ geometry has an inverted gap, while relaxation to the minimum with $C_\mathrm{2h}$ symmetry leads to a normal gap.[@kosekiViolationHundMultiplicity1985] In our calculations, pentalene with equal bond lenghts of 1.4 Å displays a small HOMO-LUMO overlap of {{< var pentalene_ideal_overlap >}}, leading to a small exchange interaction of only $2K=$ {{< var pentalene_ideal_2_exchange >}} kcal/mol (@fig-pentalene-ideal-orbitals). The dynamic spin polarization correction $\DeltaE[DSP]{SCF}$ of {{< var pentalene_ideal_dsp_scf >}} kcal/mol leads to a net predicted $\DeltaEST[DSP]{SCF}$ of {{< var pentalene_ideal_t1_s1_dsp_scf >}} kcal/mol. Adding additional correlation with CIS leads to a $\DeltaEST[DSP]{CIS}$ of {{< var pentalene_ideal_t1_s1_dsp_cis >}} kcal/mol. The reference value calculated with {{< var reference_level >}} is {{< var azaphenalene_t1_s1_ref >}} kcal/mol. The simple perturbation theory model allows us to gain additional insight into the contributions to $\DeltaE[DSP]{SCF}$ as they correspond to single excitations between doubly occupied orbitals below the HOMO and uncoccupied orbitals above the LUMO (@fig-pentalene-ideal-excitations). Three excitations, HOMO-3 → LUMO+2, HOMO-2 → LUMO+1 and HOMO-1 → LUMO+3 are identified as the main contributors to the spin polarization effect, and could potentially be tuned by substituent effects.

![Orbitals energies and frontier orbitals of pentalene at the idealized $D_\mathrm{2h}$ geometry and bond lengths of 1.4 Å.]({{< var fig_pentalene_ideal_orbitals >}}.png){#fig-pentalene-ideal-orbitals}

![Excitations and their contribution to $\DeltaE[DSP]{SCF}=$ {{< var pentalene_ideal_dsp_scf >}} kcal/mol for pentalene at the $D_\mathrm{2h}$ geometry and bond lengths of 1.4 Å.]({{< var fig_pentalene_ideal_excitations >}}.png){#fig-pentalene-ideal-excitations}

Using the relaxed $C_\mathrm{2h}$ geometry leads to dramatic changes in the frontier orbitals, with an increased HOMO-LUMO overlap of {{< var pentalene_overlap >}} and a sizeable exchange interaction of {{< var pentalene_2_exchange >}} kcal/mol (@fig-pentalene-orbitals). Consequently, the calculated gap is now normal at $\DeltaEST[DSP]{SCF}=$ {{< var pentalene_t1_s1_dsp_scf >}} kcal/mol, aggravated by a diminished $\DeltaE[DSP]{SCF}$ of only {{< var pentalene_dsp_scf >}} kcal/mol. A similar result is obtained with CIS: $\DeltaEST[DSP]{CIS}=$ {{< var pentalene_t1_s1_dsp_cis >}} kcal/mol. The reference value calculated with {{< var reference_level >}} is {{< var pentalene_t1_s1_ref >}} kcal/mol. The excitations contributing to $\DeltaE[DSP]{SCF}$ at the $D_\mathrm{2h}$ geometry have been significantly diminished, only partially alleviated by the addition of a minor stabilization from HOMO-2 → LUMO+3 (@fig-pentalene-excitations).

![Orbitals energies and frontier orbitals of pentalene at the optimized $C_\mathrm{2h}$ geometry.]({{< var fig_pentalene_orbitals >}}.png){#fig-pentalene-orbitals}

![Excitations and their contribution to $\DeltaE[DSP]{SCF}=$ {{< var pentalene_dsp_scf >}} kcal/mol for pentalene at the optimized $C_\mathrm{2h}$ geometry.]({{< var fig_pentalene_excitations >}}.png){#fig-pentalene-excitations}

Azaphenalene is the prototypical molecule that started the renewed investigations into INVEST molecules, and has been the subject of numerous studies with high-level quantum chemical methods. Can we capture the inverted gap, as first measured by Leupin and Wirz?[@leupinLowlyingElectronicallyExcited1980] The calculations reveal that the HOMO and LUMO are well separated with spatial overlap of {{< var azaphenalene_overlap >}} and a correspondingly low exchange interaction of {{< var azaphenalene_2_exchange >}} kcal/mol ([@fig-azaphenalene-orbitals]). The larger azaphenalene has 30 excitations that could possibly contribute to the large $\DeltaE[DSP]{SCF}$ of {{< var azaphenalene_dsp_scf >}} kcal/mol, but the most import by far are those from the doubly degenerate HOMO-1 to the doubly degenerate LUMO+1 ([@fig-azaphenalene-excitations]). In total, $\DeltaEST[DSP]{SCF}=$ {{< var azaphenalene_t1_s1_dsp_scf >}} kcal/mol, which is slighly smaller at the CIS level with $\DeltaEST[DSP]{CIS}=$ {{< var azaphenalene_t1_s1_dsp_cis >}} kcal/mol.

To conclude this section on the model compounds, our PPP-based protocol is able to capture the inverted gap, and also provides qualitative understanding into the physical mechanism of dynamic spin polarization through identification and visualization of the corresponding excitations. For these two compounds, the calculated $\DeltaEST[DSP]{}$ are more negative at the SCF level compared to CIS. Compared to {{< var reference_level >}}, it would seem that CIS is preferable, but to elucidate to which method is better, we now turn to larger datasets.

![Orbitals energies and frontier orbitals of azaphenalene.]({{< var fig_azaphenalene_orbitals >}}.png){#fig-azaphenalene-orbitals}

![Excitations and their contribution to $\DeltaE[DSP]{SCF}=$ {{< var azaphenalene_dsp_scf >}} kcal/mol for azaphenalene]({{< var fig_azaphenalene_excitations >}}.png){#fig-azaphenalene-excitations}

## Local chemical space of azaphenalenes and azaazulenes

While our method captures the inverted gap of azaphenalene and pentalene, it also needs to capture trends with substition to work effectively in virtual screening. We therefore calculated 256 substituted azaphenalenes that have previously been studied by some of us.[@polliceOrganicMoleculesInverted2021] The PPP level S~1~ and T~1~ excitation energies are well correlated with those from {{< var reference_level >}} with $R^2$ of {{< var local_azaphenalenes_r2_singlet >}} and {{< var local_azaphenalenes_r2_triplet >}}, respectively ([@fig-local-azaphenalenes]a,b). Unfortunately, the oscillator strenghts are not as well reproduced with an $R^2$ of {{< var local_azaphenalenes_r2_f >}}, although the Spearman $\rho$ of {{< var local_azaphenalenes_rho_f >}} indicates the values might be used to rank potential candidates ([@fig-local-azaphenalenes]c). The problems of obtaining accurate oscillator strenghts at the PPP level of theory are well known in the literature, and the problem is especially exacerbated with low oscillator strengths as here.[@misurkinElectronicStructureConjugated1974]. Crucially, the $\DeltaE{ST}$ shows good correlations, with $R^2=$ {{< var local_azaphenalenes_r2_cis >}} using $\DeltaEST[DSP]{CIS}$ ([@fig-local-azaphenalenes]f). This $R^2$ value is essentially unchanged from the gap at the CIS level without DSP ({{< var local_azaphenalenes_r2_cis_raw >}}, [@fig-local-azaphenalenes]d) and markedly better than the gap at the SCF level ({{< var local_azaphenalenes_r2_scf >}}, [@fig-local-azaphenalenes]e), showing the importance of going beyond the SCF level to include at least some configuration interaction. The results are of similar or better quality achieved by Pernal and co-workers for different set of azaphenalenes using SCF + DSP with orbitals from DFT.[@drwalRoleSpinPolarization2023] To further analyze the results in terms of a binary classification into normal/inverted, we calculated the true positives (TP), true negatives (TN), false positives (FP) and false negatives (FN), as well as a range of common classification scores (@tbl-local-azaphenalenes). The most important metrics for virtual screening are, in our opinion, the recall and the specificity. The recall mesaures the fraction of inverted molecules that the protocol captures, and is {{< var local_azaphenalenes_recall_scf >}} at the SCF level and {{< var local_azaphenalenes_recall_cis >}} at the CIS level. The specificity measurs the freaction of non-inverted molecules that the protocol identifies, and is {{< var local_azaphenalenes_specificity_scf >}} at the SCF level and {{< var local_azaphenalenes_specificity_cis >}} at the CIS level. This metrics mean that we are able to capture a large proportion of the molecules with inverted gaps, while filtering out most non-inverted molecules. The results for these azaphenalenes clearly indicate that the PPP protocol could be used to prescreen candidates of this compound class for $\DeltaE{ST}$, while further pruning with another method is likely needed for the oscillator strengths.

![Write some text here]({{< var fig_local_azaphenalenes >}}.pdf){#fig-local-azaphenalenes}

```{python}
# | label: tbl-local-azaphenalenes
# | tbl-cap: Astronomical object
# | echo: false
import pandas as pd
from IPython.display import display, Markdown

with open(tab_local_azaphenalenes) as f:
    md = f.read()
display(Markdown(md))
```

While we see success for the azaphenalenes, the situation for the azulenes is unfortunately worse. The PPP level S~1~ and T~1~ excitation energies are only moderaly correlated with those from {{< var reference_level >}} with $R^2$ of {{< var local_azulenes_r2_singlet >}} and {{< var local_azulenes_r2_triplet >}}, respectively ([@fig-local-azulenes]a,b). For the oscillator strengths, the situation is rather catastrophic with $R^2$ of {{< var local_azulenes_r2_f >}} implying no correlation whatsoever ([@fig-local-azulenes]c). The situation for the $\DeltaE{ST}$ also worse than for the azaphenalenes, with $R^2=$ {{< var local_azulenes_r2_cis >}} with CIS + DSP, a marked improvement on {{< var local_azulenes_r2_scf >}} with SCF + DSP ([@fig-local-azulenes]e,f). Unfortunately, the negative gaps are not recovered, leading to recalls of {{< var local_azulenes_recall_cis >}} as none of five inverted molecules could be identified ([@tbl-local-azulenes]). It could be speculated that the worse performance for the azulenes comes from the fact that they are non-alternants, for which the approximations are expected to be less applicable. Considering the contrasting performance for the azaphenalenes and azulenes, it seems clear that the PPP screening protocol will inevitably miss some hits belonging to particular compound classes. To investigate the global performance of the model we turned to a more diverse set with 16 different compound classes.

![Here's some more text]({{< var fig_local_azulenes >}}.pdf){#fig-local-azulenes}

```{python}
# | label: tbl-local-azulenes
# | tbl-cap: Astronomical object
# | echo: false
import pandas as pd
from IPython.display import display, Markdown

with open(tab_local_azulenes) as f:
    md = f.read()
display(Markdown(md))
```

## Screening widely

Here we used a dataset previously generated by some of us using rational design rules, comprising 16 different molecular scaffolds, the details of which will be presented elsewhere. Out of {{< var global_n_total >}} compounds, only {{< var global_n_failed >}} failed, which corresponds to a success rate of {{< var global_percent_success >}}. The S~1~ and T~1~ excitation energies are fairly reproduced with ([@fig-global-all]a,b) $R^2$ of {{< var global_all_r2_singlet >}} and {{< var global_all_r2_triplet >}} vs {{< var reference_level >}}, respectively. As seen in [@fig-global-all-gaps]a, the $\DeltaE{ST}$ at the SCF + DSP level shows a moderate $R^2$ of {{< var global_all_r2_scf >}}, which increases significantly to {{< var global_all_r2_cis >}} at the CIS + DSP level ([@fig-global-all-gaps]b). However, due to a systematic overestimation of $\DeltaE{ST}$, the F1 score is still low at {{< var global_all_f1_cis >}} and a poor recall of {{< var global_all_recall_cis >}} ([@tbl-global-all]). We therefore added a linear correction $\DeltaEST[DSP]{CIS,LC}=$ {{< var global_lc_coef >}} $\times\DeltaEST[DSP]{CIS}$ {{< var global_lc_intercept >}}, which increases the F1 score to {{< var global_all_f1_cis_corr >}} and the recall to {{< var global_all_recall_cis_corr >}} ([@fig-global-all-gaps]c). The specificity simultaneously decreases from {{< var global_all_specificity_cis >}} to {{< var global_all_specificity_cis_corr >}}, but overall the linear correction would be preferable for virtual screening where there is a stronger focus on finding rare hits. Applying the linear correction to the azaphenalenes ([@fig-local-corr]a) and ([@fig-local-corr]b) leads to an improvement in the recall of the former from {{< var local_azaphenalenes_recall_cis >}} to {{< var local_azaphenalenes_recall_cis_corr >}} while still none of the inverted azuelenes are recovered. The oscillator strengths are unfortunately poorly correlated for the rational design set, with $R^2=$ {{< var global_all_r2_oscillator >}}, although the Spearman rank correlation is more encouraging at {{< var global_all_rho_oscillator >}} ([@fig-global-all]c). Taken together, the results reinforce the conclusions from the case study of the azaphenalenes, that the PPP protocol is suitable for screening $\DeltaE{ST}$ while it struggles for oscillator strengths.

A more detailed breakdown of the correlations into the different scaffolds shows that the highest $R^2$ of {{< var global_scaffolds_r2_max >}} is obtained for {{< var global_scaffolds_r2_max_name >}}, while the lowest is obtained for {{< var global_scaffolds_r2_min_name >}} with {{< var global_scaffolds_r2_min >}}. For recall, the corresponding range is {{< var global_scaffolds_recall_max >}} for {{< var global_scaffolds_recall_max_name >}} to {{< var global_scaffolds_recall_min >}} for {{< var global_scaffolds_recall_min_name >}}. The wide range in $R^2$ and recall further reinforces that the method struggles with some particular compounds classes despite its favorable global performance. Despite these problems, the method is able to capture inverted molecules in {{< var global_scaffolds_n_captured >}} out of the 15 scaffolds where they occur. Even for the azulenes included in this screening set at least some candidates are found, in contrast to the findings above for the azaazulenes. The per-scaffold oscillator strengths are arguably sufficiently good to allow for local screening in some cases ([@fig-global-scaffolds-oscillator]). The per-scaffold S~1~ and T~1~ excitation energies can be found in ([@fig-global-scaffolds-singlet]) and ([@fig-global-scaffolds-triplet]), respectively. Notable failures are for {{< var I >}} with $R^2$ of {{< var global_scaffolds_singlet_r2_I >}} and {{< var global_scaffolds_triplet_r2_I >}}, respectively, and {{< var XII >}} and {{< var XVI >}}, where some negative T~1~ excitation energies indicate restricted-unrestricted instabilities.

![Write some text here]({{< var fig_global_all_gaps >}}.pdf){#fig-global-all-gaps}

```{python}
# | label: tbl-global-all
# | tbl-cap: Astronomical object
# | echo: false
import pandas as pd
from IPython.display import display, Markdown

with open(tab_global_all_gaps) as f:
    md = f.read()
display(Markdown(md))
```

## External validation

To further test the validity of the method, we calculated a series of compounds recently published by Garner _et. al_.[@garnerDoublebondDelocalizationNonalternant2023] The dataset comes in three parts: (1) non-alternants, (2) non-alternants with constrained high-symmetry geometires (we here follow the original terminology and call these "avoided symmetry") and (3) substituted non-alternants (here called "substituted"). Out of the subsituted compounds, four could not be calculated as they contain four-coordinate P atoms for which our PPP model lacks parameters. We here opted to use the {{< var reference_level_garner >}}/aug-cc-pVDZ data from the original paper as reference as it is the highest level which has the most complete coverage of the dataset.

Our computed $\DeltaE{ST}$ with the linearly corrected CIS + DSP show fair correlations with the {{< var reference_level_garner >}} values, with $R^2$s of {{< var garner_gaps_r2_non_alternants >}}, {{< var garner_gaps_r2_avoided_symmetry >}} and {{< var garner_gaps_r2_substituted >}}, respectively ([@fig-garner-gaps]). Gratifyingly, the recalls are {{< var garner_gaps_recall_non_alternants >}}, {{< var garner_gaps_recall_avoided_symmetry >}} and {{< var garner_gaps_recall_substituted >}}, respectively, showing that we can recover a large part of the inverted molecules found with the much more expensive {{< var reference_level_garner >}} method ([@tbl-garner-gaps]). Although some of the compounds are also present in the rationally designed set that we used for the linear correction above, we believe that they are not sufficiently many to compromise the use of the Garner data as external validation ([@fig-garner-gaps]). We also indicate in the plots which compounds show instabilities ([@fig-garner-gaps]). There seems to be no clear deterioration in the performance, although also the reference data might exhibit instabilities as there is no mention of any stability analysis in the original manuscript.[@garnerDoublebondDelocalizationNonalternant2023]

In summary, the external validation shows that the PPP method can recover inverted molecules at a fraction of the cost of more expensive methods such as {{< var reference_level_garner >}}.

![Write some text here]({{< var fig_garner_gaps >}}.pdf){#fig-garner-gaps}

```{python}
# | label: tbl-garner-gaps
# | tbl-cap: Astronomical object
# | echo: false
import pandas as pd
from IPython.display import display, Markdown

with open(tab_garner_gaps) as f:
    md = f.read()
display(Markdown(md))
```

# Conclusions and outlook

To conclude, we have shown that the Pariser-Parr-Pople theory, a simple semi-empirical $\pi$-electron theory with a minimal basis, can be used to screen for molecules with inverted singlet-triplets energy gaps, both locally and globally. Unfortunately, the method does not seem capable of screening for oscillator strengths with the same accuracy, with exceptions for some scaffolds. The chief limitation of the method is that it only includes the $\pi$-electron system and therefore struggles with: (1) functional groups without a clear $\sigma$--$\pi$ separation (found in common functional groups such as sulfonyls) (2) inductive effects and (3) neglect of $n \rightarrow \pi^*$ transitions. Further limitations include (4) lack of solvation model and need for (5) already optimized geometries. We believe that at at least some of these limitations could be mitigated by a reparametrization specifically targeting inverted gaps, while geometries from the fast GFN-xTB method would be suitable.

Taken together, we foresee that the presented methodology can be used for ultrahigh-throughput virtual screening campaigns and in inverse design algorithms, followed by curation of hits using more accurate quantum-chemical methods. Active learning schemes could also be used with machine learning corrections to the PPP singlet-triplet gaps. The method has great potential to accelerate the discovery of the next generation of OLED materials based on INVEST.

# Acknowledgements

A. A.-G. acknowledges support of Natural Resources Canada and the Canada 150 Research Chairs program and CIFAR, as well as the generous support of Anders G. Frøseth.

# References

::: {#refs}
:::

\newpage
\setcounter{page}{1}
\setcounter{section}{0}
\setcounter{figure}{0}
\setcounter{table}{0}
\renewcommand{\thepage}{S\arabic{page}}
\renewcommand{\thesection}{S\arabic{section}}
\renewcommand{\thetable}{S\arabic{table}}
\renewcommand{\thefigure}{S\arabic{figure}}

# Supporting information

Hello world

![Write some text here]({{< var fig_global_all >}}.pdf){#fig-global-all}

![Write some text here]({{< var fig_global_scaffolds_gaps >}}.png){#fig-global-scaffolds-gaps}

![Write some text here]({{< var fig_global_scaffolds_singlet >}}.png){#fig-global-scaffolds-singlet}

![Write some text here]({{< var fig_global_scaffolds_triplet >}}.png){#fig-global-scaffolds-triplet}

![Write some text here]({{< var fig_global_scaffolds_oscillator >}}.png){#fig-global-scaffolds-oscillator}

![Write some text here]({{< var fig_local_corr >}}.pdf){#fig-local-corr}

```{python}
# | label: tbl-global-scaffolds
# | tbl-cap: Astronomical object
# | echo: false
import pandas as pd
from IPython.display import display, Markdown

with open(tab_global_scaffolds) as f:
    md = f.read()
display(Markdown(md))
```

### 2023-07-28 workflow/notebooks/calculate.py

@@ -1,0 +1,87 @@
+import time
+from os import PathLike
+
+import numpy as np
+import pandas as pd
+from coulson.interface import mol_from_xyz, process_rdkit_mol
+from coulson.ppp import (
+    PPPCalculator,
+    calculate_dsp,
+    calculate_exchange,
+    homo_lumo_overlap,
+)
+
+
+def calculate_compound(filename: str | PathLike) -> dict[str, list[int]]:
+    """Calculate excitation roperties of molecule."""
+    # Process filename with RDKit mol detection from XYZ
+    start_time = time.time()
+    mol = mol_from_xyz(filename)
+    input_data, mask = process_rdkit_mol(mol)
+    if not any(mask) is True:
+        raise ValueError("Molecule does not have pi system.")
+
+    # Perform SCF calculation
+    ppp = PPPCalculator(input_data)
+    ppp.scf(max_iter=500)
+
+    # Calcualte exchange integral
+    exchange = calculate_exchange(ppp)
+
+    # Calculate HOMO-LUMO overlap
+    overlap = homo_lumo_overlap(ppp)
+
+    # Calculate S1 energies and oscillator strengths
+    ppp.ci(n_states=3)
+    energy_s1_cis = ppp.ci_energies[0] - ppp.electronic_energy
+    oscillator_strength = ppp.oscillator_strengths[0]
+
+    # Calculate T1 energies
+    ppp.ci(n_states=3, multiplicity="triplet")
+    energy_t1_cis = ppp.ci_energies[0] - ppp.electronic_energy
+
+    # Calculate DSP contribution
+    dsp_scf, _ = calculate_dsp(ppp)
+    dsp_cis, _ = calculate_dsp(
+        ppp, ci=True, energy_s_1=energy_s1_cis, energy_t_1=energy_t1_cis
+    )
+    end_time = time.time()
+    run_time = end_time - start_time
+
+    results = {
+        "s1_cis": [energy_s1_cis],
+        "t1_cis": [energy_t1_cis],
+        "exchange_integral": [exchange],
+        "dsp_scf": [dsp_scf],
+        "dsp_cis": [dsp_cis],
+        "homo_lumo_overlap": [overlap],
+        "oscillator_strength": [oscillator_strength],
+        "run_time": run_time,
+    }
+
+    return results
+
+
+def main():
+    # Perform calculation
+    try:
+        results = calculate_compound(snakemake.input[0])
+    except ValueError:
+        results = {
[…]

### 2023-07-28 workflow/notebooks/utils.py

@@ -1,0 +1,409 @@
+import json
+import re
+from io import BytesIO
+from numbers import Integral, Real
+from os import PathLike
+from typing import Any
+
+import matplotlib as mpl
+import matplotlib.pyplot as plt
+import numpy as np
+import pandas as pd
+import PIL
+import scipy.stats
+import sklearn.metrics
+from coulson.data import HARTREE_TO_EV
+from sklearn.linear_model import LinearRegression
+
+AXIS_LABELS: dict[str, str] = {
+    "t1_s1_ref": r"$\Delta E_{\mathrm{ST,REF}}$ (eV)",
+    "t1_s1_ppp": r"$\Delta E_{\mathrm{ST,CIS}}$ (eV)",
+    "t1_s1_dsp_scf": r"$\Delta E_{\mathrm{ST,SCF}}^{\mathrm{DSP}}$ (eV)",
+    "t1_s1_dsp_cis": r"$\Delta E_{\mathrm{ST,CIS}}^{\mathrm{DSP}}$ (eV)",
+    "t1_s1_dsp_cis_corr": r"$\Delta E_{\mathrm{ST,CIS,corr}}^{\mathrm{DSP}}$ (eV)",
+    "exchange": r"$K$ (eV)",
+    "dsp_scf": r"$\Delta E_{\mathrm{SCF}}^{\mathrm{DSP}}$ (eV)",
+    "dsp_cis": r"$\Delta E_{\mathrm{CIS}}^{\mathrm{DSP}}$ (eV)",
+    "homo_lumo_overlap": r"$O_\mathrm{HL}$",
+    "oscillator_strength": r"$f_\mathrm{CIS}$",
+    "oscillator_strength_ref": r"$f_\mathrm{REF}$",
+    "s1_cis": r"$\Delta E_\mathrm{S,CIS}} (eV)$",
+    "s1_ref": r"$\Delta E_\mathrm{S,REF}} (eV)$",
+    "t1_cis": r"$\Delta E_\mathrm{T,CIS}} (eV)$",
+    "t1_ref": r"$\Delta E_\mathrm{T,REF}} (eV)$",
+    "t1_s1_cc2_aug_dz": r"$\Delta E_{\mathrm{ST,REF}}$ (eV)",
+    "t1_s1_cc2_dz": r"$\Delta E_{\mathrm{ST,REF}}$ (eV)",
[…]

### 2023-07-28 workflow/scripts/calculate.py

@@ -1,50 +1,0 @@
-from coulson.interface import process_rdkit_mol, mol_from_xyz
-from coulson.ppp import (
-    calculate_exchange,
-    calculate_dsp,
-    homo_lumo_overlap,
-    PPPCalculator,
-)
-import pandas as pd
-
-idx = snakemake.wildcards.sample
-
-mol = mol_from_xyz(snakemake.input[0])
-input_data, mask = process_rdkit_mol(mol)
-ppp = PPPCalculator(input_data)
-ppp.scf(max_iter=500)
-
-# Calcualte exchange integral
-exchange = calculate_exchange(ppp)
-
-# Calculate HOMO-LUMO overlap
-overlap = homo_lumo_overlap(ppp)
-
-# Calculate S1 energies and oscillator strengths
-ppp.ci(n_states=3)
-energy_s1_cis = ppp.ci_energies[1] - ppp.ci_energies[0]
-
-oscillator_strength = ppp.oscillator_strengths[0]
-
-# Calculate T1 energies
-ppp.ci(n_states=3, multiplicity="triplet")
-energy_t1_cis = ppp.ci_energies[1] - ppp.ci_energies[0]
-
-# Calculate DSP contribution
-dsp_scf, excitations = calculate_dsp(ppp)
-dsp_cis, excitations = calculate_dsp(
-    ppp, ci=True, energy_s_1=energy_s1_cis, energy_t_1=energy_t1_cis
-)
-
-data = {
-    "id": [idx],
-    "s1_cis": [energy_s1_cis],
-    "t1_cis": [energy_t1_cis],
-    "exchange_integral": [exchange],
-    "dsp_scf": [dsp_scf],
-    "dsp_cis": [dsp_cis],
-    "homo_lumo_overlap": [overlap],
-    "oscillator_strength": [oscillator_strength],
-}
-df = pd.DataFrame(data).set_index("id")
-df.to_csv(snakemake.output[0])

### 2023-07-28 manuscript/manuscript.qmd
Clean up repo from snakemake template

@@
 title: "Ultrafast computational screening of molecules with inverted singlet-triplet energy gaps using semi-empirical quantum chemistry"
+abstract: |
+  Molecules with an inverted energy gap between their first excited states of singlet and triplet multiplicity have promising applications in the next generation of organic light-emitting diode (OLED) materials.
+Unfortunately, such molecules are rare and only a handful of examples are currently known.
+High-throughput virtual screening could assist in finding novel classes of molecules, but current efforts are hampered by the high computational cost of the required quantum-chemical methods.
+We present a method based on the semi-empirical Pariser-Parr-Pople theory augmented by perturbation theory and show that it reproduces inverted gaps at a fraction of the cost of the current employed methods.
+Our study opens the doors for ultra-high-throughput virtual screening and inverse design to accelerate the discovery and development of the new generation of OLED materials.
 bibliography: [references.bib] 
@@
 As the transition T~1~ $\rightarrow$ S~0~, from the lowest excited triplet state to the singlet ground state, is quantum-mechanically forbidden, the OLED molecule can formally only emit from its first excited state of singlet multiplicity, S~1~.
-The second generation of OLEDs exploited ways of increasing the rate of this spin-forbidden phosphoresence.
+The second generation of OLEDs exploited ways of increasing the rate of this spin-forbidden phosphorescence.
 The third generation of OLEDs are based on Thermally Activated Delayed Fluorescence (TADF), where the S~1~ state is partially populated from a near-lying T~1~ state.
-For this to happpen, the gap $\DeltaE{ST}$ between the states needs to sufficiently small so that the thermal activation competes with non-radiative decay processes from T~1~.
+For this to happen, the gap $\DeltaE{ST}$ between the states needs to sufficiently small so that the thermal activation competes with non-radiative decay processes from T~1~.
 However, as the T~1~ population is still substantial, problems with stability and lower than ideal intensity due to non-radiative decay persist.
@@
 Motivated by the need for faster methods for HTVS of INVEST compounds, we wondered if it would be possible to perform much simpler calculations as a pre-screening step for the more expensive methods.
-Based on prior work in the literature from the 1970s and 1980s, we find that the inverted gaps can be recovered already with the semi-empirical Pariser-Parr-Pople method using configuaration interaction singles (CIS) and adding key double excitations.
-The PPP method considers only the $\pi$ electrons in a minimal valence basis and approximates all integrals from experimental data and a few empirical parameters, making it computationally extremly cheap.
+Based on prior work in the literature from the 1970s and 1980s, we find that the inverted gaps can be recovered already with the semi-empirical Pariser-Parr-Pople method using configuration interaction singles (CIS) and adding key double excitations.
+The PPP method considers only the $\pi$ electrons in a minimal valence basis and approximates all integrals from experimental data and a few empirical parameters, making it computationally extremely cheap.
 At the same time, it retains the conceptual clarity of Hückel theory and allows for straightforward interpretation of the inverted gap in terms of the well-known concept of dynamic spin polarization (DSP).
@@
 
-where $C = 0.545$ is a parameter that was fit in the original publicaiton to reproduce experimental excitation energies.[@beveridgeParametrizationSemiempiricalPi1971] The overlap integral $S_{rs}$ is determined from the overlap of Slater *p* orbitals with exponents
+where $C = 0.545$ is a parameter that was fit in the original publication to reproduce experimental excitation energies.[@beveridgeParametrizationSemiempiricalPi1971] The overlap integral $S_{rs}$ is determined from the overlap of Slater *p* orbitals with exponents
 
@@
 Borden and Davidson explained this effect in terms of dynamic spin polarization (DSP),[@bordenTheoreticalStudiesDiradicals1981] in which the electrons of a pair of disjoint non-bonding molecular orbitals (NBMOs, [@fig-dsp]a) experience stronger electron correlation effects in the singlet state as compared to the triplet state ([@fig-dsp]b).
-As outlined by Kollmar and Staemmler,[@kollmarViolationHundRule1978] as well as Malrieu and co-workers,[@benamorSpinPolarizationElectronic2020] an alternative view of the same phenomenon is through admixture of singly-excited configurations into the electronic wavefunction of the ground state ([@fig-dsp]c).
+For a pedagogic introduction to the topic, the reader is referred to an excellent article by Karafiloglou[@karafiloglouDoubleDynamicSpin1989] As outlined by Kollmar and Staemmler,[@kollmarViolationHundRule1978] as well as Malrieu and co-workers,[@benamorSpinPolarizationElectronic2020] an alternative view of the same phenomenon is through admixture of singly-excited configurations into the electronic wavefunction of the ground state ([@fig-dsp]c).
 The DSP effect for the S~1~ excited state can correspondingly be described by admixture of doubly excited configurations,[@kollmarViolationHundRule1978] as they are singly excited with respect to the S~1~ state, and doubly excited with respect to the S~0~ ground state ([@fig-dsp]d).
@@
 
-![Test to write something here](../data/figures/dynamic_spin_polarization.pdf){#fig-dsp}
+![Dynamic spin polarization stabilizes the open-shell singlet state over the open-shell triplet.
+(a) Transformation of the canonical frontier molecular orbitals creates a set of disjoint non-bonding molecular orbitals for cyclobutadiene.
+Singlet and triplet occupations are shown.
+(b) Dynamic spin polarization preferentially stabilizes the singlet state of cyclobutadiene through additive spin polarization for the singlet and competitive spin polarization for the triplet.
+(c) An alternative description of the spin polarization phenomenon is via configuration interaction and admixture of excited configurations into the ground state.
+(d) In the same way as cyclobutadiene is stabilized by DSP in the singlet ground state, molecules with inverted singlet-triplet gaps in the excited state are also stabilized by DSP that can be described by admixture of excited configurations.
+For singly excited states, this means addition of doubly excited states.](../data/figures/dynamic_spin_polarization.pdf){#fig-dsp}
 
@@
 We further integrated [coulson]{.smallcaps} with [PySCF]{.smallcaps}[@sunRecentDevelopmentsPySCF2020] to provide alternative algorithms for the SCF convergence and CIS and found the results consistent within numerical accuracy.
-Oscillator strengths were calculated with the dipole length approximation.[@jorgensenTimedependentHartreeFockCalculations1970;@klessingerExcitedStatesPhotochemistry1995;@damourGroundExcitedStateDipole2023] Some compounds exhibit negative triplet exitation energies and in some case even negative singlet excitation energies, which indicates a restricted-unrestricted instability of the ground state wavefunction.
+Oscillator strengths were calculated with the dipole length approximation.[@jorgensenTimedependentHartreeFockCalculations1970;@klessingerExcitedStatesPhotochemistry1995;@damourGroundExcitedStateDipole2023] Some compounds exhibit negative triplet excitation energies and in some case even negative singlet excitation energies, which indicates a restricted-unrestricted instability of the ground state wavefunction.
 We have here taken the pragmatic approach to ignore these issues, while we will highlight below some examples where it occurs and if it has any effect on the overall results.
@@
 The results should anyway be indicative of the gap between the lowest singlet and triplet states with open-shell character, regardless of if they are lower in energy than the closed-shell singlet.
+
+We used four separate datasets for this study.
+The first is a set of azaphenalenes previously studied by some of us, comprising 257 substituted compounds, for which excitation energies were computed with (2nd-order) algebraic-diagrammatic construction, {{< var reference_level >}}/[Robert: Fill out], and geometries were optimized with B3LYP/[Robert: Fill out][@polliceOrganicMoleculesInverted2021] The second is a set substituted 138 aza-azulenes with {{< var reference_level >}}/[Robert: Fill out] excitation energies and optimized with the B97-3c composite method.[Robert: Add ref to dataset if it exists.] The third is a set of 16 rationally designed scaffolds, which including substitutions amounts to {{< var global_n_total >}} unique compounds, optimized at the B97-3c level and with excitation energies at the {{< var reference_level >}}/[Robert: Fill out] level.[Robert: Add ref to dataset if it exists.] The fourth is a set 315 non-alternant hydrocarbons including substitutions, divided into three subsets of size 76, 187 and 52, optimized at the ωB97X-D/def2-TZVP level and with excitation energies at the CC2/aug-cc-pVTZ level.
 
@@
 We first demonstrate that our method is capable of capturing the inverted $\DeltaE{ST}$ for some model compounds.
-As shown by Toyota and co-workers, pentalene at the ideal  $D_\mathrm{2h}$ geometry has an inverted gap, while relaxation to the minimum with $C_\mathrm{2h}$ symmetry leads to a normal gap.[@kosekiViolationHundMultiplicity1985] In our calculations, pentalene with equal bond lenghts of 1.4 Å displays a small HOMO-LUMO overlap of {{< var pentalene_ideal_overlap >}}, leading to a small exchange interaction of only $2K=$ {{< var pentalene_ideal_2_exchange >}} kcal/mol (@fig-pentalene-ideal-orbitals).
+As shown by Toyota and co-workers, pentalene at the ideal  $D_\mathrm{2h}$ geometry has an inverted gap, while relaxation to the minimum with $C_\mathrm{2h}$ symmetry leads to a normal gap.[@kosekiViolationHundMultiplicity1985] In our calculations, pentalene with equal bond lengths of 1.4 Å displays a small HOMO-LUMO overlap of {{< var pentalene_ideal_overlap >}}, leading to a small exchange interaction of only $2K=$ {{< var pentalene_ideal_2_exchange >}} kcal/mol (@fig-pentalene-ideal-orbitals).
 The dynamic spin polarization correction $\DeltaE[DSP]{SCF}$ of {{< var pentalene_ideal_dsp_scf >}} kcal/mol leads to a net predicted $\DeltaEST[DSP]{SCF}$ of {{< var pentalene_ideal_t1_s1_dsp_scf >}} kcal/mol.
@@
 The reference value calculated with {{< var reference_level >}} is {{< var azaphenalene_t1_s1_ref >}} kcal/mol.
-The simple perturbation theory model allows us to gain additional insight into the contributions to $\DeltaE[DSP]{SCF}$ as they correspond to single excitations between doubly occupied orbitals below the HOMO and uncoccupied orbitals above the LUMO (@fig-pentalene-ideal-excitations).
+The simple perturbation theory model allows us to gain additional insight into the contributions to $\DeltaE[DSP]{SCF}$ as they correspond to single excitations between doubly occupied orbitals below the HOMO and unoccupied orbitals above the LUMO (@fig-pentalene-ideal-excitations).
 Three excitations, HOMO-3 → LUMO+2, HOMO-2 → LUMO+1 and HOMO-1 → LUMO+3 are identified as the main contributors to the spin polarization effect, and could potentially be tuned by substituent effects.
@@
 The larger azaphenalene has 30 excitations that could possibly contribute to the large $\DeltaE[DSP]{SCF}$ of {{< var azaphenalene_dsp_scf >}} kcal/mol, but the most import by far are those from the doubly degenerate HOMO-1 to the doubly degenerate LUMO+1 ([@fig-azaphenalene-excitations]).
-In total, $\DeltaEST[DSP]{SCF}=$ {{< var azaphenalene_t1_s1_dsp_scf >}} kcal/mol, which is slighly smaller at the CIS level with $\DeltaEST[DSP]{CIS}=$ {{< var azaphenalene_t1_s1_dsp_cis >}} kcal/mol.
+In total, $\DeltaEST[DSP]{SCF}=$ {{< var azaphenalene_t1_s1_dsp_scf >}} kcal/mol, which is slightly smaller at the CIS level with $\DeltaEST[DSP]{CIS}=$ {{< var azaphenalene_t1_s1_dsp_cis >}} kcal/mol.
 
@@
 
-While our method captures the inverted gap of azaphenalene and pentalene, it also needs to capture trends with substition to work effectively in virtual screening.
+While our method captures the inverted gap of azaphenalene and pentalene, it also needs to capture trends with substitution to work effectively in virtual screening.
 We therefore calculated 256 substituted azaphenalenes that have previously been studied by some of us.[@polliceOrganicMoleculesInverted2021] The PPP level S~1~ and T~1~ excitation energies are well correlated with those from {{< var reference_level >}} with $R^2$ of {{< var local_azaphenalenes_r2_singlet >}} and {{< var local_azaphenalenes_r2_triplet >}}, respectively ([@fig-local-azaphenalenes]a,b).
-Unfortunately, the oscillator strenghts are not as well reproduced with an $R^2$ of {{< var local_azaphenalenes_r2_f >}}, although the Spearman $\rho$ of {{< var local_azaphenalenes_rho_f >}} indicates the values might be used to rank potential candidates ([@fig-local-azaphenalenes]c).
-The problems of obtaining accurate oscillator strenghts at the PPP level of theory are well known in the literature, and the problem is especially exacerbated with low oscillator strengths as here.[@misurkinElectronicStructureConjugated1974].
+Unfortunately, the oscillator strengths are not as well reproduced with an $R^2$ of {{< var local_azaphenalenes_r2_f >}}, although the Spearman $\rho$ of {{< var local_azaphenalenes_rho_f >}} indicates the values might be used to rank potential candidates ([@fig-local-azaphenalenes]c).
+The problems of obtaining accurate oscillator strengths at the PPP level of theory are well known in the literature, and the problem is especially exacerbated with low oscillator strengths as here.[@misurkinElectronicStructureConjugated1974].
 Crucially, the $\DeltaE{ST}$ shows good correlations, with $R^2=$ {{< var local_azaphenalenes_r2_cis >}} using $\DeltaEST[DSP]{CIS}$ ([@fig-local-azaphenalenes]f).
@@
 The most important metrics for virtual screening are, in our opinion, the recall and the specificity.
-The recall mesaures the fraction of inverted molecules that the protocol captures, and is {{< var local_azaphenalenes_recall_scf >}} at the SCF level and {{< var local_azaphenalenes_recall_cis >}} at the CIS level.
-The specificity measurs the freaction of non-inverted molecules that the protocol identifies, and is {{< var local_azaphenalenes_specificity_scf >}} at the SCF level and {{< var local_azaphenalenes_specificity_cis >}} at the CIS level.
+The recall measures the fraction of inverted molecules that the protocol captures, and is {{< var local_azaphenalenes_recall_scf >}} at the SCF level and {{< var local_azaphenalenes_recall_cis >}} at the CIS level.
+The specificity measures the fraction of non-inverted molecules that the protocol identifies, and is {{< var local_azaphenalenes_specificity_scf >}} at the SCF level and {{< var local_azaphenalenes_specificity_cis >}} at the CIS level.
 This metrics mean that we are able to capture a large proportion of the molecules with inverted gaps, while filtering out most non-inverted molecules.
@@
 While we see success for the azaphenalenes, the situation for the azulenes is unfortunately worse.
-The PPP level S~1~ and T~1~ excitation energies are only moderaly correlated with those from {{< var reference_level >}} with $R^2$ of {{< var local_azulenes_r2_singlet >}} and {{< var local_azulenes_r2_triplet >}}, respectively ([@fig-local-azulenes]a,b).
+The PPP level S~1~ and T~1~ excitation energies are only moderately correlated with those from {{< var reference_level >}} with $R^2$ of {{< var local_azulenes_r2_singlet >}} and {{< var local_azulenes_r2_triplet >}}, respectively ([@fig-local-azulenes]a,b).
 For the oscillator strengths, the situation is rather catastrophic with $R^2$ of {{< var local_azulenes_r2_f >}} implying no correlation whatsoever ([@fig-local-azulenes]c).
@@
 Out of {{< var global_n_total >}} compounds, only {{< var global_n_failed >}} failed, which corresponds to a success rate of {{< var global_percent_success >}}.
+The total runtime using four cores on the M2 processor was {{< var time_cpu_hours >}} CPU hours.
 The S~1~ and T~1~ excitation energies are fairly reproduced with ([@fig-global-all]a,b) $R^2$ of {{< var global_all_r2_singlet >}} and {{< var global_all_r2_triplet >}} vs {{< var reference_level >}}, respectively.
@@
 The specificity simultaneously decreases from {{< var global_all_specificity_cis >}} to {{< var global_all_specificity_cis_corr >}}, but overall the linear correction would be preferable for virtual screening where there is a stronger focus on finding rare hits.
-Applying the linear correction to the azaphenalenes ([@fig-local-corr]a) and ([@fig-local-corr]b) leads to an improvement in the recall of the former from {{< var local_azaphenalenes_recall_cis >}} to {{< var local_azaphenalenes_recall_cis_corr >}} while still none of the inverted azuelenes are recovered.
+Applying the linear correction to the azaphenalenes ([@fig-local-corr]a) and ([@fig-local-corr]b) leads to an improvement in the recall of the former from {{< var local_azaphenalenes_recall_cis >}} to {{< var local_azaphenalenes_recall_cis_corr >}} while still none of the inverted azulenes are recovered.
 The oscillator strengths are unfortunately poorly correlated for the rational design set, with $R^2=$ {{< var global_all_r2_oscillator >}}, although the Spearman rank correlation is more encouraging at {{< var global_all_rho_oscillator >}} ([@fig-global-all]c).
@@
 To further test the validity of the method, we calculated a series of compounds recently published by Garner _et.
-al_.[@garnerDoublebondDelocalizationNonalternant2023] The dataset comes in three parts: (1) non-alternants, (2) non-alternants with constrained high-symmetry geometires (we here follow the original terminology and call these "avoided symmetry") and (3) substituted non-alternants (here called "substituted").
-Out of the subsituted compounds, four could not be calculated as they contain four-coordinate P atoms for which our PPP model lacks parameters.
+al_.[@garnerDoublebondDelocalizationNonalternant2023] The dataset comes in three parts: (1) non-alternants, (2) non-alternants with constrained high-symmetry geometries (we here follow the original terminology and call these "avoided symmetry") and (3) substituted non-alternants (here called "substituted").
+Out of the substituted compounds, four could not be calculated as they contain four-coordinate P atoms for which our PPP model lacks parameters.
 We here opted to use the {{< var reference_level_garner >}}/aug-cc-pVDZ data from the original paper as reference as it is the highest level which has the most complete coverage of the dataset.
@@
 
-A. A.-G. acknowledges support of Natural Resources Canada and the Canada 150 Research Chairs program and CIFAR, as well as the generous support of Anders G. Frøseth.
+K.J.
+acknowledges funding through an International Postdoc grant from the Swedish Research Council (No.
+2020-00314).
+A.
+A.-G.
+acknowledges support of Natural Resources Canada and the Canada 150 Research Chairs program and CIFAR, as well as the generous support of Anders G.
+Frøseth.
 
@@
 
-Hello world
+![Singlet and triplet excitation energies as well as oscillator strengths for the rationally designed dataset *vs.* the reference level.]({{< var fig_global_all >}}.pdf){#fig-global-all}
 
-![Write some text here]({{< var fig_global_all >}}.pdf){#fig-global-all}
+![Singlet-triplet gap split by scaffold for the rationally designed dataset *vs.* the reference level.]({{< var fig_global_scaffolds_gaps >}}.png){#fig-global-scaffolds-gaps}
 
-![Write some text here]({{< var fig_global_scaffolds_gaps >}}.png){#fig-global-scaffolds-gaps}
+![Singlet excitation energies split by scaffold for the rationally designed dataset *vs.* the reference level.]({{< var fig_global_scaffolds_singlet >}}.png){#fig-global-scaffolds-singlet}
 
-![Write some text here]({{< var fig_global_scaffolds_singlet >}}.png){#fig-global-scaffolds-singlet}
+![Triplet excitation energies split by scaffold for the rationally designed dataset *vs.* the reference level.]({{< var fig_global_scaffolds_triplet >}}.png){#fig-global-scaffolds-triplet}
 
-![Write some text here]({{< var fig_global_scaffolds_triplet >}}.png){#fig-global-scaffolds-triplet}
+![Oscillator strengths split by scaffold for the rationally designed dataset *vs.* the reference level.]({{< var fig_global_scaffolds_oscillator >}}.png){#fig-global-scaffolds-oscillator}
 
-![Write some text here]({{< var fig_global_scaffolds_oscillator >}}.png){#fig-global-scaffolds-oscillator}
-
-![Write some text here]({{< var fig_local_corr >}}.pdf){#fig-local-corr}
+![Singlet-triplet gaps with CIS + DSP and linear correction for the (a) azaphenalenes and (b) azaazulenes *vs.* the reference level.]({{< var fig_local_corr >}}.pdf){#fig-local-corr}
 
@@
 # | label: tbl-global-scaffolds
-# | tbl-cap: Astronomical object
+# | tbl-cap: Metrics split by scaffold for the rationally designed dataset.
 # | echo: false

### 2023-07-29 manuscript/manuscript.qmd
Update

@@
     orcid: 0000-0001-8836-6266
+    email: "r.pollice@rug.nl"
+    corresponding: true
     affiliations: 
@@
 
-![Write some text here]({{< var fig_local_azaphenalenes >}}.pdf){#fig-local-azaphenalenes}
+![Excitation properties for azaphenalenes _vs._ the reference level.
+(a) Singlet excitation energies.
+(b) Triplet excitation energies.
+(c) Oscillator strengths.
+(d) Singlet-triplet energy gaps with CIS.
+(e) Singlet-triplet energy gaps with SCF + DSP.
+(f) Singlet-triplet energy gaps with CIS + DSP.]({{< var fig_local_azaphenalenes >}}.pdf){#fig-local-azaphenalenes}
 
@@
 # | label: tbl-local-azaphenalenes
-# | tbl-cap: Astronomical object
+# | tbl-cap: Metrics for azaphenalenes.
 # | echo: false
@@
 
-![Here's some more text]({{< var fig_local_azulenes >}}.pdf){#fig-local-azulenes}
+![Excitation properties for azaazulenes _vs._ the reference level.
+(a) Singlet excitation energies.
+(b) Triplet excitation energies.
+(c) Oscillator strengths.
+(d) Singlet-triplet energy gaps with CIS.
+(e) Singlet-triplet energy gaps with SCF + DSP.
+(f) Singlet-triplet energy gaps with CIS + DSP.]({{< var fig_local_azulenes >}}.pdf){#fig-local-azulenes}
 
@@
 # | label: tbl-local-azulenes
-# | tbl-cap: Astronomical object
+# | tbl-cap: Metrics for azaazulenes.
 # | echo: false
@@
 
-![Write some text here]({{< var fig_global_all_gaps >}}.pdf){#fig-global-all-gaps}
+![Singlet-triplet energy gaps for rationally design set _vs._ reference level for (a) SCF + DSP, (b) CIS + DSP and (c) CIS + DSP and linear correction.]({{< var fig_global_all_gaps >}}.pdf){#fig-global-all-gaps}
 
@@
 # | label: tbl-global-all
-# | tbl-cap: Astronomical object
+# | tbl-cap: Metrics for rationally designed set.
 # | echo: false
@@
 
-![Write some text here]({{< var fig_garner_gaps >}}.pdf){#fig-garner-gaps}
+![Singlet-triplet energy gaps for CIS + DSP and linear correction _vs._ reference level for the subsets of (a) non-alternants (b) non-alternants constrained to higher symmetry and (c) substituted non-alternants.
+Compounds contained in the rationally designed sets are marked with dashed circles and those with negative triplet excitation energies are marked with a semi-transparent red circle.]({{< var fig_garner_gaps >}}.pdf){#fig-garner-gaps}
 
@@
 # | label: tbl-garner-gaps
-# | tbl-cap: Astronomical object
+# | tbl-cap: Metrics for the dataset from Garner _et al._
 # | echo: false
@@
 
-![Singlet-triplet gap split by scaffold for the rationally designed dataset *vs.* the reference level.]({{< var fig_global_scaffolds_gaps >}}.png){#fig-global-scaffolds-gaps}
+![Singlet-triplet energy gaps split by scaffold for the rationally designed dataset *vs.* the reference level.]({{< var fig_global_scaffolds_gaps >}}.png){#fig-global-scaffolds-gaps}
 
@@
 
-![Singlet-triplet gaps with CIS + DSP and linear correction for the (a) azaphenalenes and (b) azaazulenes *vs.* the reference level.]({{< var fig_local_corr >}}.pdf){#fig-local-corr}
+![Singlet-triplet energy gaps with CIS + DSP and linear correction for the (a) azaphenalenes and (b) azaazulenes *vs.* the reference level.]({{< var fig_local_corr >}}.pdf){#fig-local-corr}
 

### 2023-07-29 manuscript/manuscript.tex
Update

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  number,
  3p]{elsarticle}

\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

% Energy commands
\newcommand{\E}[2][]{E^{\mathrm{#1}}_{\mathrm{#2}}}
\newcommand{\DeltaE}[2][]{\Delta E^{\mathrm{#1}}_{\mathrm{#2}}}
\newcommand{\DeltaEST}[2][]{\Delta E^{\mathrm{#1}}_{\mathrm{ST,#2}}}
\newcommand{\EState}[2][]{E (\phi_{\mathrm{#2}}^{#1})}

% Integrals
\newcommand{\Integral}[1]{(i | K_x #1 K_y | k )^2}
\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[many]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\journal{Journal of Physical Chemistry A}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Ultrafast computational screening of molecules with inverted singlet-triplet energy gaps using semi-empirical quantum chemistry},
  pdfauthor={Kjell Jorner; Robert Pollice; Cyrille Lavigne; Alán Aspuru Guzik},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\setlength{\parindent}{6pt}
\begin{document}

\begin{frontmatter}
\title{Ultrafast computational screening of molecules with inverted
singlet-triplet energy gaps using semi-empirical quantum chemistry}
\author[1,2,3,4]{Kjell Jorner%
\corref{cor1}%
}
 \ead{kjell.jorner@chem.ethz.ch} 
\author[5,3,4]{Robert Pollice%
\corref{cor1}%
}
 \ead{r.pollice@rug.nl} 
\author[3,4]{Cyrille Lavigne%
%
}

\author[3,4,6,7,8,10]{Alán Aspuru Guzik%
%
}

\affiliation[1]{organization={ETH Zürich, Institute of Chemical and
Bioengineering, Department of Chemistry and Applied
Biosciences},addressline={Vladimir-Prelog-Weg
1},city={Zürich},country={Switzerland},countrysep={,},postcode={CH-8093},postcodesep={}}
\affiliation[2]{organization={Chalmers University of
Technology, Department of Chemistry and Chemical
Engineering},addressline={Kemigården
4},city={Gothenburg},country={Sweden},countrysep={,},postcode={SE-41258},postcodesep={}}
\affiliation[3]{organization={University of Toronto, Chemical Physics
Theory Group, Department of Chemistry},addressline={80 St.~George
St.},city={Toronto},country={Canada},countrysep={,},postcode={M5S
3H6},postcodesep={}}
\affiliation[4]{organization={University of Toronto, Department of
Computer Science},addressline={40 St.~George
St.},city={Toronto},country={Canada},countrysep={,},postcode={M5S
2E4},postcodesep={}}
\affiliation[5]{organization={University of Groningen, Stratingh
Institute for Chemistry},addressline={Nijenborgh
4},city={Groningen},country={The
Netherlands},countrysep={,},postcode={9747 AG},postcodesep={}}
\affiliation[6]{organization={University of Toronto, Department of
Chemical Engineering \& Applied Chemistry},addressline={200 College
St.},city={Toronto},country={Canada},countrysep={,},postcode={M5S
3E5},postcodesep={}}
\affiliation[7]{organization={University of Toronto, Department of
Materials Science \& Engineering},addressline={184 College
St.},city={Toronto},country={Canada},countrysep={,},postcode={M5S
3E4},postcodesep={}}
\affiliation[8]{organization={Vector Institute for Artificial
Intelligence},addressline={661 University Ave. Suite
710},city={Toronto},country={Canada},countrysep={,},postcode={M5G
1M1},postcodesep={}}
\affiliation[9]{organization={Lebovic Fellow, Canadian Institute for
Advanced Research
(CIFAR)},city={Toronto},country={Canada},countrysep={,},postcode={M5G
1M1},postcodesep={}}
\affiliation[10]{organization={Acceleration Consortium, University of
Toronto},city={Toronto},country={Canada},countrysep={,},postcodesep={}}

\cortext[cor1]{Corresponding author}

\begin{abstract}
Molecules with an inverted energy gap between their first excited states
of singlet and triplet multiplicity have promising applications in the
next generation of organic light-emitting diode (OLED) materials.
Unfortunately, such molecules are rare and only a handful of examples
are currently known. High-throughput virtual screening could assist in
finding novel classes of molecules, but current efforts are hampered by
the high computational cost of the required quantum-chemical methods. We
present a method based on the semi-empirical Pariser-Parr-Pople theory
augmented by perturbation theory and show that it reproduces inverted
gaps at a fraction of the cost of the current employed methods. Our
study opens the doors for ultra-high-throughput virtual screening and
inverse design to accelerate the discovery and development of the new
generation of OLED materials.
\end{abstract}

\end{frontmatter}
    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[breakable, interior hidden, frame hidden, sharp corners, boxrule=0pt, enhanced, borderline west={3pt}{0pt}{shadecolor}]}{\end{tcolorbox}}\fi

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Organic light-emitting diodes (OLEDs) is a technology for generating
light from electricity using organic
molecules.\textsuperscript{\protect\hyperlink{ref-yangRecentAdvancesOrganic2017}{1}}
The first generation of OLEDs were limited to an efficiency of max 25\%
due to spin statistics. As the transition T\textsubscript{1}
\(\rightarrow\) S\textsubscript{0}, from the lowest excited triplet
state to the singlet ground state, is quantum-mechanically forbidden,
the OLED molecule can formally only emit from its first excited state of
singlet multiplicity, S\textsubscript{1}. The second generation of OLEDs
exploited ways of increasing the rate of this spin-forbidden
phosphorescence. The third generation of OLEDs are based on Thermally
Activated Delayed Fluorescence (TADF), where the S\textsubscript{1}
state is partially populated from a near-lying T\textsubscript{1} state.
For this to happen, the gap \(\DeltaE{ST}\) between the states needs to
sufficiently small so that the thermal activation competes with
non-radiative decay processes from T\textsubscript{1}. However, as the
T\textsubscript{1} population is still substantial, problems with
stability and lower than ideal intensity due to non-radiative decay
persist. The logical step for the next generation of OLEDs is to emit
directly from an S\textsubscript{1} state that lies below the
T\textsubscript{1} state, potentially converting all of the electrically
generated excitons into photons, \emph{i.e.}, 100\% efficiency.

Molecules that fulfill this criterion have been branded with the acronym
INVErted Singlet Triplet energy gap (INVEST) and are exceedingly
rare.\textsuperscript{\protect\hyperlink{ref-liOrganicMoleculesInverted2022}{2}}
They are said to violate Hund's rule of maximum multiplicity, which
describes that the electronic state with the highest spin (in this case
the triplet) should be lowest in energy. While examples of molecules
violating Hund's rule in the excited state have been known since the
1970s and
1980s,\textsuperscript{\protect\hyperlink{ref-kollmarViolationHundRule1978}{3}--\protect\hyperlink{ref-kosekiViolationHundMultiplicity1985}{5}}
it is only recently that they have received considerable attention for
application in
OLEDs.\textsuperscript{\protect\hyperlink{ref-desilvaInvertedSingletTriplet2019}{6}--\protect\hyperlink{ref-aizawaDelayedFluorescenceInverted2022}{8}}
As they are rare, recent efforts have used high-throughput virtual
screening (HTVS) with computational chemistry to identify new compounds
with inverted gaps, focusing on expanding the local chemical space
around specific
scaffolds\textsuperscript{\protect\hyperlink{ref-polliceOrganicMoleculesInverted2021}{7},\protect\hyperlink{ref-aizawaDelayedFluorescenceInverted2022}{8}}
or scanning larger (combinatorially or experimentally derived) datasets
to identify novel
scaffolds.\textsuperscript{\protect\hyperlink{ref-terenceblaskovitsSymmetryInducedSinglet2023}{9}--\protect\hyperlink{ref-garnerDoublebondDelocalizationNonalternant2023}{11}}
These HTVS efforts have been successful at identifying several new
INVEST candidates, some of which have also been synthesized and tested
experimentally.\textsuperscript{\protect\hyperlink{ref-aizawaDelayedFluorescenceInverted2022}{8}}

While this early success of HTVS is highly encouraging, it is hampered
by the cost of the computational methods. Excitation energies are
routinely calculated by time-dependent density functional theory
(TD-DFT), which has become a workhorse of computational photophysics
over the last
decades.\textsuperscript{\protect\hyperlink{ref-casidaProgressTimeDependentDensityFunctional2012}{12}}
Unfortunately, it has been shown that TD-DFT fails to capture the
inverted gap of INVEST compounds, as the method only considers single
electronic
excitations.\textsuperscript{\protect\hyperlink{ref-desilvaInvertedSingletTriplet2019}{6}}
Inclusion of at least double excitations are necessary to reproduce the
inverted gaps, corresponding to methods such as double hybrid TD-DFT or
excited state coupled-cluster methods such as second-order approximate
coupled cluster singles and doubles (CC2) or equation of motion coupled
cluster with single and double excitations (EOM-CCSD). In line with the
results of benchmarking
studies,\textsuperscript{\protect\hyperlink{ref-ricciSingletTripletExcited2021}{13},\protect\hyperlink{ref-sancho-garciaViolationHundRule2022}{14}}
recent HTVS studies have used methods such as the complete active space
self-consistent field (CASSCF) and CC2 for preliminary screening, while
confirming inverted gaps with more expensive methods such as multistate
complete active space second-order perturbation theory (MS-CASPT2) and
EOM-CCSD. These methods are not only expensive, but in the case of the
CAS methods, also require the choice of an orbital active space,
something which is not routinely automatized.

Motivated by the need for faster methods for HTVS of INVEST compounds,
we wondered if it would be possible to perform much simpler calculations
as a pre-screening step for the more expensive methods. Based on prior
work in the literature from the 1970s and 1980s, we find that the
inverted gaps can be recovered already with the semi-empirical
Pariser-Parr-Pople method using configuration interaction singles (CIS)
and adding key double excitations. The PPP method considers only the
\(\pi\) electrons in a minimal valence basis and approximates all
integrals from experimental data and a few empirical parameters, making
it computationally extremely cheap. At the same time, it retains the
conceptual clarity of Hückel theory and allows for straightforward
interpretation of the inverted gap in terms of the well-known concept of
dynamic spin polarization (DSP). We show that the method is capable of
finding hits both in the local chemical space around known scaffolds,
and of identifying hits in more diverse datasets. While there are clear
limitations to the method, we believe that it will open the doors for
ultra high-throughput virtual screening (UHTVS) of compound libraries
orders of magnitude larger than today. The method is also perfectly
applicable as a scoring function in inverse design algorithms.

\hypertarget{theory}{%
\section{Theory}\label{theory}}

While the recent literature has focused on applying high-level \emph{ab
initio} and DFT methods to study INVEST compounds, we believe that some
conceptual clarity has been lost in process. Instead, we apply the
simplest possible electronic structure method that can capture the
physics of the problem. The Pariser-Parr-Pople (PPP) method is an
extension of the semi-empirical Hückel molecular orbital theory for
\(\pi\)-electron systems that includes electron
correlation.\textsuperscript{\protect\hyperlink{ref-pariserSemiEmpiricalTheory1953a}{15}--\protect\hyperlink{ref-popleElectronInteractionUnsaturated1953}{17}}
It originated in the 1950s, with new parametrizations being developed
mainly in the 1960s and 1970s, and was used in the dye industry at least
until the 1980s before the advent of more accurate ab initio
methods.\textsuperscript{\protect\hyperlink{ref-griffithsPracticalAspectsColour1982}{18}}
The PPP method retains the conceptual clarity of Hückel theory, while at
the same time including the electron correlation that is necessary to
capture the inverted gap.

\hypertarget{pariser-parr-pople-theory}{%
\subsection{Pariser-Parr-Pople theory}\label{pariser-parr-pople-theory}}

Within Pople's formulation of PPP theory, Roothan's equations are solved
self-consistently.\textsuperscript{\protect\hyperlink{ref-popleElectronInteractionUnsaturated1953}{17}}.
Under the Zero Differential Overlap (ZDO) approximation, overlap
integrals are neglected for orbitals on different centers, leading to a
simplification of the Roothan equations as \(\mathbf{S}\) becomes the
unit matrix \(\mathbf{I}\)

\begin{equation}\protect\hypertarget{eq-roothan}{}{
\mathbf{FC} = \mathbf{SCE} = \mathbf{ICE} = \mathbf{CE}
}\label{eq-roothan}\end{equation}

where \(\mathbf{F}\) is the Fock matrix, \(\mathbf{C}\) contains the
molecular orbital coefficients and \(\mathbf{E}\) is the diagonal matrix
of the orbital energy eigenvalues. Given \(\mathbf{F}\), the
corresponding \(\mathbf{C}\) and \(\mathbf{E}\) can then be determined
either self-consistently (following
Pople\textsuperscript{\protect\hyperlink{ref-popleElectronInteractionUnsaturated1953}{17}})
or via configuration interaction (following Pariser and
Parr\textsuperscript{\protect\hyperlink{ref-pariserSemiEmpiricalTheory1953a}{15},\protect\hyperlink{ref-pariserSemiEmpiricalTheory1953}{16}}),
most often starting from a guess \(\mathbf{C}\) from the corresponding
Hückel model. The main effort is then to construct \(\mathbf{F}\), which
has the following matrix
elements\textsuperscript{\protect\hyperlink{ref-beveridgeParametrizationSemiempiricalPi1971}{19}}
\begin{equation}\protect\hypertarget{eq-f-rr}{}{
F_{rr} = H_{rr} + \sum_{t} P_{tt} \gamma_{rt} - \frac{1}{2} P_{rr} \gamma_{rr} 
}\label{eq-f-rr}\end{equation}

\begin{equation}\protect\hypertarget{eq-f-rs}{}{
F_{rs} = H_{rs} - \frac{1}{2} P_{rs} \gamma_{rs}
}\label{eq-f-rs}\end{equation}

where \(P_{rs}\) are elements of the density matrix \(\mathbf{P}\) and
\(\gamma_{rr}\) and \(\gamma_{rs}\) are parameters called the
\emph{one-center and two-center electron repulsion integrals},
respectively. The \emph{core resonance integral} matrix elements of
\(\mathbf{H}\) are given by
\begin{equation}\protect\hypertarget{eq-h-rr}{}{
H_{rr} = \alpha_r = U_r - \sum_{s \neq r} Z_s \gamma_{rs}
}\label{eq-h-rr}\end{equation}
\begin{equation}\protect\hypertarget{eq-h-rs}{}{
H_{rs} = \beta_{rs}
}\label{eq-h-rs}\end{equation}

where \(\alpha_{r}\) is the \emph{one-center core resonance integral},
\(U_r\) is a parameter of model called the \emph{atomic valence state
potential}, \(Z_s\) is the effective nuclear charge and \(\beta_{rs}\)
is the \emph{two-center resonance integral}, another parameter of the
model.

\hypertarget{parametrization}{%
\subsection{Parametrization}\label{parametrization}}

The four parameters are thus \(U_{r}\), \(\gamma_{rr}\), \(\gamma_{rs}\)
and \(\beta_{rs}\). They are derived either from experiment, first
principles, or a combination of both. The Pariser-Parr approximation
leads to \begin{equation}\protect\hypertarget{eq-u-r}{}{
U_{r} = - \mathrm{IP}_v
}\label{eq-u-r}\end{equation}

where \(\mathrm{IP}_v\) is the \emph{valence-state ionization potential}
of the orbital and atom in questions (\emph{e.g.} a 2\emph{p} orbital of
an \emph{sp}\textsuperscript{2}-hybridized C
atom).\textsuperscript{\protect\hyperlink{ref-mullikenNewElectroaffinityScale1934}{20}}
The \(\mathrm{IP}_v\) values (and the corresponding electron affinities
\(\mathrm{EA}_v\)) are tabulated by Hinze and Jaffé as determined from
experimental
data.\textsuperscript{\protect\hyperlink{ref-hinzeElectronegativity1962}{21},\protect\hyperlink{ref-hinzeElectronegativityOrbitalElectronegativity1962}{22}}
\(\gamma_{rr}\) also enjoys an almost universally adopted approximation
due to Pariser and Parr:
\begin{equation}\protect\hypertarget{eq-gamma-rr}{}{
\gamma_{rr} = \mathrm{IP}_v - \mathrm{EA}_v
}\label{eq-gamma-rr}\end{equation}

For \(\gamma_{rs}\) and \(\beta_{rs}\), there is much less consensus.
Formulas for \(\gamma_{rs}\) have been suggested by, among others,
Mataga and
Nishimoto\textsuperscript{\protect\hyperlink{ref-matagaElectronicStructureSpectra1957}{23}}
and
Ohno\textsuperscript{\protect\hyperlink{ref-ohnoRemarksPariserParrPopleMethod1964}{24}},
but here we follow the approach by Beveridge and
Hinze\textsuperscript{\protect\hyperlink{ref-beveridgeParametrizationSemiempiricalPi1971}{19}}:
\begin{equation}\protect\hypertarget{eq-gamma-rs}{}{
\gamma_{rs} = \frac{1}{a_{rs} \exp{\left( -r_{rs}^2/2a_{rs}^2\right)} + r_{rs}}
}\label{eq-gamma-rs}\end{equation}

where \(r_{rs}\) is the distance between the two atoms \(r\) and \(s\),
and \begin{equation}\protect\hypertarget{eq-a}{}{
a_{rs} = \frac{2}{\gamma_{rr} + \gamma_{ss}}
}\label{eq-a}\end{equation}

For \(\beta_{rs}\), expressions have been given by, \emph{e.g.},
Linderberg\textsuperscript{\protect\hyperlink{ref-linderbergConsistencyRequirementPariserParrPople1967}{25}}
and
Jug\textsuperscript{\protect\hyperlink{ref-jugOperatorEquationsApproximate1972}{26}},
based on first principles and by
Dewar\textsuperscript{\protect\hyperlink{ref-chungGroundStatesConjugated1965}{27}}
based on experimental data. Here we again follow Beveridge and
Hinze,\textsuperscript{\protect\hyperlink{ref-beveridgeParametrizationSemiempiricalPi1971}{19}}
who derived, following
Ohno\textsuperscript{\protect\hyperlink{ref-ohnoRemarksPariserParrPopleMethod1964}{24}}:

\begin{equation}\protect\hypertarget{eq-beta-rs}{}{
\beta_{rs} = \frac{1}{2} \left( Z_r + Z_s \right) S_{rs} \left( \gamma_{rs} - \frac{2C}{r_{rs}} \right)
}\label{eq-beta-rs}\end{equation}

where \(C = 0.545\) is a parameter that was fit in the original
publication to reproduce experimental excitation
energies.\textsuperscript{\protect\hyperlink{ref-beveridgeParametrizationSemiempiricalPi1971}{19}}
The overlap integral \(S_{rs}\) is determined from the overlap of Slater
\emph{p} orbitals with exponents

\[
\zeta_r = \frac{1280}{501} \gamma_{rr}
\]

according to recursion formulas given by
Mulliken.\textsuperscript{\protect\hyperlink{ref-mullikenFormulasNumericalTables1949}{28}}
Thus, the four parameters \(U_{r}\), \(\gamma_{rr}\), \(\gamma_{rs}\)
and \(\beta_{rs}\) are expressed completely in terms of the
corresponding valence state ionization potentials \(\mathrm{IP}_v\) and
electron affinities \(\mathrm{EA}_v\) as well as the completely
empirical parameter \(C\).

The distance dependence of \(\beta_{rs}\) and \(\gamma_{rs}\) allows for
treatment of bond-length alternation beyond idealized geometries. In
addition, we have followed the
literature\textsuperscript{\protect\hyperlink{ref-murrellSemiempiricalSelfconsistentfieldMolecular1972}{29}}
and added an angle-dependence to \(\beta_{rs}\) according to

\[
\beta_{rs}^\prime = \beta_{rs} \cos{\theta}
\]

where \(\theta\) is the twist angle between the two \emph{p} orbitals.
We have determined \(\theta\) as the average of all possible dihedrals
involving the two atoms \(r\) and \(s\).

\hypertarget{excited-state-calculations}{%
\subsection{Excited state
calculations}\label{excited-state-calculations}}

The energy difference of the vertically excited S\textsubscript{1} and
T\textsubscript{1} states can be determined by Roothan's expression at
the SCF
level\textsuperscript{\protect\hyperlink{ref-roothaanNewDevelopmentsMolecular1951}{30}}
\begin{equation}\protect\hypertarget{eq-exchange}{}{
\DeltaEST{SCF} = \E{S} - \E{T} = 2 \left( xy \vert yx \right) = 2 \left( x \lvert K_y \rvert x \right) = 2K 
}\label{eq-exchange}\end{equation}

where \(K\) is the exchange integral between the two orbitals involved
in the single excitation (normally the HOMO and the LUMO). As defined
here, a positive \(\DeltaE{ST}\) corresponds to the usual situation with
the triplet being lower in energy than the singlet, while a positive
\(\DeltaE{ST}\) corresponds to an inverted gap. The exchange integral
depends strongly on the spatial overlap between the the two orbitals,
which in the ZDO approximation can be computed
as:\textsuperscript{\protect\hyperlink{ref-peachExcitationEnergiesDensity2008}{31}}

\[
O_{xy} = \sum_{r} |c_{r,x}||c_{r,y}|
\]

where \(|c_{r,i}|\) is the absolute value of the coefficient of orbital
\(i\) on center \(r\). When the overlap is zero, the exchange
interaction vanishes and \(\DeltaEST{SCF}\) is zero. Unfortunately, so
is the oscillator strength, \(f\), between the two
states,\textsuperscript{\protect\hyperlink{ref-klessingerExcitedStatesPhotochemistry1995}{32}}
which is non-optimal for OLED materials that should emit light.
Consequently, there is a trade-off between having a small HOMO-LUMO
overlap to reduce the exchange interaction, while still maintaining a
sufficient oscillator strength.

A more precise expression for \(\DeltaE{ST}\) can be obtained by
configuration interaction singles
(CIS).\textsuperscript{\protect\hyperlink{ref-beveridgeParametrizationSemiempiricalPi1971}{19}}
While CIS adds some correlation for the excited states, it is necessary
to add additional excitations beyonds singles to capture the inverted
\(\DeltaE{ST}\). Fortunately, the most important excitations have
already been identified in the literature long ago. Singlet-triplet
inversion occurs also for the ground state, where several violations of
Hund's rule are well known, for example for bond-equalized
cyclobutadiene at the \emph{D}\textsubscript{4h} geometry and pentalene
at the \emph{D}\textsubscript{2h} geometry. Borden and Davidson
explained this effect in terms of dynamic spin polarization
(DSP),\textsuperscript{\protect\hyperlink{ref-bordenTheoreticalStudiesDiradicals1981}{33}}
in which the electrons of a pair of disjoint non-bonding molecular
orbitals (NBMOs, Figure~\ref{fig-dsp}a) experience stronger electron
correlation effects in the singlet state as compared to the triplet
state (Figure~\ref{fig-dsp}b). For a pedagogic introduction to the
topic, the reader is referred to an excellent article by
Karafiloglou\textsuperscript{\protect\hyperlink{ref-karafiloglouDoubleDynamicSpin1989}{34}}
As outlined by Kollmar and
Staemmler,\textsuperscript{\protect\hyperlink{ref-kollmarViolationHundRule1978}{3}}
as well as Malrieu and
co-workers,\textsuperscript{\protect\hyperlink{ref-benamorSpinPolarizationElectronic2020}{35}}
an alternative view of the same phenomenon is through admixture of
singly-excited configurations into the electronic wavefunction of the
ground state (Figure~\ref{fig-dsp}c). The DSP effect for the
S\textsubscript{1} excited state can correspondingly be described by
admixture of doubly excited
configurations,\textsuperscript{\protect\hyperlink{ref-kollmarViolationHundRule1978}{3}}
as they are singly excited with respect to the S\textsubscript{1} state,
and doubly excited with respect to the S\textsubscript{0} ground state
(Figure~\ref{fig-dsp}d).

As first shown by Kollmar and
Staemmler,\textsuperscript{\protect\hyperlink{ref-kollmarViolationHundRule1978}{3}}
and recently re-emphasized by Pernal and
co-workers,\textsuperscript{\protect\hyperlink{ref-drwalRoleSpinPolarization2023}{36}}
the \(\DeltaE{ST}\) can be approximated by a combination of two terms,
the first one being the exchange interaction \(2K\) (given by
Equation~\ref{eq-exchange}) and a correction term \(\DeltaE[DSP]{}\) due
to the DSP, which can be approximated with perturbation
theory:\textsuperscript{\protect\hyperlink{ref-kollmarViolationHundRule1978}{3}}

\begin{figure}

{\centering \includegraphics{../data/figures/dynamic_spin_polarization.pdf}

}

\caption{\label{fig-dsp}Dynamic spin polarization stabilizes the
open-shell singlet state over the open-shell triplet. (a) Transformation
of the canonical frontier molecular orbitals creates a set of disjoint
non-bonding molecular orbitals for cyclobutadiene. Singlet and triplet
occupations are shown. (b) Dynamic spin polarization preferentially
stabilizes the singlet state of cyclobutadiene through additive spin
polarization for the singlet and competitive spin polarization for the
triplet. (c) An alternative description of the spin polarization
phenomenon is via configuration interaction and admixture of excited
configurations into the ground state. (d) In the same way as
cyclobutadiene is stabilized by DSP in the singlet ground state,
molecules with inverted singlet-triplet gaps in the excited state are
also stabilized by DSP that can be described by admixture of excited
configurations. For singly excited states, this means addition of doubly
excited states.}

\end{figure}

\[
\DeltaE[DSP]{SCF} =  - \sum_{i}\sum_{k}
    \frac{1}{2} \left( 
    \frac{3 \Integral{-}}{\EState[1]{S} - \EState{S}}
    - \frac{\Integral{-}}{\EState[1]{T} - \EState{T}}
    - \frac{2 \Integral{+}}{\EState[2]{T} - \EState{T}}
    \right)
\]

Here, \(K_x\) and \(K_y\) are exchange operators,
\(E(\phi_{\mathrm{S}})\) and \(E(\phi_{\mathrm{T}})\) are the energies
of the singlet and triplet excited determinants, while
\(E (\phi_{\mathrm{S}}^{1})\), \(E (\phi_{\mathrm{T}}^{1})\) and
\(E(\phi_{\mathrm{T}}^{2})\) are the energies of doubly excited
determinants. They can be written in terms of the correspond orbital
energies, \(\epsilon_i\), provided the same orbitals are used for both
the singlet and triplet
states.\textsuperscript{\protect\hyperlink{ref-benamorSpinPolarizationElectronic2020}{35}}

\[
\DeltaE[DSP]{SCF} =  - \sum_{i}\sum_{k}
    \frac{1}{2} \left( 
    \frac{3 \Integral{-}}{\epsilon_k - \epsilon_i}
    - \frac{\Integral{-}}{\epsilon_k - \epsilon_i}
    - \frac{2 \Integral{+}}{\epsilon_k - \epsilon_i}
    \right)
\]

The simplest approximation for \(\DeltaE{ST}\) taking DSP into account
is then

\[
\DeltaEST[DSP]{SCF} = 2K + \DeltaE[DSP]{SCF}
\]

Alternatively, \(\DeltaE[DSP]{}\) can be calculated with respect to the
CIS
states,\textsuperscript{\protect\hyperlink{ref-head-gordonDoublesCorrectionElectronic1994}{37}}

\[
\begin{split}
\DeltaE[DSP]{CIS} =  - \sum_{i}\sum_{k}
    \frac{1}{2} \left( 
    \frac{3 \Integral{-}}{\epsilon_k + \epsilon_y - \epsilon_y - \epsilon_i - \DeltaE{S,CIS}}
    - \frac{\Integral{-}}{\epsilon_k + \epsilon_y - \epsilon_y - \epsilon_i - \DeltaE{T,CIS}} \right. \\
    \left. - \frac{2 \Integral{+}}{\epsilon_k + \epsilon_y - \epsilon_y - \epsilon_i - \DeltaE{T,CIS}}
    \right)
\end{split}
\]

and added as a correction to the energy gap at the CIS level to yield a
possibly more accurate value

\[
\DeltaEST[DSP]{CIS} = \DeltaEST{CIS} + \DeltaE[DSP]{CIS}
\]

\hypertarget{methods}{%
\section{Methods}\label{methods}}

The PPP method was implemented in the Python package
\textsc{coulson}\textsuperscript{\protect\hyperlink{ref-jornerCoulson}{38}}
which is freely available on GitHub with a permissive MIT license.
Further details on \textsc{coulson} will be reported elsewhere. The PPP
wavefunction was converged with the self-consistent field method using
the naive variational principle. To speed up the calculations, we
derived a spline interpolation of the overlap
integrals,\textsuperscript{\protect\hyperlink{ref-mullikenFormulasNumericalTables1949}{28}}
which will be reported separately. All calculations have been performed
on a MacBook Air laptop computer with an M2 processor. We further
integrated \textsc{coulson} with
\textsc{PySCF}\textsuperscript{\protect\hyperlink{ref-sunRecentDevelopmentsPySCF2020}{39}}
to provide alternative algorithms for the SCF convergence and CIS and
found the results consistent within numerical accuracy. Oscillator
strengths were calculated with the dipole length
approximation.\textsuperscript{\protect\hyperlink{ref-klessingerExcitedStatesPhotochemistry1995}{32},\protect\hyperlink{ref-jorgensenTimedependentHartreeFockCalculations1970}{40},\protect\hyperlink{ref-damourGroundExcitedStateDipole2023}{41}}
Some compounds exhibit negative triplet excitation energies and in some
case even negative singlet excitation energies, which indicates a
restricted-unrestricted instability of the ground state wavefunction. We
have here taken the pragmatic approach to ignore these issues, while we
will highlight below some examples where it occurs and if it has any
effect on the overall results. This approach can be partly justified as
the reference data also does not include any stability analysis. The
results should anyway be indicative of the gap between the lowest
singlet and triplet states with open-shell character, regardless of if
they are lower in energy than the closed-shell singlet.

We used four separate datasets for this study. The first is a set of
azaphenalenes previously studied by some of us, comprising 257
substituted compounds, for which excitation energies were computed with
(2nd-order) algebraic-diagrammatic construction, ADC(2)/{[}Robert: Fill
out{]}, and geometries were optimized with B3LYP/{[}Robert: Fill
out{]}\textsuperscript{\protect\hyperlink{ref-polliceOrganicMoleculesInverted2021}{7}}
The second is a set substituted 138 aza-azulenes with ADC(2)/{[}Robert:
Fill out{]} excitation energies and optimized with the B97-3c composite
method.{[}Robert: Add ref to dataset if it exists.{]} The third is a set
of 16 rationally designed scaffolds, which including substitutions
amounts to 68695 unique compounds, optimized at the B97-3c level and
with excitation energies at the ADC(2)/{[}Robert: Fill out{]}
level.{[}Robert: Add ref to dataset if it exists.{]} The fourth is a set
315 non-alternant hydrocarbons including substitutions, divided into
three subsets of size 76, 187 and 52, optimized at the ωB97X-D/def2-TZVP
level and with excitation energies at the CC2/aug-cc-pVTZ level.

\hypertarget{results-and-discussion}{%
\section{Results and discussion}\label{results-and-discussion}}

\hypertarget{qualitative-understanding-of-the-inverted-gap}{%
\subsection{Qualitative understanding of the inverted
gap}\label{qualitative-understanding-of-the-inverted-gap}}

We first demonstrate that our method is capable of capturing the
inverted \(\DeltaE{ST}\) for some model compounds. As shown by Toyota
and co-workers, pentalene at the ideal \(D_\mathrm{2h}\) geometry has an
inverted gap, while relaxation to the minimum with \(C_\mathrm{2h}\)
symmetry leads to a normal
gap.\textsuperscript{\protect\hyperlink{ref-kosekiViolationHundMultiplicity1985}{5}}
In our calculations, pentalene with equal bond lengths of 1.4 Å displays
a small HOMO-LUMO overlap of 0.24, leading to a small exchange
interaction of only \(2K=\) 2.99 kcal/mol
(Figure~\ref{fig-pentalene-ideal-orbitals}). The dynamic spin
polarization correction \(\DeltaE[DSP]{SCF}\) of -8.96 kcal/mol leads to
a net predicted \(\DeltaEST[DSP]{SCF}\) of -5.97 kcal/mol. Adding
additional correlation with CIS leads to a \(\DeltaEST[DSP]{CIS}\) of
-4.07 kcal/mol. The reference value calculated with ADC(2) is -3.70
kcal/mol. The simple perturbation theory model allows us to gain
additional insight into the contributions to \(\DeltaE[DSP]{SCF}\) as
they correspond to single excitations between doubly occupied orbitals
below the HOMO and unoccupied orbitals above the LUMO
(Figure~\ref{fig-pentalene-ideal-excitations}). Three excitations,
HOMO-3 → LUMO+2, HOMO-2 → LUMO+1 and HOMO-1 → LUMO+3 are identified as
the main contributors to the spin polarization effect, and could
potentially be tuned by substituent effects.

\begin{figure}

{\centering \includegraphics{../results/figures/pentalene_ideal_orbitals.png}

}

\caption{\label{fig-pentalene-ideal-orbitals}Orbitals energies and
frontier orbitals of pentalene at the idealized \(D_\mathrm{2h}\)
geometry and bond lengths of 1.4 Å.}

\end{figure}

\begin{figure}

{\centering \includegraphics{../results/figures/pentalene_ideal_excitations.png}

}

\caption{\label{fig-pentalene-ideal-excitations}Excitations and their
contribution to \(\DeltaE[DSP]{SCF}=\) -8.96 kcal/mol for pentalene at
the \(D_\mathrm{2h}\) geometry and bond lengths of 1.4 Å.}

\end{figure}

Using the relaxed \(C_\mathrm{2h}\) geometry leads to dramatic changes
in the frontier orbitals, with an increased HOMO-LUMO overlap of 0.86
and a sizeable exchange interaction of 15.88 kcal/mol
(Figure~\ref{fig-pentalene-orbitals}). Consequently, the calculated gap
is now normal at \(\DeltaEST[DSP]{SCF}=\) 13.16 kcal/mol, aggravated by
a diminished \(\DeltaE[DSP]{SCF}\) of only -2.72 kcal/mol. A similar
result is obtained with CIS: \(\DeltaEST[DSP]{CIS}=\) 21.08 kcal/mol.
The reference value calculated with ADC(2) is 19.92 kcal/mol. The
excitations contributing to \(\DeltaE[DSP]{SCF}\) at the
\(D_\mathrm{2h}\) geometry have been significantly diminished, only
partially alleviated by the addition of a minor stabilization from
HOMO-2 → LUMO+3 (Figure~\ref{fig-pentalene-excitations}).

\begin{figure}

{\centering \includegraphics{../results/figures/pentalene_orbitals.png}

}

\caption{\label{fig-pentalene-orbitals}Orbitals energies and frontier
orbitals of pentalene at the optimized \(C_\mathrm{2h}\) geometry.}

\end{figure}

\begin{figure}

{\centering \includegraphics{../results/figures/pentalene_excitations.png}

}

\caption{\label{fig-pentalene-excitations}Excitations and their
contribution to \(\DeltaE[DSP]{SCF}=\) -2.72 kcal/mol for pentalene at
the optimized \(C_\mathrm{2h}\) geometry.}

\end{figure}

Azaphenalene is the prototypical molecule that started the renewed
investigations into INVEST molecules, and has been the subject of
numerous studies with high-level quantum chemical methods. Can we
capture the inverted gap, as first measured by Leupin and
Wirz?\textsuperscript{\protect\hyperlink{ref-leupinLowlyingElectronicallyExcited1980}{4}}
The calculations reveal that the HOMO and LUMO are well separated with
spatial overlap of 0.14 and a correspondingly low exchange interaction
of 0.64 kcal/mol (Figure~\ref{fig-azaphenalene-orbitals}). The larger
azaphenalene has 30 excitations that could possibly contribute to the
large \(\DeltaE[DSP]{SCF}\) of -11.34 kcal/mol, but the most import by
far are those from the doubly degenerate HOMO-1 to the doubly degenerate
LUMO+1 (Figure~\ref{fig-azaphenalene-excitations}). In total,
\(\DeltaEST[DSP]{SCF}=\) -10.71 kcal/mol, which is slightly smaller at
the CIS level with \(\DeltaEST[DSP]{CIS}=\) -7.41 kcal/mol.

To conclude this section on the model compounds, our PPP-based protocol
is able to capture the inverted gap, and also provides qualitative
understanding into the physical mechanism of dynamic spin polarization
through identification and visualization of the corresponding
excitations. For these two compounds, the calculated
\(\DeltaEST[DSP]{}\) are more negative at the SCF level compared to CIS.
Compared to ADC(2), it would seem that CIS is preferable, but to
elucidate to which method is better, we now turn to larger datasets.

\begin{figure}

{\centering \includegraphics{../results/figures/azaphenalene_orbitals.png}

}

\caption{\label{fig-azaphenalene-orbitals}Orbitals energies and frontier
orbitals of azaphenalene.}

\end{figure}

\begin{figure}

{\centering \includegraphics{../results/figures/azaphenalene_excitations.png}

}

\caption{\label{fig-azaphenalene-excitations}Excitations and their
contribution to \(\DeltaE[DSP]{SCF}=\) -11.34 kcal/mol for azaphenalene}

\end{figure}

\hypertarget{local-chemical-space-of-azaphenalenes-and-azaazulenes}{%
\subsection{Local chemical space of azaphenalenes and
azaazulenes}\label{local-chemical-space-of-azaphenalenes-and-azaazulenes}}

While our method captures the inverted gap of azaphenalene and
pentalene, it also needs to capture trends with substitution to work
effectively in virtual screening. We therefore calculated 256
substituted azaphenalenes that have previously been studied by some of
us.\textsuperscript{\protect\hyperlink{ref-polliceOrganicMoleculesInverted2021}{7}}
The PPP level S\textsubscript{1} and T\textsubscript{1} excitation
energies are well correlated with those from ADC(2) with \(R^2\) of 0.90
and 0.94, respectively (Figure~\ref{fig-local-azaphenalenes}a,b).
Unfortunately, the oscillator strengths are not as well reproduced with
an \(R^2\) of 0.54, although the Spearman \(\rho\) of 0.82 indicates the
values might be used to rank potential candidates
(Figure~\ref{fig-local-azaphenalenes}c). The problems of obtaining
accurate oscillator strengths at the PPP level of theory are well known
in the literature, and the problem is especially exacerbated with low
oscillator strengths as
here.\textsuperscript{\protect\hyperlink{ref-misurkinElectronicStructureConjugated1974}{42}}.
Crucially, the \(\DeltaE{ST}\) shows good correlations, with \(R^2=\)
0.81 using \(\DeltaEST[DSP]{CIS}\)
(Figure~\ref{fig-local-azaphenalenes}f). This \(R^2\) value is
essentially unchanged from the gap at the CIS level without DSP (0.81,
Figure~\ref{fig-local-azaphenalenes}d) and markedly better than the gap
at the SCF level (0.52, Figure~\ref{fig-local-azaphenalenes}e), showing
the importance of going beyond the SCF level to include at least some
configuration interaction. The results are of similar or better quality
achieved by Pernal and co-workers for different set of azaphenalenes
using SCF + DSP with orbitals from
DFT.\textsuperscript{\protect\hyperlink{ref-drwalRoleSpinPolarization2023}{36}}
To further analyze the results in terms of a binary classification into
normal/inverted, we calculated the true positives (TP), true negatives
(TN), false positives (FP) and false negatives (FN), as well as a range
of common classification scores (Table~\ref{tbl-local-azaphenalenes}).
The most important metrics for virtual screening are, in our opinion,
the recall and the specificity. The recall measures the fraction of
inverted molecules that the protocol captures, and is 0.58 at the SCF
level and 0.56 at the CIS level. The specificity measures the fraction
of non-inverted molecules that the protocol identifies, and is 0.94 at
the SCF level and 0.97 at the CIS level. This metrics mean that we are
able to capture a large proportion of the molecules with inverted gaps,
while filtering out most non-inverted molecules. The results for these
azaphenalenes clearly indicate that the PPP protocol could be used to
prescreen candidates of this compound class for \(\DeltaE{ST}\), while
further pruning with another method is likely needed for the oscillator
strengths.

\begin{figure}

{\centering \includegraphics{../results/figures/local_azaphenalenes.pdf}

}

\caption{\label{fig-local-azaphenalenes}Excitation properties for
azaphenalenes \emph{vs.} the reference level. (a) Singlet excitation
energies. (b) Triplet excitation energies. (c) Oscillator strengths. (d)
Singlet-triplet energy gaps with CIS. (e) Singlet-triplet energy gaps
with SCF + DSP. (f) Singlet-triplet energy gaps with CIS + DSP.}

\end{figure}

\hypertarget{tbl-local-azaphenalenes}{}
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 22\tabcolsep) * \real{0.0490}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 22\tabcolsep) * \real{0.0882}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 22\tabcolsep) * \real{0.0980}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 22\tabcolsep) * \real{0.0588}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 22\tabcolsep) * \real{0.1078}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 22\tabcolsep) * \real{0.1176}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 22\tabcolsep) * \real{0.0980}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 22\tabcolsep) * \real{0.1471}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 22\tabcolsep) * \real{0.0588}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 22\tabcolsep) * \real{0.0588}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 22\tabcolsep) * \real{0.0588}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 22\tabcolsep) * \real{0.0588}}@{}}
\caption{\label{tbl-local-azaphenalenes}Metrics for
azaphenalenes.}\tabularnewline
\toprule()
\begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
\(R^2\)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
\(\rho\)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
F1
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
ROC-AUC
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
Accuracy
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
Recall
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
Specificity
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
TP
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
TN
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
FP
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
FN
\end{minipage} \\
\midrule()
\endfirsthead
\toprule()
\begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
\(R^2\)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
\(\rho\)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
F1
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
ROC-AUC
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
Accuracy
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
Recall
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
Specificity
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
TP
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
TN
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
FP
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
FN
\end{minipage} \\
\midrule()
\endhead
SCF & 0.52 & 0.72 & 0.73 & 0.76 & 0.62 & 0.58 & 0.94 & 128 & 32 & 2 &
94 \\
CIS & 0.81 & 0.83 & 0.72 & 0.77 & 0.62 & 0.56 & 0.97 & 125 & 33 & 1 &
97 \\
\bottomrule()
\end{longtable}

While we see success for the azaphenalenes, the situation for the
azulenes is unfortunately worse. The PPP level S\textsubscript{1} and
T\textsubscript{1} excitation energies are only moderately correlated
with those from ADC(2) with \(R^2\) of 0.44 and 0.47, respectively
(Figure~\ref{fig-local-azulenes}a,b). For the oscillator strengths, the
situation is rather catastrophic with \(R^2\) of 0.00 implying no
correlation whatsoever (Figure~\ref{fig-local-azulenes}c). The situation
for the \(\DeltaE{ST}\) also worse than for the azaphenalenes, with
\(R^2=\) 0.46 with CIS + DSP, a marked improvement on 0.18 with SCF +
DSP (Figure~\ref{fig-local-azulenes}e,f). Unfortunately, the negative
gaps are not recovered, leading to recalls of 0.00 as none of five
inverted molecules could be identified (Table~\ref{tbl-local-azulenes}).
It could be speculated that the worse performance for the azulenes comes
from the fact that they are non-alternants, for which the approximations
are expected to be less applicable. Considering the contrasting
performance for the azaphenalenes and azulenes, it seems clear that the
PPP screening protocol will inevitably miss some hits belonging to
particular compound classes. To investigate the global performance of
the model we turned to a more diverse set with 16 different compound
classes.

\begin{figure}

{\centering \includegraphics{../results/figures/local_azulenes.pdf}

}

\caption{\label{fig-local-azulenes}Excitation properties for azaazulenes
\emph{vs.} the reference level. (a) Singlet excitation energies. (b)
Triplet excitation energies. (c) Oscillator strengths. (d)
Singlet-triplet energy gaps with CIS. (e) Singlet-triplet energy gaps
with SCF + DSP. (f) Singlet-triplet energy gaps with CIS + DSP.}

\end{figure}

\hypertarget{tbl-local-azulenes}{}
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 22\tabcolsep) * \real{0.0490}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 22\tabcolsep) * \real{0.0882}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 22\tabcolsep) * \real{0.0980}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 22\tabcolsep) * \real{0.0588}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 22\tabcolsep) * \real{0.1078}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 22\tabcolsep) * \real{0.1176}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 22\tabcolsep) * \real{0.0980}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 22\tabcolsep) * \real{0.1471}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 22\tabcolsep) * \real{0.0588}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 22\tabcolsep) * \real{0.0588}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 22\tabcolsep) * \real{0.0588}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 22\tabcolsep) * \real{0.0588}}@{}}
\caption{\label{tbl-local-azulenes}Metrics for
azaazulenes.}\tabularnewline
\toprule()
\begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
\(R^2\)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
\(\rho\)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
F1
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
ROC-AUC
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
Accuracy
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
Recall
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
Specificity
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
TP
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
TN
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
FP
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
FN
\end{minipage} \\
\midrule()
\endfirsthead
\toprule()
\begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
\(R^2\)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
\(\rho\)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
F1
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
ROC-AUC
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
Accuracy
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
Recall
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
Specificity
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
TP
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
TN
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
FP
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
FN
\end{minipage} \\
\midrule()
\endhead
SCF & 0.18 & 0.41 & nan & 0.50 & 0.96 & 0.00 & 1.00 & 0 & 133 & 0 & 5 \\
CIS & 0.46 & 0.65 & nan & 0.50 & 0.96 & 0.00 & 1.00 & 0 & 133 & 0 & 5 \\
\bottomrule()
\end{longtable}

\hypertarget{screening-widely}{%
\subsection{Screening widely}\label{screening-widely}}

Here we used a dataset previously generated by some of us using rational
design rules, comprising 16 different molecular scaffolds, the details
of which will be presented elsewhere. Out of 68695 compounds, only 5
failed, which corresponds to a success rate of 99.99. The total runtime
using four cores on the M2 processor was 12.61 CPU hours. The
S\textsubscript{1} and T\textsubscript{1} excitation energies are fairly
reproduced with (Figure~\ref{fig-global-all}a,b) \(R^2\) of 0.66 and
0.70 vs ADC(2), respectively. As seen in
Figure~\ref{fig-global-all-gaps}a, the \(\DeltaE{ST}\) at the SCF + DSP
level shows a moderate \(R^2\) of 0.46, which increases significantly to
0.71 at the CIS + DSP level (Figure~\ref{fig-global-all-gaps}b).
However, due to a systematic overestimation of \(\DeltaE{ST}\), the F1
score is still low at 0.27 and a poor recall of 0.16
(Table~\ref{tbl-global-all}). We therefore added a linear correction
\(\DeltaEST[DSP]{CIS,LC}=\) 0.53 \(\times\DeltaEST[DSP]{CIS}\) -0.15,
which increases the F1 score to 0.55 and the recall to 0.51
(Figure~\ref{fig-global-all-gaps}c). The specificity simultaneously
decreases from 0.99 to 0.96, but overall the linear correction would be
preferable for virtual screening where there is a stronger focus on
finding rare hits. Applying the linear correction to the azaphenalenes
(Figure~\ref{fig-local-corr}a) and (Figure~\ref{fig-local-corr}b) leads
to an improvement in the recall of the former from 0.56 to 0.94 while
still none of the inverted azulenes are recovered. The oscillator
strengths are unfortunately poorly correlated for the rational design
set, with \(R^2=\) 0.21, although the Spearman rank correlation is more
encouraging at 0.58 (Figure~\ref{fig-global-all}c). Taken together, the
results reinforce the conclusions from the case study of the
azaphenalenes, that the PPP protocol is suitable for screening
\(\DeltaE{ST}\) while it struggles for oscillator strengths.

A more detailed breakdown of the correlations into the different
scaffolds shows that the highest \(R^2\) of 0.82 is obtained for
dicyclopenta{[}a,e{]}cyclooctene, while the lowest is obtained for
bowtiene with 0.07. For recall, the corresponding range is 0.92 for
zurlene to 0.01 for phenazulene. The wide range in \(R^2\) and recall
further reinforces that the method struggles with some particular
compounds classes despite its favorable global performance. Despite
these problems, the method is able to capture inverted molecules in 15
out of the 15 scaffolds where they occur. Even for the azulenes included
in this screening set at least some candidates are found, in contrast to
the findings above for the azaazulenes. The per-scaffold oscillator
strengths are arguably sufficiently good to allow for local screening in
some cases (Figure~\ref{fig-global-scaffolds-oscillator}). The
per-scaffold S\textsubscript{1} and T\textsubscript{1} excitation
energies can be found in (Figure~\ref{fig-global-scaffolds-singlet}) and
(Figure~\ref{fig-global-scaffolds-triplet}), respectively. Notable
failures are for cyclobuta-1,3-diene with \(R^2\) of 0.29 and 0.27,
respectively, and phenazulene and dicyclopenta{[}a,c{]}cyclooctene,
where some negative T\textsubscript{1} excitation energies indicate
restricted-unrestricted instabilities.

\begin{figure}

{\centering \includegraphics{../results/figures/global_all_gaps.pdf}

}

\caption{\label{fig-global-all-gaps}Singlet-triplet energy gaps for
rationally design set \emph{vs.} reference level for (a) SCF + DSP, (b)
CIS + DSP and (c) CIS + DSP and linear correction.}

\end{figure}

\hypertarget{tbl-global-all}{}
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 22\tabcolsep) * \real{0.1404}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 22\tabcolsep) * \real{0.0789}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 22\tabcolsep) * \real{0.0877}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 22\tabcolsep) * \real{0.0526}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 22\tabcolsep) * \real{0.0965}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 22\tabcolsep) * \real{0.1053}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 22\tabcolsep) * \real{0.0877}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 22\tabcolsep) * \real{0.1316}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 22\tabcolsep) * \real{0.0526}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 22\tabcolsep) * \real{0.0614}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 22\tabcolsep) * \real{0.0526}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 22\tabcolsep) * \real{0.0526}}@{}}
\caption{\label{tbl-global-all}Metrics for rationally designed
set.}\tabularnewline
\toprule()
\begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
\(R^2\)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
\(\rho\)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
F1
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
ROC-AUC
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
Accuracy
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
Recall
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
Specificity
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
TP
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
TN
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
FP
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
FN
\end{minipage} \\
\midrule()
\endfirsthead
\toprule()
\begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
\(R^2\)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
\(\rho\)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
F1
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
ROC-AUC
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
Accuracy
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
Recall
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
Specificity
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
TP
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
TN
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
FP
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
FN
\end{minipage} \\
\midrule()
\endhead
SCF + DSP & 0.46 & 0.60 & 0.40 & 0.63 & 0.91 & 0.27 & 0.99 & 2073 &
60490 & 561 & 5566 \\
CIS + DSP & 0.71 & 0.85 & 0.27 & 0.58 & 0.90 & 0.16 & 0.99 & 1227 &
60736 & 315 & 6412 \\
CIS + DSP + LC & 0.71 & 0.85 & 0.55 & 0.73 & 0.91 & 0.51 & 0.96 & 3867 &
58592 & 2459 & 3772 \\
\bottomrule()
\end{longtable}

\hypertarget{external-validation}{%
\subsection{External validation}\label{external-validation}}

To further test the validity of the method, we calculated a series of
compounds recently published by Garner \emph{et.
al}.\textsuperscript{\protect\hyperlink{ref-garnerDoublebondDelocalizationNonalternant2023}{11}}
The dataset comes in three parts: (1) non-alternants, (2) non-alternants
with constrained high-symmetry geometries (we here follow the original
terminology and call these ``avoided symmetry'') and (3) substituted
non-alternants (here called ``substituted''). Out of the substituted
compounds, four could not be calculated as they contain four-coordinate
P atoms for which our PPP model lacks parameters. We here opted to use
the CC2/aug-cc-pVDZ data from the original paper as reference as it is
the highest level which has the most complete coverage of the dataset.

Our computed \(\DeltaE{ST}\) with the linearly corrected CIS + DSP show
fair correlations with the CC2 values, with \(R^2\)s of 0.68, 0.69 and
0.56, respectively (Figure~\ref{fig-garner-gaps}). Gratifyingly, the
recalls are 0.50, 0.80 and 0.74, respectively, showing that we can
recover a large part of the inverted molecules found with the much more
expensive CC2 method (Table~\ref{tbl-garner-gaps}). Although some of the
compounds are also present in the rationally designed set that we used
for the linear correction above, we believe that they are not
sufficiently many to compromise the use of the Garner data as external
validation (Figure~\ref{fig-garner-gaps}). We also indicate in the plots
which compounds show instabilities (Figure~\ref{fig-garner-gaps}). There
seems to be no clear deterioration in the performance, although also the
reference data might exhibit instabilities as there is no mention of any
stability analysis in the original
manuscript.\textsuperscript{\protect\hyperlink{ref-garnerDoublebondDelocalizationNonalternant2023}{11}}

In summary, the external validation shows that the PPP method can
recover inverted molecules at a fraction of the cost of more expensive
methods such as CC2.

\begin{figure}

{\centering \includegraphics{../results/figures/garner_gaps.pdf}

}

\caption{\label{fig-garner-gaps}Singlet-triplet energy gaps for CIS +
DSP and linear correction \emph{vs.} reference level for the subsets of
(a) non-alternants (b) non-alternants constrained to higher symmetry and
(c) substituted non-alternants. Compounds contained in the rationally
designed sets are marked with dashed circles and those with negative
triplet excitation energies are marked with a semi-transparent red
circle.}

\end{figure}

\hypertarget{tbl-garner-gaps}{}
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 22\tabcolsep) * \real{0.1565}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 22\tabcolsep) * \real{0.0783}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 22\tabcolsep) * \real{0.0870}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 22\tabcolsep) * \real{0.0522}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 22\tabcolsep) * \real{0.0957}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 22\tabcolsep) * \real{0.1043}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 22\tabcolsep) * \real{0.0870}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 22\tabcolsep) * \real{0.1304}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 22\tabcolsep) * \real{0.0522}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 22\tabcolsep) * \real{0.0522}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 22\tabcolsep) * \real{0.0522}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 22\tabcolsep) * \real{0.0522}}@{}}
\caption{\label{tbl-garner-gaps}Metrics for the dataset from Garner
\emph{et al.}}\tabularnewline
\toprule()
\begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
\(R^2\)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
\(\rho\)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
F1
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
ROC-AUC
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
Accuracy
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
Recall
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
Specificity
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
TP
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
TN
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
FP
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
FN
\end{minipage} \\
\midrule()
\endfirsthead
\toprule()
\begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
\(R^2\)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
\(\rho\)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
F1
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
ROC-AUC
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
Accuracy
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
Recall
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
Specificity
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
TP
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
TN
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
FP
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
FN
\end{minipage} \\
\midrule()
\endhead
Non-alternants & 0.68 & 0.78 & 0.67 & 0.75 & 0.97 & 0.50 & 1.00 & 2 & 75
& 0 & 2 \\
Avoided symmetry & 0.69 & 0.84 & 0.83 & 0.87 & 0.89 & 0.80 & 0.94 & 12 &
30 & 2 & 3 \\
Substituted & 0.56 & 0.78 & 0.80 & 0.85 & 0.90 & 0.74 & 0.96 & 14 & 50 &
2 & 5 \\
\bottomrule()
\end{longtable}

\hypertarget{conclusions-and-outlook}{%
\section{Conclusions and outlook}\label{conclusions-and-outlook}}

To conclude, we have shown that the Pariser-Parr-Pople theory, a simple
semi-empirical \(\pi\)-electron theory with a minimal basis, can be used
to screen for molecules with inverted singlet-triplets energy gaps, both
locally and globally. Unfortunately, the method does not seem capable of
screening for oscillator strengths with the same accuracy, with
exceptions for some scaffolds. The chief limitation of the method is
that it only includes the \(\pi\)-electron system and therefore
struggles with: (1) functional groups without a clear
\(\sigma\)--\(\pi\) separation (found in common functional groups such
as sulfonyls) (2) inductive effects and (3) neglect of
\(n \rightarrow \pi^*\) transitions. Further limitations include (4)
lack of solvation model and need for (5) already optimized geometries.
We believe that at at least some of these limitations could be mitigated
by a reparametrization specifically targeting inverted gaps, while
geometries from the fast GFN-xTB method would be suitable.

Taken together, we foresee that the presented methodology can be used
for ultrahigh-throughput virtual screening campaigns and in inverse
design algorithms, followed by curation of hits using more accurate
quantum-chemical methods. Active learning schemes could also be used
with machine learning corrections to the PPP singlet-triplet gaps. The
method has great potential to accelerate the discovery of the next
generation of OLED materials based on INVEST.

\hypertarget{acknowledgements}{%
\section{Acknowledgements}\label{acknowledgements}}

K.J. acknowledges funding through an International Postdoc grant from
the Swedish Research Council (No.~2020-00314). A. A.-G. acknowledges
support of Natural Resources Canada and the Canada 150 Research Chairs
program and CIFAR, as well as the generous support of Anders G. Frøseth.

\hypertarget{references}{%
\section{References}\label{references}}

\hypertarget{refs}{}
\begin{CSLReferences}{0}{0}
\leavevmode\vadjust pre{\hypertarget{ref-yangRecentAdvancesOrganic2017}{}}%
\CSLLeftMargin{(1) }%
\CSLRightInline{Yang, Z.; Mao, Z.; Xie, Z.; Zhang, Y.; Liu, S.; Zhao,
J.; Xu, J.; Chi, Z.; Aldred, M. P. Recent Advances in Organic Thermally
Activated Delayed Fluorescence Materials. \emph{Chem. Soc. Rev.}
\textbf{2017}, \emph{46} (3), 915--1016.
\url{https://doi.org/10.1039/C6CS00368K}.}

\leavevmode\vadjust pre{\hypertarget{ref-liOrganicMoleculesInverted2022}{}}%
\CSLLeftMargin{(2) }%
\CSLRightInline{Li, J.; Li, Z.; Liu, H.; Gong, H.; Zhang, J.; Yao, Y.;
Guo, Q. Organic Molecules with Inverted Singlet-Triplet Gaps.
\emph{Front. Chem.} \textbf{2022}, \emph{10}, 999856.
\url{https://doi.org/10.3389/fchem.2022.999856}.}

\leavevmode\vadjust pre{\hypertarget{ref-kollmarViolationHundRule1978}{}}%
\CSLLeftMargin{(3) }%
\CSLRightInline{Kollmar, H.; Staemmler, V. Violation of {Hund}'s Rule by
Spin Polarization in Molecules. \emph{Theoret. Chim. Acta}
\textbf{1978}, \emph{48} (3), 223--239.
\url{https://doi.org/10.1007/BF00549021}.}

\leavevmode\vadjust pre{\hypertarget{ref-leupinLowlyingElectronicallyExcited1980}{}}%
\CSLLeftMargin{(4) }%
\CSLRightInline{Leupin, W.; Wirz, J. Low-Lying Electronically Excited
States of Cycl{[}3.3.3{]}azine, a Bridged 12π-Perimeter. \emph{J. Am.
Chem. Soc.} \textbf{1980}, \emph{102} (19), 6068--6075.
\url{https://doi.org/10.1021/ja00539a016}.}

\leavevmode\vadjust pre{\hypertarget{ref-kosekiViolationHundMultiplicity1985}{}}%
\CSLLeftMargin{(5) }%
\CSLRightInline{Koseki, S.; Nakajima, T.; Toyota, A. Violation of
{Hund}'s Multiplicity Rule in the Electronically Excited States of
Conjugated Hydrocarbons. \emph{Can. J. Chem.} \textbf{1985}, \emph{63}
(7), 1572--1579. \url{https://doi.org/10.1139/v85-267}.}

\leavevmode\vadjust pre{\hypertarget{ref-desilvaInvertedSingletTriplet2019}{}}%
\CSLLeftMargin{(6) }%
\CSLRightInline{de Silva, P. Inverted {Singlet}--{Triplet Gaps} and
{Their Relevance} to {Thermally Activated Delayed Fluorescence}.
\emph{J. Phys. Chem. Lett.} \textbf{2019}, \emph{10} (18), 5674--5679.
\url{https://doi.org/10.1021/acs.jpclett.9b02333}.}

\leavevmode\vadjust pre{\hypertarget{ref-polliceOrganicMoleculesInverted2021}{}}%
\CSLLeftMargin{(7) }%
\CSLRightInline{Pollice, R.; Friederich, P.; Lavigne, C.; Gomes, G. dos
P.; Aspuru-Guzik, A. Organic Molecules with Inverted Gaps Between First
Excited Singlet and Triplet States and Appreciable Fluorescence Rates.
\emph{Matter} \textbf{2021}, \emph{4} (5), 1654--1682.
\url{https://doi.org/10.1016/j.matt.2021.02.017}.}

\leavevmode\vadjust pre{\hypertarget{ref-aizawaDelayedFluorescenceInverted2022}{}}%
\CSLLeftMargin{(8) }%
\CSLRightInline{Aizawa, N.; Pu, Y.-J.; Harabuchi, Y.; Nihonyanagi, A.;
Ibuka, R.; Inuzuka, H.; Dhara, B.; Koyama, Y.; Nakayama, K.; Maeda, S.;
Araoka, F.; Miyajima, D. Delayed Fluorescence from Inverted Singlet and
Triplet Excited States. \emph{Nature} \textbf{2022}, \emph{609} (7927),
502--506. \url{https://doi.org/10.1038/s41586-022-05132-y}.}

\leavevmode\vadjust pre{\hypertarget{ref-terenceblaskovitsSymmetryInducedSinglet2023}{}}%
\CSLLeftMargin{(9) }%
\CSLRightInline{Terence Blaskovits, J.; Garner, M. H.; Corminboeuf, C.
Symmetry‐{Induced Singlet}‐{Triplet Inversions} in {Non}‐{Alternant
Hydrocarbons}**. \emph{Angew. Chem. Int. Ed.} \textbf{2023}, \emph{62}
(15), e202218156. \url{https://doi.org/10.1002/anie.202218156}.}

\leavevmode\vadjust pre{\hypertarget{ref-omarIdentificationUnknownInverted2023}{}}%
\CSLLeftMargin{(10) }%
\CSLRightInline{Omar, O.; Xie, X.; Troisi, A.; Padula, D.
\emph{Identification of {Unknown Inverted Singlet-Triplet Cores} by
{High--Throughput Virtual Screening}}.
\url{https://doi.org/10.26434/chemrxiv-2023-h7tsp-v3}.}

\leavevmode\vadjust pre{\hypertarget{ref-garnerDoublebondDelocalizationNonalternant2023}{}}%
\CSLLeftMargin{(11) }%
\CSLRightInline{Garner, M. H.; Blaskovits, J. T.; Corminboeuf, C.
\emph{Double-bond delocalization in non-alternant hydrocarbons induces
inverted singlet-triplet gaps}.
\url{https://doi.org/10.26434/chemrxiv-2023-568ks}.}

\leavevmode\vadjust pre{\hypertarget{ref-casidaProgressTimeDependentDensityFunctional2012}{}}%
\CSLLeftMargin{(12) }%
\CSLRightInline{Casida, M. E.; Huix-Rotllant, M.
\href{http://www.annualreviews.org/doi/full/10.1146/annurev-physchem-032511-143803}{Progress
in {Time-Dependent Density-Functional Theory}}. \emph{Annu. Rev. Phys.
Chem.} \textbf{2012}, \emph{63}, 287--323.}

\leavevmode\vadjust pre{\hypertarget{ref-ricciSingletTripletExcited2021}{}}%
\CSLLeftMargin{(13) }%
\CSLRightInline{Ricci, G.; San‐Fabián, E.; Olivier, Y.; Sancho‐García,
J. C. Singlet‐{Triplet Excited}‐{State Inversion} in {Heptazine} and
{Related Molecules}: {Assessment} of {TD}‐{DFT} and {\emph{Ab Initio}}
{Methods}. \emph{ChemPhysChem} \textbf{2021}, cphc.202000926.
\url{https://doi.org/10.1002/cphc.202000926}.}

\leavevmode\vadjust pre{\hypertarget{ref-sancho-garciaViolationHundRule2022}{}}%
\CSLLeftMargin{(14) }%
\CSLRightInline{Sancho-García, J. C.; Brémond, E.; Ricci, G.;
Pérez-Jiménez, A. J.; Olivier, Y.; Adamo, C. Violation of {Hund}'s Rule
in Molecules: {Predicting} the Excited-State Energy Inversion by
{TD-DFT} with Double-Hybrid Methods. \emph{J. Chem. Phys.}
\textbf{2022}, \emph{156} (3), 034105.
\url{https://doi.org/10.1063/5.0076545}.}

\leavevmode\vadjust pre{\hypertarget{ref-pariserSemiEmpiricalTheory1953a}{}}%
\CSLLeftMargin{(15) }%
\CSLRightInline{Pariser, R.; Parr, R. G. A {Semi}-{Empirical Theory} of
the {Electronic Spectra} and {Electronic Structure} of {Complex
Unsaturated Molecules}. {I}. \emph{J. Chem. Phys.} \textbf{1953},
\emph{21} (3), 466--471. \url{https://doi.org/10.1063/1.1698929}.}

\leavevmode\vadjust pre{\hypertarget{ref-pariserSemiEmpiricalTheory1953}{}}%
\CSLLeftMargin{(16) }%
\CSLRightInline{Pariser, R.; Parr, R. G. A {Semi}-{Empirical Theory} of
the {Electronic Spectra} and {Electronic Structure} of {Complex
Unsaturated Molecules}. {II}. \emph{J. Chem. Phys.} \textbf{1953},
\emph{21} (5), 767--776. \url{https://doi.org/10.1063/1.1699030}.}

\leavevmode\vadjust pre{\hypertarget{ref-popleElectronInteractionUnsaturated1953}{}}%
\CSLLeftMargin{(17) }%
\CSLRightInline{Pople, J. A. Electron Interaction in Unsaturated
Hydrocarbons. \emph{Trans. Faraday Soc.} \textbf{1953}, \emph{49}, 1375.
\url{https://doi.org/10.1039/tf9534901375}.}

\leavevmode\vadjust pre{\hypertarget{ref-griffithsPracticalAspectsColour1982}{}}%
\CSLLeftMargin{(18) }%
\CSLRightInline{Griffiths, J. Practical Aspects of Colour Prediction of
Organic Dye Molecules. \emph{Dyes Pigments} \textbf{1982}, \emph{3}
(2-3), 211--233. \url{https://doi.org/10.1016/0143-7208(82)80024-7}.}

\leavevmode\vadjust pre{\hypertarget{ref-beveridgeParametrizationSemiempiricalPi1971}{}}%
\CSLLeftMargin{(19) }%
\CSLRightInline{Beveridge, D. L.; Hinze, J. Parametrization of
Semiempirical .pi.-Electron Molecular Orbital Calculations.
.pi.{Systems} Containing Carbon, Nitrogen, Oxygen, and Fluorine.
\emph{J. Am. Chem. Soc.} \textbf{1971}, \emph{93} (13), 3107--3114.
\url{https://doi.org/10.1021/ja00742a002}.}

\leavevmode\vadjust pre{\hypertarget{ref-mullikenNewElectroaffinityScale1934}{}}%
\CSLLeftMargin{(20) }%
\CSLRightInline{Mulliken, R. S. A {New Electroaffinity Scale};
{Together} with {Data} on {Valence States} and on {Valence Ionization
Potentials} and {Electron Affinities}. \emph{J. Chem. Phys.}
\textbf{1934}, \emph{2} (11), 782--793.
\url{https://doi.org/10.1063/1.1749394}.}

\leavevmode\vadjust pre{\hypertarget{ref-hinzeElectronegativity1962}{}}%
\CSLLeftMargin{(21) }%
\CSLRightInline{Hinze, J. A. Electronegativity. PhD thesis, {University
of Cincinnati}, {Ann Arbor}, 1962.
\url{http://myaccess.library.utoronto.ca/login?qurl=https\%3A\%2F\%2Fwww.proquest.com\%2Fdissertations-theses\%2Felectronegativity\%2Fdocview\%2F288031630\%2Fse-2\%3Faccountid\%3D14771}.}

\leavevmode\vadjust pre{\hypertarget{ref-hinzeElectronegativityOrbitalElectronegativity1962}{}}%
\CSLLeftMargin{(22) }%
\CSLRightInline{Hinze, Jürgen.; Jaffe, H. H. \textbf{Electronegativity.}
{\textbf{I}}\textbf{.} {\textbf{Orbital Electronegativity}} \textbf{of}
{\textbf{Neutral Atoms}}. \emph{J. Am. Chem. Soc.} \textbf{1962},
\emph{84} (4), 540--546. \url{https://doi.org/10.1021/ja00863a008}.}

\leavevmode\vadjust pre{\hypertarget{ref-matagaElectronicStructureSpectra1957}{}}%
\CSLLeftMargin{(23) }%
\CSLRightInline{Mataga, N.; Nishimoto, K. Electronic {Structure} and
{Spectra} of {Nitrogen Heterocycles}. \emph{Z. Für Phys. Chem.}
\textbf{1957}, \emph{13}, 140--157.
\url{https://doi.org/10.1524/zpch.1957.13.3_4.140}.}

\leavevmode\vadjust pre{\hypertarget{ref-ohnoRemarksPariserParrPopleMethod1964}{}}%
\CSLLeftMargin{(24) }%
\CSLRightInline{Ohno, K. Some Remarks on the {Pariser-Parr-Pople}
Method. \emph{Theor. Chim. Acta} \textbf{1964}, \emph{2} (3), 219--227.
\url{https://doi.org/10.1007/BF00528281}.}

\leavevmode\vadjust pre{\hypertarget{ref-linderbergConsistencyRequirementPariserParrPople1967}{}}%
\CSLLeftMargin{(25) }%
\CSLRightInline{Linderberg, J. Consistency Requirement in the
{Pariser-Parr-Pople} Model. \emph{Chem. Phys. Lett.} \textbf{1967},
\emph{1} (2), 39--41.
\url{https://doi.org/10.1016/0009-2614(67)80061-7}.}

\leavevmode\vadjust pre{\hypertarget{ref-jugOperatorEquationsApproximate1972}{}}%
\CSLLeftMargin{(26) }%
\CSLRightInline{Jug, K. Operator Equations in Approximate Molecular
Orbital Theories: {II}. {Integral} Approximations and Charge
Distributions. \emph{Theor. Chim. Acta} \textbf{1972}, \emph{26} (3),
231--236. \url{https://doi.org/10.1007/BF00529308}.}

\leavevmode\vadjust pre{\hypertarget{ref-chungGroundStatesConjugated1965}{}}%
\CSLLeftMargin{(27) }%
\CSLRightInline{Chung, A. L. H.; Dewar, M. J. S. Ground {States} of
{Conjugated Molecules}. {I}. {Semiempirical SCF MO Treatment} and {Its
Application} to {Aromatic Hydrocarbons}. \emph{J. Chem. Phys.}
\textbf{1965}, \emph{42} (2), 756--766.
\url{https://doi.org/10.1063/1.1696003}.}

\leavevmode\vadjust pre{\hypertarget{ref-mullikenFormulasNumericalTables1949}{}}%
\CSLLeftMargin{(28) }%
\CSLRightInline{Mulliken, R. S.; Rieke, C. A.; Orloff, D.; Orloff, H.
Formulas and {Numerical Tables} for {Overlap Integrals}. \emph{J. Chem.
Phys.} \textbf{1949}, \emph{17} (12), 1248--1267.
\url{https://doi.org/10.1063/1.1747150}.}

\leavevmode\vadjust pre{\hypertarget{ref-murrellSemiempiricalSelfconsistentfieldMolecular1972}{}}%
\CSLLeftMargin{(29) }%
\CSLRightInline{Murrell, J. N.; Harget, A. J. \emph{Semi-Empirical
Self-Consistent-Field Molecular Orbital Theory of Molecules};
{Wiley-Interscience}: {London, New York}, 1972.}

\leavevmode\vadjust pre{\hypertarget{ref-roothaanNewDevelopmentsMolecular1951}{}}%
\CSLLeftMargin{(30) }%
\CSLRightInline{Roothaan, C. C. J. New {Developments} in {Molecular
Orbital Theory}. \emph{Rev. Mod. Phys.} \textbf{1951}, \emph{23} (2),
69--89. \url{https://doi.org/10.1103/RevModPhys.23.69}.}

\leavevmode\vadjust pre{\hypertarget{ref-peachExcitationEnergiesDensity2008}{}}%
\CSLLeftMargin{(31) }%
\CSLRightInline{Peach, M. J. G.; Benfield, P.; Helgaker, T.; Tozer, D.
J. Excitation Energies in Density Functional Theory: {An} Evaluation and
a Diagnostic Test. \emph{The Journal of Chemical Physics} \textbf{2008},
\emph{128} (4), 044118. \url{https://doi.org/10.1063/1.2831900}.}

\leavevmode\vadjust pre{\hypertarget{ref-klessingerExcitedStatesPhotochemistry1995}{}}%
\CSLLeftMargin{(32) }%
\CSLRightInline{Klessinger, M.; Michl, J. \emph{Excited States and
Photochemistry of Organic Molecules}; {VCH}: {New York}, 1995.}

\leavevmode\vadjust pre{\hypertarget{ref-bordenTheoreticalStudiesDiradicals1981}{}}%
\CSLLeftMargin{(33) }%
\CSLRightInline{Borden, W. T.; Davidson, E. R. Theoretical Studies of
Diradicals Containing Four π Electrons. \emph{Acc. Chem. Res.}
\textbf{1981}, \emph{14} (3), 69--76.
\url{https://doi.org/10.1021/ar00063a002}.}

\leavevmode\vadjust pre{\hypertarget{ref-karafiloglouDoubleDynamicSpin1989}{}}%
\CSLLeftMargin{(34) }%
\CSLRightInline{Karafiloglou, P. The Double (or Dynamic) Spin
Polarization in π Diradicals. \emph{J. Chem. Educ.} \textbf{1989},
\emph{66} (10), 816. \url{https://doi.org/10.1021/ed066p816}.}

\leavevmode\vadjust pre{\hypertarget{ref-benamorSpinPolarizationElectronic2020}{}}%
\CSLLeftMargin{(35) }%
\CSLRightInline{Ben Amor, N.; Noûs, C.; Trinquier, G.; Malrieu, J.-P.
Spin Polarization as an Electronic Cooperative Effect. \emph{J. Chem.
Phys.} \textbf{2020}, \emph{153} (4), 044118.
\url{https://doi.org/10.1063/5.0011582}.}

\leavevmode\vadjust pre{\hypertarget{ref-drwalRoleSpinPolarization2023}{}}%
\CSLLeftMargin{(36) }%
\CSLRightInline{Drwal, D.; Matousek, M.; Golub, P.; Tucholska, A.;
Hapka, M.; Brabec, J.; Veis, L.; Pernal, K. The Role of Spin
Polarization and Dynamic Correlation in Singlet-Triplet Gap Inversion of
Heptazine Derivatives. \textbf{2023}.
\url{https://doi.org/10.48550/ARXIV.2307.09075}.}

\leavevmode\vadjust pre{\hypertarget{ref-head-gordonDoublesCorrectionElectronic1994}{}}%
\CSLLeftMargin{(37) }%
\CSLRightInline{Head-Gordon, M.; Rico, R. J.; Oumi, M.; Lee, T. J. A
Doubles Correction to Electronic Excited States from Configuration
Interaction in the Space of Single Substitutions. \emph{Chem. Phys.
Lett.} \textbf{1994}, \emph{219} (1-2), 21--29.
\url{https://doi.org/10.1016/0009-2614(94)00070-0}.}

\leavevmode\vadjust pre{\hypertarget{ref-jornerCoulson}{}}%
\CSLLeftMargin{(38) }%
\CSLRightInline{Jorner, K.
\emph{\href{https://github.com/kjelljorner/coulson}{\textsc{Coulson}}}.}

\leavevmode\vadjust pre{\hypertarget{ref-sunRecentDevelopmentsPySCF2020}{}}%
\CSLLeftMargin{(39) }%
\CSLRightInline{Sun, Q.; Zhang, X.; Banerjee, S.; Bao, P.; Barbry, M.;
Blunt, N. S.; Bogdanov, N. A.; Booth, G. H.; Chen, J.; Cui, Z.-H.;
Eriksen, J. J.; Gao, Y.; Guo, S.; Hermann, J.; Hermes, M. R.; Koh, K.;
Koval, P.; Lehtola, S.; Li, Z.; Liu, J.; Mardirossian, N.; McClain, J.
D.; Motta, M.; Mussard, B.; Pham, H. Q.; Pulkin, A.; Purwanto, W.;
Robinson, P. J.; Ronca, E.; Sayfutyarova, E. R.; Scheurer, M.; Schurkus,
H. F.; Smith, J. E. T.; Sun, C.; Sun, S.-N.; Upadhyay, S.; Wagner, L.
K.; Wang, X.; White, A.; Whitfield, J. D.; Williamson, M. J.; Wouters,
S.; Yang, J.; Yu, J. M.; Zhu, T.; Berkelbach, T. C.; Sharma, S.;
Sokolov, A. Yu.; Chan, G. K.-L. Recent Developments in the {PySCF}
Program Package. \emph{J. Chem. Phys.} \textbf{2020}, \emph{153} (2),
024109. \url{https://doi.org/10.1063/5.0006074}.}

\leavevmode\vadjust pre{\hypertarget{ref-jorgensenTimedependentHartreeFockCalculations1970}{}}%
\CSLLeftMargin{(40) }%
\CSLRightInline{Jørgensen, P.; Linderberg, J. Time-Dependent
{Hartree-Fock} Calculations in the {Pariser-Parr-Pople} Model.
{Applications} to Aniline, Azulene and Pyridine: {APPLICATIONS TO
ANILINE}, {AZULENE AND PYRIDINE}. \emph{Int. J. Quantum Chem.}
\textbf{1970}, \emph{4} (6), 587--602.
\url{https://doi.org/10.1002/qua.560040606}.}

\leavevmode\vadjust pre{\hypertarget{ref-damourGroundExcitedStateDipole2023}{}}%
\CSLLeftMargin{(41) }%
\CSLRightInline{Damour, Y.; Quintero-Monsebaiz, R.; Caffarel, M.;
Jacquemin, D.; Kossoski, F.; Scemama, A.; Loos, P.-F. Ground- and
{Excited-State Dipole Moments} and {Oscillator Strengths} of {Full
Configuration Interaction Quality}. \emph{J. Chem. Theory Comput.}
\textbf{2023}, \emph{19} (1), 221--234.
\url{https://doi.org/10.1021/acs.jctc.2c01111}.}

\leavevmode\vadjust pre{\hypertarget{ref-misurkinElectronicStructureConjugated1974}{}}%
\CSLLeftMargin{(42) }%
\CSLRightInline{Misurkin, I. A.; Ovchinnikov, A. A. The {Electronic
Structure} of {Conjugated Systems} in {Terms} of the
{Pariser}--{Parr}--{Pole Approximation}. \emph{Russ. Chem. Rev.}
\textbf{1974}, \emph{43} (12), 1072--1088.
\url{https://doi.org/10.1070/RC1974v043n12ABEH001881}.}

\end{CSLReferences}

\newpage
\setcounter{page}{1}
\setcounter{section}{0}
\setcounter{figure}{0}
\setcounter{table}{0}
\renewcommand{\thepage}{S\arabic{page}}
\renewcommand{\thesection}{S\arabic{section}}
\renewcommand{\thetable}{S\arabic{table}}
\renewcommand{\thefigure}{S\arabic{figure}}

\hypertarget{supporting-information}{%
\section{Supporting information}\label{supporting-information}}

\begin{figure}

{\centering \includegraphics{../results/figures/global_all.pdf}

}

\caption{\label{fig-global-all}Singlet and triplet excitation energies
as well as oscillator strengths for the rationally designed dataset
\emph{vs.} the reference level.}

\end{figure}

\begin{figure}

{\centering \includegraphics{../results/figures/global_scaffolds_gaps.png}

}

\caption{\label{fig-global-scaffolds-gaps}Singlet-triplet energy gaps
split by scaffold for the rationally designed dataset \emph{vs.} the
reference level.}

\end{figure}

\begin{figure}

{\centering \includegraphics{../results/figures/global_scaffolds_singlet.png}

}

\caption{\label{fig-global-scaffolds-singlet}Singlet excitation energies
split by scaffold for the rationally designed dataset \emph{vs.} the
reference level.}

\end{figure}

\begin{figure}

{\centering \includegraphics{../results/figures/global_scaffolds_triplet.png}

}

\caption{\label{fig-global-scaffolds-triplet}Triplet excitation energies
split by scaffold for the rationally designed dataset \emph{vs.} the
reference level.}

\end{figure}

\begin{figure}

{\centering \includegraphics{../results/figures/global_scaffolds_oscillator.png}

}

\caption{\label{fig-global-scaffolds-oscillator}Oscillator strengths
split by scaffold for the rationally designed dataset \emph{vs.} the
reference level.}

\end{figure}

\begin{figure}

{\centering \includegraphics{../results/figures/local_corr.pdf}

}

\caption{\label{fig-local-corr}Singlet-triplet energy gaps with CIS +
DSP and linear correction for the (a) azaphenalenes and (b) azaazulenes
\emph{vs.} the reference level.}

\end{figure}

\hypertarget{tbl-global-scaffolds}{}
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 22\tabcolsep) * \real{0.0577}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 22\tabcolsep) * \real{0.0865}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 22\tabcolsep) * \real{0.0962}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 22\tabcolsep) * \real{0.0577}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 22\tabcolsep) * \real{0.1058}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 22\tabcolsep) * \real{0.1154}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 22\tabcolsep) * \real{0.0962}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 22\tabcolsep) * \real{0.1442}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 22\tabcolsep) * \real{0.0577}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 22\tabcolsep) * \real{0.0673}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 22\tabcolsep) * \real{0.0577}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 22\tabcolsep) * \real{0.0577}}@{}}
\caption{\label{tbl-global-scaffolds}Metrics split by scaffold for the
rationally designed dataset.}\tabularnewline
\toprule()
\begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
\(R^2\)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
\(\rho\)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
F1
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
ROC-AUC
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
Accuracy
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
Recall
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
Specificity
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
TP
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
TN
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
FP
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
FN
\end{minipage} \\
\midrule()
\endfirsthead
\toprule()
\begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
\(R^2\)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
\(\rho\)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
F1
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
ROC-AUC
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
Accuracy
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
Recall
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
Specificity
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
TP
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
TN
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
FP
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
FN
\end{minipage} \\
\midrule()
\endhead
I & 0.69 & 0.93 & 0.89 & 0.90 & 0.93 & 0.80 & 1.00 & 4 & 10 & 0 & 1 \\
II & 0.33 & 0.52 & - & - & 1.00 & - & 1.00 & 0 & 56 & 0 & 0 \\
III & 0.40 & 0.55 & 0.75 & 0.80 & 0.99 & 0.60 & 1.00 & 3 & 331 & 0 &
2 \\
IV & 0.56 & 0.82 & 0.73 & 0.81 & 0.88 & 0.66 & 0.95 & 39 & 178 & 9 &
20 \\
V & 0.52 & 0.65 & 0.17 & 0.54 & 0.83 & 0.10 & 0.99 & 27 & 1208 & 13 &
245 \\
VI & 0.07 & 0.41 & 0.32 & 0.56 & 0.70 & 0.32 & 0.80 & 9 & 78 & 19 &
19 \\
VII & 0.78 & 0.90 & 0.75 & 0.80 & 0.83 & 0.67 & 0.93 & 2175 & 4860 & 392
& 1085 \\
VIII & 0.56 & 0.76 & 0.52 & 0.80 & 0.72 & 0.92 & 0.69 & 388 & 1470 & 673
& 33 \\
IX & 0.57 & 0.73 & 0.37 & 0.62 & 0.98 & 0.24 & 1.00 & 9 & 1644 & 2 &
29 \\
X & 0.62 & 0.80 & 0.65 & 0.76 & 0.82 & 0.62 & 0.90 & 215 & 841 & 97 &
134 \\
XI & 0.76 & 0.89 & 0.51 & 0.69 & 0.87 & 0.42 & 0.95 & 714 & 8436 & 417 &
972 \\
XII & 0.73 & 0.82 & 0.02 & 0.51 & 0.97 & 0.01 & 1.00 & 5 & 15085 & 17 &
439 \\
XIII & 0.49 & 0.67 & 0.19 & 0.61 & 0.84 & 0.35 & 0.87 & 72 & 3199 & 472
& 135 \\
XIV & 0.82 & 0.90 & 0.45 & 0.70 & 0.93 & 0.44 & 0.96 & 116 & 3556 & 140
& 149 \\
XV & 0.48 & 0.67 & 0.19 & 0.56 & 0.96 & 0.12 & 1.00 & 47 & 10263 & 44 &
353 \\
XVI & 0.77 & 0.88 & 0.22 & 0.60 & 0.96 & 0.22 & 0.98 & 44 & 7377 & 164 &
156 \\
\bottomrule()
\end{longtable}

\end{document}

### 2023-07-29 workflow/notebooks/utils.py
Refactor notebooks. Change threads to tasks

@@ -5,6 +5,7 @@ from numbers import Integral, Real
 from os import PathLike
 from typing import Any
 
+import cairosvg
 import matplotlib as mpl
 import matplotlib.pyplot as plt
 import numpy as np
@@ -12,7 +13,18 @@ import pandas as pd
 import PIL
 import scipy.stats
 import sklearn.metrics
-from coulson.data import HARTREE_TO_EV
+import skunk
+from coulson.data import HARTREE_TO_EV, HARTREE_TO_KCAL
+from coulson.draw import draw_mol, draw_orbital_energies
+from coulson.interface import process_rdkit_mol
+from coulson.ppp import (
+    PPPCalculator,
+    calculate_dsp,
+    calculate_exchange,
+    homo_lumo_overlap,
+)
+from rdkit import Chem
+from rdkit.Chem import AllChem
 from sklearn.linear_model import LinearRegression
 
 AXIS_LABELS: dict[str, str] = {
@@ -407,3 +419,181 @@ def add_to_variables(
         ] = value
 
     return variables_new
+
+
+def generate_orbital_figure(
+    mol_planar: Chem.Mol, ppp: PPPCalculator, exchange: float
+) -> bytes:
+    """Generate composite image of frontier orbitals and exchange value.
+
+    Args:
+        mol_planar: A version of the mol object aligned in the drawing plane
+        ppp: A PPP calculator for the mol
+        exchange: The exchange integral value
+
+    Returns:
+        png_cropped: Composite image as PNG bytes
+    """
+    # Generate orbital figures
+    fig_orbitals, _ = draw_orbital_energies(
+        energies=ppp.orbital_energies * HARTREE_TO_EV,
+        occupations=ppp.occupations,
+        axis_label="Orbital energy (eV)",
+        invert_axis=False,
+        occupation_labels=range(1, ppp.n_orbitals + 1),
+        draw_occupation_labels=True,
+    )
+
+    # Create image of the frontier orbitals
+    svg_homo = draw_mol(
+        mol_planar, properties=ppp.coefficients[ppp.homo_idx], img_format="svg"
+    )
+    svg_lumo = draw_mol(
+        mol_planar, properties=ppp.coefficients[ppp.lumo_idx], img_format="svg"
+    )
+
[…]

### 2023-07-29 workflow/notebooks/utils.py
Fix bugs. Change inner number of threads joblib

@@ -481,9 +481,9 @@ def generate_orbital_figure(
 
 
 def generate_excitation_figure(
-    mol_planar: Chem.Mok,
+    mol_planar: Chem.Mol,
     ppp: PPPCalculator,
-    excitations: dict[tuple(int, int), dict[str, float]],
+    excitations: dict[tuple[int, int], dict[str, float]],
 ):
     """Generate composite image of excitations.
 
@@ -531,7 +531,7 @@ def generate_excitation_figure(
     return png_cropped
 
 
-def compute_mol_qualitative(mol: Chem.Mol) -> tuple(dict[str, float], bytes, bytes):
+def compute_mol_qualitative(mol: Chem.Mol) -> tuple[dict[str, float], bytes, bytes]:
     """Generate excitation data and images for mol.
 
     Args:

### 2023-07-29 manuscript/manuscript.qmd
Add conda env to paper generation

@@
 We used four separate datasets for this study.
-The first is a set of azaphenalenes previously studied by some of us, comprising 257 substituted compounds, for which excitation energies were computed with (2nd-order) algebraic-diagrammatic construction, {{< var reference_level >}}/[Robert: Fill out], and geometries were optimized with B3LYP/[Robert: Fill out][@polliceOrganicMoleculesInverted2021] The second is a set substituted 138 aza-azulenes with {{< var reference_level >}}/[Robert: Fill out] excitation energies and optimized with the B97-3c composite method.[Robert: Add ref to dataset if it exists.] The third is a set of 16 rationally designed scaffolds, which including substitutions amounts to {{< var global_n_total >}} unique compounds, optimized at the B97-3c level and with excitation energies at the {{< var reference_level >}}/[Robert: Fill out] level.[Robert: Add ref to dataset if it exists.] The fourth is a set 315 non-alternant hydrocarbons including substitutions, divided into three subsets of size 76, 187 and 52, optimized at the ωB97X-D/def2-TZVP level and with excitation energies at the CC2/aug-cc-pVTZ level.
+The first is a set of azaphenalenes previously studied by some of us, comprising 257 substituted compounds, for which excitation energies were computed with (2nd-order) algebraic-diagrammatic construction, {{< var reference_level >}}/[Robert: Fill out], and geometries were optimized with B3LYP/[Robert: Fill out][@polliceOrganicMoleculesInverted2021] The second is a set substituted 138 azaazulenes with {{< var reference_level >}}/[Robert: Fill out] excitation energies and optimized with the B97-3c composite method.[Robert: Add ref to dataset if it exists.] The third is a set of 16 rationally designed scaffolds, which including substitutions amounts to {{< var global_n_total >}} unique compounds, optimized at the B97-3c level and with excitation energies at the {{< var reference_level >}}/[Robert: Fill out] level.[Robert: Add ref to dataset if it exists.] The fourth is a set 315 non-alternant hydrocarbons including substitutions, divided into three subsets of size 76, 187 and 52, optimized at the ωB97X-D/def2-TZVP level and with excitation energies at the CC2/aug-cc-pVTZ level.
 
@@
 Adding additional correlation with CIS leads to a $\DeltaEST[DSP]{CIS}$ of {{< var pentalene_ideal_t1_s1_dsp_cis >}} kcal/mol.
-The reference value calculated with {{< var reference_level >}} is {{< var azaphenalene_t1_s1_ref >}} kcal/mol.
+The reference value calculated with {{< var reference_level >}} is {{< var pentalene_ideal_t1_s1_ref >}} kcal/mol.
 The simple perturbation theory model allows us to gain additional insight into the contributions to $\DeltaE[DSP]{SCF}$ as they correspond to single excitations between doubly occupied orbitals below the HOMO and unoccupied orbitals above the LUMO (@fig-pentalene-ideal-excitations).
@@
 
-Azaphenalene is the prototypical molecule that started the renewed investigations into INVEST molecules, and has been the subject of numerous studies with high-level quantum chemical methods.
+Azaphenalene is the prototypical molecule that started the renewed investigations into INVEST molecules,[@desilvaInvertedSingletTriplet2019;@sobolewskiAreHeptazineBasedOrganic2021] and has been the subject of numerous studies with high-level quantum chemical methods.
 Can we capture the inverted gap, as first measured by Leupin and Wirz?[@leupinLowlyingElectronicallyExcited1980] The calculations reveal that the HOMO and LUMO are well separated with spatial overlap of {{< var azaphenalene_overlap >}} and a correspondingly low exchange interaction of {{< var azaphenalene_2_exchange >}} kcal/mol ([@fig-azaphenalene-orbitals]).
@@
 
-While we see success for the azaphenalenes, the situation for the azulenes is unfortunately worse.
+While we see success for the azaphenalenes, the situation for the azaazulenes is unfortunately worse.
 The PPP level S~1~ and T~1~ excitation energies are only moderately correlated with those from {{< var reference_level >}} with $R^2$ of {{< var local_azulenes_r2_singlet >}} and {{< var local_azulenes_r2_triplet >}}, respectively ([@fig-local-azulenes]a,b).
@@
 Unfortunately, the negative gaps are not recovered, leading to recalls of {{< var local_azulenes_recall_cis >}} as none of five inverted molecules could be identified ([@tbl-local-azulenes]).
-It could be speculated that the worse performance for the azulenes comes from the fact that they are non-alternants, for which the approximations are expected to be less applicable.
-Considering the contrasting performance for the azaphenalenes and azulenes, it seems clear that the PPP screening protocol will inevitably miss some hits belonging to particular compound classes.
+It could be speculated that the worse performance for the azaazulenes comes from the fact that they are non-alternants, for which the approximations are expected to be less applicable.
+Considering the contrasting performance for the azaphenalenes and azaazulenes, it seems clear that the PPP screening protocol will inevitably miss some hits belonging to particular compound classes.
 To investigate the global performance of the model we turned to a more diverse set with 16 different compound classes.
@@
 Out of {{< var global_n_total >}} compounds, only {{< var global_n_failed >}} failed, which corresponds to a success rate of {{< var global_percent_success >}}.
-The total runtime using four cores on the M2 processor was {{< var time_cpu_hours >}} CPU hours.
+To put this into context, ~25% of the CASSCF calculations and 9% of the CIS calculations failed in a recent HTVS workflow by Troisi, Padula and co-workers.[@omarIdentificationUnknownInverted2023] The total runtime using four cores on the M2 processor was {{< var time_cpu_hours >}} CPU hours, with the mean time per compounds at {{< var time_mean >}} seconds.
 The S~1~ and T~1~ excitation energies are fairly reproduced with ([@fig-global-all]a,b) $R^2$ of {{< var global_all_r2_singlet >}} and {{< var global_all_r2_triplet >}} vs {{< var reference_level >}}, respectively.
@@
 The specificity simultaneously decreases from {{< var global_all_specificity_cis >}} to {{< var global_all_specificity_cis_corr >}}, but overall the linear correction would be preferable for virtual screening where there is a stronger focus on finding rare hits.
-Applying the linear correction to the azaphenalenes ([@fig-local-corr]a) and ([@fig-local-corr]b) leads to an improvement in the recall of the former from {{< var local_azaphenalenes_recall_cis >}} to {{< var local_azaphenalenes_recall_cis_corr >}} while still none of the inverted azulenes are recovered.
+Applying the linear correction to the azaphenalenes ([@fig-local-corr]a) and ([@fig-local-corr]b) leads to an improvement in the recall of the former from {{< var local_azaphenalenes_recall_cis >}} to {{< var local_azaphenalenes_recall_cis_corr >}} while still none of the inverted azaazulenes are recovered.
 The oscillator strengths are unfortunately poorly correlated for the rational design set, with $R^2=$ {{< var global_all_r2_oscillator >}}, although the Spearman rank correlation is more encouraging at {{< var global_all_rho_oscillator >}} ([@fig-global-all]c).
@@
 Despite these problems, the method is able to capture inverted molecules in {{< var global_scaffolds_n_captured >}} out of the 15 scaffolds where they occur.
-Even for the azulenes included in this screening set at least some candidates are found, in contrast to the findings above for the azaazulenes.
+Even for the azaazulenes included in this screening set at least some candidates are found, in contrast to the findings above for the azaazulenes.
 The per-scaffold oscillator strengths are arguably sufficiently good to allow for local screening in some cases ([@fig-global-scaffolds-oscillator]).
@@
 
+# Data availability
+
+The code to perform all the calculations and to generate the PDF of the manuscript are found at https://https://github.com/kjelljorner/ppp-invest.
+Snakemake[@molderSustainableDataAnalysis2021] was used as workflow manager, ensuring reproducible data generation.
+An archive of the workflow run used to generate this manuscript can be found on https://doi.org/[doi] together with all the output data generated.
+
 # References

### 2023-07-29 manuscript/manuscript.qmd
Fix issue with elsarticle

@@
 
+Data was handled with Pandas,[@mckinneyDataStructuresStatistical2010] chemical data with the RDKit,[@RDKitOpensourceCheminformatics].
+Plots were generated with Matplotlib.[@hunterMatplotlib2DGraphics2007] Numerical calculations used NumPy[@harris2020array] and SciPy[@scipy].
+Calculations and visualizations are heavily indebted to working with Jupyter Notebooks[@grangerJupyterThinkingStorytelling2021] and integrating them into a Snakemake[@molderSustainableDataAnalysis2021] workflow for reproducible computation.
+
 # Results and discussion
@@
 Adding additional correlation with CIS leads to a $\DeltaEST[DSP]{CIS}$ of {{< var pentalene_ideal_t1_s1_dsp_cis >}} kcal/mol.
-The reference value calculated with {{< var reference_level >}} is {{< var pentalene_ideal_t1_s1_ref >}} kcal/mol.
 The simple perturbation theory model allows us to gain additional insight into the contributions to $\DeltaE[DSP]{SCF}$ as they correspond to single excitations between doubly occupied orbitals below the HOMO and unoccupied orbitals above the LUMO (@fig-pentalene-ideal-excitations).

### 2023-07-30 manuscript/manuscript.qmd
Add analysis of negative examples

@@
 High-throughput virtual screening could assist in finding novel classes of molecules, but current efforts are hampered by the high computational cost of the required quantum-chemical methods.
-We present a method based on the semi-empirical Pariser-Parr-Pople theory augmented by perturbation theory and show that it reproduces inverted gaps at a fraction of the cost of the current employed methods.
+We present a method based on the semi-empirical Pariser-Parr-Pople theory augmented by perturbation theory and show that it reproduces inverted gaps at a fraction of the cost of the currently employed methods.
 Our study opens the doors for ultra-high-throughput virtual screening and inverse design to accelerate the discovery and development of the new generation of OLED materials.
@@
 We show that the method is capable of finding hits both in the local chemical space around known scaffolds, and of identifying hits in more diverse datasets.
-While there are clear limitations to the method, we believe that it will open the doors for ultra high-throughput virtual screening (UHTVS) of compound libraries orders of magnitude larger than today.
+While there are clear limitations to the method, we believe that it will open the doors for ultra-high-throughput virtual screening (UHTVS) of compound libraries orders of magnitude larger than today.
 The method is also perfectly applicable as a scoring function in inverse design algorithms.
@@
 \zeta_r = \frac{1280}{501} \gamma_{rr}
-$$
+$${#eq-zeta-r}
 
@@
 \beta_{rs}^\prime = \beta_{rs} \cos{\theta}
-$$
+$${#eq-beta-rs}
 
@@
 O_{xy} = \sum_{r} |c_{r,x}||c_{r,y}|
-$$
+$${#eq-overlap}
 
@@
 Fortunately, the most important excitations have already been identified in the literature long ago.
-Singlet-triplet inversion occurs also for the ground state, where several violations of Hund’s rule are well known, for example for bond-equalized cyclobutadiene at the *D*~4h~ geometry and pentalene at the *D*~2h~ geometry.
+Singlet-triplet inversion occurs also for the ground state, where several violations of Hund’s rule are well known, for example for bond-equalized cyclobutadiene at the *D*~4h~ geometry.
 Borden and Davidson explained this effect in terms of dynamic spin polarization (DSP),[@bordenTheoreticalStudiesDiradicals1981] in which the electrons of a pair of disjoint non-bonding molecular orbitals (NBMOs, [@fig-dsp]a) experience stronger electron correlation effects in the singlet state as compared to the triplet state ([@fig-dsp]b).
-For a pedagogic introduction to the topic, the reader is referred to an excellent article by Karafiloglou[@karafiloglouDoubleDynamicSpin1989] As outlined by Kollmar and Staemmler,[@kollmarViolationHundRule1978] as well as Malrieu and co-workers,[@benamorSpinPolarizationElectronic2020] an alternative view of the same phenomenon is through admixture of singly-excited configurations into the electronic wavefunction of the ground state ([@fig-dsp]c).
+For a pedagogic introduction to the topic, the reader is referred to an excellent article by Karafiloglou.[@karafiloglouDoubleDynamicSpin1989] As outlined by Kollmar and Staemmler,[@kollmarViolationHundRule1978] as well as Malrieu and co-workers,[@benamorSpinPolarizationElectronic2020] an alternative view of the same phenomenon is through admixture of singly-excited configurations into the electronic wavefunction of the ground state ([@fig-dsp]c).
 The DSP effect for the S~1~ excited state can correspondingly be described by admixture of doubly excited configurations,[@kollmarViolationHundRule1978] as they are singly excited with respect to the S~1~ state, and doubly excited with respect to the S~0~ ground state ([@fig-dsp]d).
@@
     \right)
-$$
+$${#eq-dsp-scf}
 
 Here, $K_x$ and $K_y$ are exchange operators, $E(\phi_{\mathrm{S}})$ and $E(\phi_{\mathrm{T}})$ are the energies of the singlet and triplet excited determinants, while $E (\phi_{\mathrm{S}}^{1})$, $E (\phi_{\mathrm{T}}^{1})$ and  $E(\phi_{\mathrm{T}}^{2})$ are the energies of doubly excited determinants.
-They can be written in terms of the correspond orbital energies, $\epsilon_i$, provided the same orbitals are used for both the singlet and triplet states.[@benamorSpinPolarizationElectronic2020]
+They can be written in terms of the corresponding orbital energies, $\epsilon_i$, provided the same orbitals are used for both the singlet and triplet states.[@benamorSpinPolarizationElectronic2020]
 
@@
     \right)
-$$
+$${#eq-dsp-scf-orbital}
 
@@
 \DeltaEST[DSP]{SCF} = 2K + \DeltaE[DSP]{SCF}
-$$
+$${#eq-dest-dsp-scf}
 
@@
     \frac{1}{2} \left( 
-    \frac{3 \Integral{-}}{\epsilon_k + \epsilon_y - \epsilon_y - \epsilon_i - \DeltaE{S,CIS}}
-    - \frac{\Integral{-}}{\epsilon_k + \epsilon_y - \epsilon_y - \epsilon_i - \DeltaE{T,CIS}} \right. \\
-    \left. - \frac{2 \Integral{+}}{\epsilon_k + \epsilon_y - \epsilon_y - \epsilon_i - \DeltaE{T,CIS}}
+    \frac{3 \Integral{-}}{\epsilon_k + \epsilon_y - \epsilon_x - \epsilon_i - \DeltaE{S,CIS}}
+    - \frac{\Integral{-}}{\epsilon_k + \epsilon_y - \epsilon_x - \epsilon_i - \DeltaE{T,CIS}} \right. \\
+    \left. - \frac{2 \Integral{+}}{\epsilon_k + \epsilon_y - \epsilon_x - \epsilon_i - \DeltaE{T,CIS}}
     \right)
 \end{split}
-$$
+$${#eq-dsp-cis}
 
@@
 \DeltaEST[DSP]{CIS} = \DeltaEST{CIS} + \DeltaE[DSP]{CIS}
-$$
+$${#eq-dest-dsp-cis}
 
@@
 We used four separate datasets for this study.
-The first is a set of azaphenalenes previously studied by some of us, comprising 257 substituted compounds, for which excitation energies were computed with (2nd-order) algebraic-diagrammatic construction, {{< var reference_level >}}/[Robert: Fill out], and geometries were optimized with B3LYP/[Robert: Fill out][@polliceOrganicMoleculesInverted2021] The second is a set substituted 138 azaazulenes with {{< var reference_level >}}/[Robert: Fill out] excitation energies and optimized with the B97-3c composite method.[Robert: Add ref to dataset if it exists.] The third is a set of 16 rationally designed scaffolds, which including substitutions amounts to {{< var global_n_total >}} unique compounds, optimized at the B97-3c level and with excitation energies at the {{< var reference_level >}}/[Robert: Fill out] level.[Robert: Add ref to dataset if it exists.] The fourth is a set 315 non-alternant hydrocarbons including substitutions, divided into three subsets of size 76, 187 and 52, optimized at the ωB97X-D/def2-TZVP level and with excitation energies at the CC2/aug-cc-pVTZ level.
+The first is a set of azaphenalenes previously studied by some of us, comprising 256 substituted compounds, for which excitation energies were computed with (2nd-order) algebraic-diagrammatic construction, {{< var reference_level >}}/[Robert: Fill out], and geometries were optimized with B3LYP/[Robert: Fill out][@polliceOrganicMoleculesInverted2021] The second is a set substituted 138 azaazulenes with {{< var reference_level >}}/[Robert: Fill out] excitation energies and optimized with the B97-3c composite method.[Robert: Add ref to dataset if it exists.] The third is a set of 16 rationally designed scaffolds, which including substitutions amounts to {{< var global_n_total >}} unique compounds, optimized at the B97-3c level and with excitation energies at the {{< var reference_level >}}/[Robert: Fill out] level.[Robert: Add ref to dataset if it exists.] The fourth is a set 315 non-alternant hydrocarbons including substitutions, divided into three subsets of size 76, 187 and 52, optimized at the $\omega$B97X-D/def2-TZVP level and with excitation energies at the CC2/aug-cc-pVTZ level.
 
-Data was handled with Pandas,[@mckinneyDataStructuresStatistical2010] chemical data with the RDKit,[@RDKitOpensourceCheminformatics].
+::: {.callout-important}
+Robert: I have 138 azazulenes with data in the csv file but 144 xyz files.
+:::
+
+Data was handled with Pandas,[@mckinneyDataStructuresStatistical2010] and chemical structures with the RDKit[@RDKitOpensourceCheminformatics].
 Plots were generated with Matplotlib.[@hunterMatplotlib2DGraphics2007] Numerical calculations used NumPy[@harris2020array] and SciPy[@scipy].
@@
 Adding additional correlation with CIS leads to a $\DeltaEST[DSP]{CIS}$ of {{< var pentalene_ideal_t1_s1_dsp_cis >}} kcal/mol.
-The simple perturbation theory model allows us to gain additional insight into the contributions to $\DeltaE[DSP]{SCF}$ as they correspond to single excitations between doubly occupied orbitals below the HOMO and unoccupied orbitals above the LUMO (@fig-pentalene-ideal-excitations).
+The simple perturbation theory model allows us to gain additional insight into the contributions to $\DeltaE[DSP]{SCF}$ as they correspond to single excitations from doubly occupied orbitals below the HOMO to unoccupied orbitals above the LUMO (@fig-pentalene-ideal-excitations).
 Three excitations, HOMO-3 → LUMO+2, HOMO-2 → LUMO+1 and HOMO-1 → LUMO+3 are identified as the main contributors to the spin polarization effect, and could potentially be tuned by substituent effects.
@@
 Azaphenalene is the prototypical molecule that started the renewed investigations into INVEST molecules,[@desilvaInvertedSingletTriplet2019;@sobolewskiAreHeptazineBasedOrganic2021] and has been the subject of numerous studies with high-level quantum chemical methods.
-Can we capture the inverted gap, as first measured by Leupin and Wirz?[@leupinLowlyingElectronicallyExcited1980] The calculations reveal that the HOMO and LUMO are well separated with spatial overlap of {{< var azaphenalene_overlap >}} and a correspondingly low exchange interaction of {{< var azaphenalene_2_exchange >}} kcal/mol ([@fig-azaphenalene-orbitals]).
+Can we capture the inverted gap, as first measured by Leupin and Wirz in 1980?[@leupinLowlyingElectronicallyExcited1980] The calculations reveal that the HOMO and LUMO are well separated with spatial overlap of {{< var azaphenalene_overlap >}} and a correspondingly low exchange interaction of only {{< var azaphenalene_2_exchange >}} kcal/mol ([@fig-azaphenalene-orbitals]).
 The larger azaphenalene has 30 excitations that could possibly contribute to the large $\DeltaE[DSP]{SCF}$ of {{< var azaphenalene_dsp_scf >}} kcal/mol, but the most import by far are those from the doubly degenerate HOMO-1 to the doubly degenerate LUMO+1 ([@fig-azaphenalene-excitations]).
@@
 Unfortunately, the oscillator strengths are not as well reproduced with an $R^2$ of {{< var local_azaphenalenes_r2_f >}}, although the Spearman $\rho$ of {{< var local_azaphenalenes_rho_f >}} indicates the values might be used to rank potential candidates ([@fig-local-azaphenalenes]c).
-The problems of obtaining accurate oscillator strengths at the PPP level of theory are well known in the literature, and the problem is especially exacerbated with low oscillator strengths as here.[@misurkinElectronicStructureConjugated1974].
-Crucially, the $\DeltaE{ST}$ shows good correlations, with $R^2=$ {{< var local_azaphenalenes_r2_cis >}} using $\DeltaEST[DSP]{CIS}$ ([@fig-local-azaphenalenes]f).
-This $R^2$ value is essentially unchanged from the gap at the CIS level without DSP ({{< var local_azaphenalenes_r2_cis_raw >}}, [@fig-local-azaphenalenes]d) and markedly better than the gap at the SCF level ({{< var local_azaphenalenes_r2_scf >}}, [@fig-local-azaphenalenes]e), showing the importance of going beyond the SCF level to include at least some configuration interaction.
-The results are of similar or better quality achieved by Pernal and co-workers for different set of azaphenalenes using SCF + DSP with orbitals from DFT.[@drwalRoleSpinPolarization2023] To further analyze the results in terms of a binary classification into normal/inverted, we calculated the true positives (TP), true negatives (TN), false positives (FP) and false negatives (FN), as well as a range of common classification scores (@tbl-local-azaphenalenes).
+The problems of obtaining accurate oscillator strengths at the PPP level of theory are well known in the literature, and the problem is especially exacerbated with low oscillator strengths as here.[@misurkinElectronicStructureConjugated1974] Crucially, the $\DeltaE{ST}$ shows good correlations, with $R^2=$ {{< var local_azaphenalenes_r2_cis >}} using $\DeltaEST[DSP]{CIS}$ ([@fig-local-azaphenalenes]f).
+This $R^2$ value is essentially unchanged from the gap at the CIS level without DSP ({{< var local_azaphenalenes_r2_cis_raw >}}, [@fig-local-azaphenalenes]d) and markedly better than for the gap at the SCF level ({{< var local_azaphenalenes_r2_scf >}}, [@fig-local-azaphenalenes]e), showing the importance of going beyond the SCF level to include at least some configuration interaction.
+The results are of similar or better quality recently achieved by Pernal and co-workers for a different set of azaphenalenes using SCF + DSP with orbitals from DFT.[@drwalRoleSpinPolarization2023] To further analyze the results in terms of a binary classification into normal/inverted, we calculated the true positives (TP), true negatives (TN), false positives (FP) and false negatives (FN), as well as a range of common classification scores (@tbl-local-azaphenalenes).
+For definitions, see [@eq-f1] to [@eq-roc-auc].
 The most important metrics for virtual screening are, in our opinion, the recall and the specificity.
@@
 The specificity measures the fraction of non-inverted molecules that the protocol identifies, and is {{< var local_azaphenalenes_specificity_scf >}} at the SCF level and {{< var local_azaphenalenes_specificity_cis >}} at the CIS level.
-This metrics mean that we are able to capture a large proportion of the molecules with inverted gaps, while filtering out most non-inverted molecules.
-The results for these azaphenalenes clearly indicate that the PPP protocol could be used to prescreen candidates of this compound class for $\DeltaE{ST}$, while further pruning with another method is likely needed for the oscillator strengths.
+These metrics mean that we are able to capture a large proportion of the molecules with inverted gaps, while filtering out most non-inverted molecules.
+The results for these azaphenalenes indicate that the PPP protocol could be used to prescreen candidates of this compound class for $\DeltaE{ST}$, while further pruning with another method is likely needed for the oscillator strengths.
 
@@
 For the oscillator strengths, the situation is rather catastrophic with $R^2$ of {{< var local_azulenes_r2_f >}} implying no correlation whatsoever ([@fig-local-azulenes]c).
-The situation for the $\DeltaE{ST}$ also worse than for the azaphenalenes, with $R^2=$ {{< var local_azulenes_r2_cis >}} with CIS + DSP, a marked improvement on {{< var local_azulenes_r2_scf >}} with SCF + DSP ([@fig-local-azulenes]e,f).
+The situation for the $\DeltaE{ST}$ is also worse than for the azaphenalenes, with $R^2=$ {{< var local_azulenes_r2_cis >}} with CIS + DSP, a marked improvement on {{< var local_azulenes_r2_scf >}} with SCF + DSP ([@fig-local-azulenes]e,f).
 Unfortunately, the negative gaps are not recovered, leading to recalls of {{< var local_azulenes_recall_cis >}} as none of five inverted molecules could be identified ([@tbl-local-azulenes]).
@@
 Here we used a dataset previously generated by some of us using rational design rules, comprising 16 different molecular scaffolds, the details of which will be presented elsewhere.
-Out of {{< var global_n_total >}} compounds, only {{< var global_n_failed >}} failed, which corresponds to a success rate of {{< var global_percent_success >}}.
+Out of {{< var global_n_total >}} compounds, only {{< var global_n_failed >}} failed to compute, which corresponds to a success rate of {{< var global_percent_success >}}%.
 To put this into context, ~25% of the CASSCF calculations and 9% of the CIS calculations failed in a recent HTVS workflow by Troisi, Padula and co-workers.[@omarIdentificationUnknownInverted2023] The total runtime using four cores on the M2 processor was {{< var time_cpu_hours >}} CPU hours, with the mean time per compounds at {{< var time_mean >}} seconds.
@@
 As seen in [@fig-global-all-gaps]a, the $\DeltaE{ST}$ at the SCF + DSP level shows a moderate $R^2$ of {{< var global_all_r2_scf >}}, which increases significantly to {{< var global_all_r2_cis >}} at the CIS + DSP level ([@fig-global-all-gaps]b).
-However, due to a systematic overestimation of $\DeltaE{ST}$, the F1 score is still low at {{< var global_all_f1_cis >}} and a poor recall of {{< var global_all_recall_cis >}} ([@tbl-global-all]).
+However, due to a systematic overestimation of $\DeltaE{ST}$, the F1 score is still low at {{< var global_all_f1_cis >}} and the recall is only {{< var global_all_recall_cis >}} ([@tbl-global-all]).
 We therefore added a linear correction $\DeltaEST[DSP]{CIS,LC}=$ {{< var global_lc_coef >}} $\times\DeltaEST[DSP]{CIS}$ {{< var global_lc_intercept >}}, which increases the F1 score to {{< var global_all_f1_cis_corr >}} and the recall to {{< var global_all_recall_cis_corr >}} ([@fig-global-all-gaps]c).
-The specificity simultaneously decreases from {{< var global_all_specificity_cis >}} to {{< var global_all_specificity_cis_corr >}}, but overall the linear correction would be preferable for virtual screening where there is a stronger focus on finding rare hits.
-Applying the linear correction to the azaphenalenes ([@fig-local-corr]a) and ([@fig-local-corr]b) leads to an improvement in the recall of the former from {{< var local_azaphenalenes_recall_cis >}} to {{< var local_azaphenalenes_recall_cis_corr >}} while still none of the inverted azaazulenes are recovered.
+The specificity simultaneously decreases from {{< var global_all_specificity_cis >}} to {{< var global_all_specificity_cis_corr >}}, but overall the linear correction would be preferable for virtual screening where there is a strong focus on finding rare hits.
+Applying the linear correction to the azaphenalenes ([@fig-local-corr]a) and azaazulenes ([@fig-local-corr]b) leads to an improvement in the recall of the former from {{< var local_azaphenalenes_recall_cis >}} to {{< var local_azaphenalenes_recall_cis_corr >}} while still none of the inverted azaazulenes are recovered.
 The oscillator strengths are unfortunately poorly correlated for the rational design set, with $R^2=$ {{< var global_all_r2_oscillator >}}, although the Spearman rank correlation is more encouraging at {{< var global_all_rho_oscillator >}} ([@fig-global-all]c).
@@
 Despite these problems, the method is able to capture inverted molecules in {{< var global_scaffolds_n_captured >}} out of the 15 scaffolds where they occur.
-Even for the azaazulenes included in this screening set at least some candidates are found, in contrast to the findings above for the azaazulenes.
+Even for the azulenes included in this screening set at least some candidates are found, in contrast to the findings above for the azaazulenes.
 The per-scaffold oscillator strengths are arguably sufficiently good to allow for local screening in some cases ([@fig-global-scaffolds-oscillator]).
-The per-scaffold S~1~ and T~1~ excitation energies can be found in ([@fig-global-scaffolds-singlet]) and ([@fig-global-scaffolds-triplet]), respectively.
-Notable failures are for {{< var I >}} with $R^2$ of {{< var global_scaffolds_singlet_r2_I >}} and {{< var global_scaffolds_triplet_r2_I >}}, respectively, and {{< var XII >}} and {{< var XVI >}}, where some negative T~1~ excitation energies indicate restricted-unrestricted instabilities.
+The per-scaffold S~1~ and T~1~ excitation energies can be found in [@fig-global-scaffolds-singlet] and [@fig-global-scaffolds-triplet], respectively.
+Notable failures are seen for {{< var I >}} with $R^2$ of {{< var global_scaffolds_singlet_r2_I >}} and {{< var global_scaffolds_triplet_r2_I >}}, respectively, and {{< var XII >}} and {{< var XVI >}}, where some negative T~1~ excitation energies indicate restricted-unrestricted instabilities.
+A closer inspection of the dataset reveals that {{< var negative_all_nnegtriplet_rational_design >}} out of {{< var negative_all_ntotal_rational_design >}} calculated compounds show negative triplet excitation energies ({{< var negative_all_negtriplet_rational_design >}}%, see [@tbl-negative]).
+This can be compared to {{< var negative_all_negtriplet_azaphenalenes >}}% for the azaphenalenes and {{< var negative_all_negtriplet_azulenes >}}% for the azulenes ([@fig-negative-all] and [@tbl-negative]).
+Broken down over scaffolds, the results indicate that some of the outliers in the calculated $\DeltaE{ST}$ might be explained this way ([@fig-negative-scaffolds]), although the absolute numbers are very small compared to the total number of compounds.
 
@@
 
-To conclude, we have shown that the Pariser-Parr-Pople theory, a simple semi-empirical $\pi$-electron theory with a minimal basis, can be used to screen for molecules with inverted singlet-triplets energy gaps, both locally and globally.
+To conclude, we have shown that the Pariser-Parr-Pople theory, a simple semi-empirical $\pi$-electron theory with a minimal valence basis, can be used to screen for molecules with inverted singlet-triplets energy gaps, both locally and globally.
 Unfortunately, the method does not seem capable of screening for oscillator strengths with the same accuracy, with exceptions for some scaffolds.
@@
 
-Taken together, we foresee that the presented methodology can be used for ultrahigh-throughput virtual screening campaigns and in inverse design algorithms, followed by curation of hits using more accurate quantum-chemical methods.
+Taken together, we foresee that the presented methodology can be used for ultra-high-throughput virtual screening campaigns and in inverse design algorithms, followed by curation of hits using more accurate quantum-chemical methods.
 Active learning schemes could also be used with machine learning corrections to the PPP singlet-triplet gaps.
@@
 
+::: {.callout-important}
+
+Please add acknowledgements.
+
+:::
+
 # Data availability
 
-The code to perform all the calculations and to generate the PDF of the manuscript are found at https://https://github.com/kjelljorner/ppp-invest.
+The code to perform all the calculations and to generate the PDF of the manuscript are found at <https://https://github.com/kjelljorner/ppp-invest>.
 Snakemake[@molderSustainableDataAnalysis2021] was used as workflow manager, ensuring reproducible data generation.
-An archive of the workflow run used to generate this manuscript can be found on https://doi.org/[doi] together with all the output data generated.
+An archive of the workflow run used to generate this manuscript can be found on <https://doi.org/[doi]> together with all the output data generated.
 
@@
 \newpage
+\appendix
 \setcounter{page}{1}
@@
 \setcounter{table}{0}
+\setcounter{equation}{0}
 \renewcommand{\thepage}{S\arabic{page}}
 \renewcommand{\thesection}{S\arabic{section}}
+\renewcommand{\thefigure}{S\arabic{figure}}
 \renewcommand{\thetable}{S\arabic{table}}
-\renewcommand{\thefigure}{S\arabic{figure}}
+\renewcommand{\theequation}{S\arabic{equation}}
 
 # Supporting information
+
+## Metrics
+
+The F1 score is the harmonic mean of the precision and sensitivity, and is defined as
+
+$$
+\frac{2\mathrm{TP}}{2\mathrm{TP} + \mathrm{FP} + \mathrm{FN}}
+$$ {#eq-f1}
+
+The accuracy is defined as
+
+$$
+\frac{\mathrm{TP} + \mathrm{TN}}{\mathrm{TP} + \mathrm{TN} + \mathrm{FP} + \mathrm{FN}}
+$$ {#eq-accuracy}
+
+The specificity, or true negative rate (TNR), is defined as
+
+$$
+\frac{\mathrm{TN}}{\mathrm{TN} + \mathrm{FP}}
+$$ {#eq-specificity}
+
+The recall, or true positive rate (TPR), is defined as
+$$
+\frac{\mathrm{TP}}{\mathrm{TP} + \mathrm{FN}}
+$$ {#eq-recall}
+
+The fallout, or false positive rate (FPR), is defined as
+$$
+\frac{\mathrm{FP}}{\mathrm{FP} + \mathrm{TN}}
+$$ {#eq-recall}
+
+The ROC-AUC, or the area under the so-called receiver operating characteristic (ROC) curve, is given as
+
+$$
+\int_{x=0}^1 \mathrm{TPR}(\mathrm{FPR}^{-1}(x))dx
+$$ {#eq-roc-auc}
+
+_i.e._, the area under the curve with the FPR on the $x$ axis and TPR on the $y$ axis.
 
@@
 ```
+
+![Singlet-triplet energy gaps with CIS + DSP *vs.* the reference level with points with negative triplet excitation energies marked for the (a) azaphenalenes and (b) azaazulenes (c) rationally designed datasets.]({{< var fig_negative_all >}}.pdf){#fig-negative-all}
+
+![Singlet-triplet energy gaps with CIS + DSP *vs.* the reference level with points with negative triplet excitation energies marked split by scaffold for the rationally designed dataset.]({{< var fig_negative_scaffolds >}}.png){#fig-negative-scaffolds}
+
+```{python}
+# | label: tbl-negative
+# | tbl-cap: Statistics of compounds with negative triplet excitation energies.
+# | echo: false
+import pandas as pd
+from IPython.display import display, Markdown
+
+with open(tab_negative) as f:
+    md = f.read()
+display(Markdown(md))
+```

### 2023-08-06 manuscript/manuscript.qmd
Process Robert comments

@@
       - ref: ac
+      - ref: cifar
 
@@
 
-Organic light-emitting diodes (OLEDs) is a technology for generating light from electricity using organic molecules.[@yangRecentAdvancesOrganic2017] The first generation of OLEDs were limited to an efficiency of max 25% due to spin statistics.
-As the transition T~1~ $\rightarrow$ S~0~, from the lowest excited triplet state to the singlet ground state, is quantum-mechanically forbidden, the OLED molecule can formally only emit from its first excited state of singlet multiplicity, S~1~.
+Organic light-emitting diodes (OLEDs) is a technology for generating light from electricity using organic molecules.[@yangRecentAdvancesOrganic2017] The first generation of OLEDs were based on fluorescent organic molecules and limited to an efficiency of 25% due to spin statistics.
+As the transition T~1~ $\rightarrow$ S~0~, from the lowest excited triplet state to the singlet ground state, is spin-forbidden, the OLED molecule can only emit efficiently from its first excited state of singlet multiplicity, S~1~.
 The second generation of OLEDs exploited ways of increasing the rate of this spin-forbidden phosphorescence.
 The third generation of OLEDs are based on Thermally Activated Delayed Fluorescence (TADF), where the S~1~ state is partially populated from a near-lying T~1~ state.
-For this to happen, the gap $\DeltaE{ST}$ between the states needs to sufficiently small so that the thermal activation competes with non-radiative decay processes from T~1~.
-However, as the T~1~ population is still substantial, problems with stability and lower than ideal intensity due to non-radiative decay persist.
-The logical step for the next generation of OLEDs is to emit directly from an S~1~ state that lies below the T~1~ state, potentially converting all of the electrically generated excitons into photons, _i.e._, 100% efficiency.
+For this to happen, the gap $\DeltaE{ST}$ between the states needs to be sufficiently small so that the thermal activation competes with non-radiative decay processes from T~1~.
+However, as the T~1~ population is still substantial, problems with stability and lower than ideal quantum yields due to non-radiative decay persist.
+The logical step for the next generation of OLEDs is to emit directly from an S~1~ state that lies below the T~1~ state, potentially converting all of the electrically generated excitons into photons, _i.e._, 100% efficiency, while avoiding decay reactions from the T~1~ state.
 
-Molecules that fulfill this criterion have been branded with the acronym INVErted Singlet Triplet energy gap (INVEST) and are exceedingly rare.[@liOrganicMoleculesInverted2022] They are said to violate Hund's rule of maximum multiplicity, which describes that the electronic state with the highest spin (in this case the triplet) should be lowest in energy.
-While examples of molecules violating Hund's rule in the excited state have been known since the 1970s and 1980s,[@kollmarViolationHundRule1978; @leupinLowlyingElectronicallyExcited1980; @kosekiViolationHundMultiplicity1985] it is only recently that they have received considerable attention for application in OLEDs.[@desilvaInvertedSingletTriplet2019;@polliceOrganicMoleculesInverted2021;@aizawaDelayedFluorescenceInverted2022] As they are rare, recent efforts have used high-throughput virtual screening (HTVS) with computational chemistry to identify new compounds with inverted gaps, focusing on expanding the local chemical space around specific scaffolds[@polliceOrganicMoleculesInverted2021;@aizawaDelayedFluorescenceInverted2022] or scanning larger (combinatorially or experimentally derived) datasets to identify novel scaffolds.[@terenceblaskovitsSymmetryInducedSinglet2023;@omarIdentificationUnknownInverted2023;@garnerDoublebondDelocalizationNonalternant2023] These HTVS efforts have been successful at identifying several new INVEST candidates, some of which have also been synthesized and tested experimentally.[@aizawaDelayedFluorescenceInverted2022]
+Molecules that fulfill this criterion have been branded with the acronym INVErted Singlet Triplet energy gap (INVEST) and are exceedingly rare.[@liOrganicMoleculesInverted2022] They are said to violate Hund's rule of maximum multiplicity as applied to the S~1~ and T~1~ states, which describes that the electronic state with the highest spin (in this case the triplet) should be lowest in energy.
+While examples of molecules violating Hund's rule in the excited state have been known since the 1970s and 1980s,[@kollmarViolationHundRule1978; @leupinLowlyingElectronicallyExcited1980; @kosekiViolationHundMultiplicity1985;@toyotaViolationHundRule1988;@toyotaViolationHundMultiplicity1986] it is only recently that they have received considerable attention for application in OLEDs.[@desilvaInvertedSingletTriplet2019;@polliceOrganicMoleculesInverted2021;@aizawaDelayedFluorescenceInverted2022] As they are rare, recent efforts have used high-throughput virtual screening (HTVS) with computational chemistry to identify new compounds with inverted gaps, focusing on expanding the local chemical space around specific scaffolds[@polliceOrganicMoleculesInverted2021;@sanz-rodrigoNegativeSingletTriplet2021;@ricciSingletTripletExcited2021;@sobolewskiAreHeptazineBasedOrganic2021;@aizawaDelayedFluorescenceInverted2022;@ricciEstablishingDesignStrategies2022;@tuckovaOriginInvertedSinglet2022] or scanning larger (combinatorially or experimentally derived) datasets to identify novel scaffolds.[@terenceblaskovitsSymmetryInducedSinglet2023;@omarIdentificationUnknownInverted2023;@garnerDoublebondDelocalizationNonalternant2023] These HTVS efforts have been successful at identifying several new INVEST candidates, some of which have also been synthesized and tested experimentally.[@aizawaDelayedFluorescenceInverted2022]
 
@@
 We show that the method is capable of finding hits both in the local chemical space around known scaffolds, and of identifying hits in more diverse datasets.
-While there are clear limitations to the method, we believe that it will open the doors for ultra-high-throughput virtual screening (UHTVS) of compound libraries orders of magnitude larger than today.
+While there are clear limitations to the method, we believe that it will open the doors for ultra-high-throughput virtual screening (UHTVS) of compound libraries several orders of magnitude larger than today.
 The method is also perfectly applicable as a scoring function in inverse design algorithms.
@@
 
-While the recent literature has focused on applying high-level *ab initio* and DFT methods to study INVEST compounds, we believe that some conceptual clarity has been lost in process.
+While the recent literature has focused on applying high-level *ab initio* and DFT methods to study INVEST compounds, we believe that some conceptual clarity has been lost in the process.
 Instead, we apply the simplest possible electronic structure method that can capture the physics of the problem.
@@
 
-where $\alpha_{r}$ is the *one-center core resonance integral*, $U_r$ is a parameter of model called the *atomic valence state potential*, $Z_s$ is the effective nuclear charge and $\beta_{rs}$ is the *two-center resonance integral*, another parameter of the model.
+where $\alpha_{r}$ is the *one-center core resonance integral*, $U_r$ is a parameter of the model called the *atomic valence state potential*, $Z_s$ is the effective nuclear charge and $\beta_{rs}$ is the *two-center resonance integral*, another parameter of the model.
 
@@
 where $K$ is the exchange integral between the two orbitals involved in the single excitation (normally the HOMO and the LUMO).
-As defined here, a positive $\DeltaE{ST}$ corresponds to the usual situation with the triplet being lower in energy than the singlet, while a positive $\DeltaE{ST}$ corresponds to an inverted gap.
+As defined here, a positive $\DeltaE{ST}$ corresponds to the usual situation with the triplet being lower in energy than the singlet, while a negative $\DeltaE{ST}$ corresponds to an inverted gap.
 The exchange integral depends strongly on the spatial overlap between the the two orbitals, which in the ZDO approximation can be computed as:[@peachExcitationEnergiesDensity2008]
@@
 When the overlap is zero, the exchange interaction vanishes and $\DeltaEST{SCF}$ is zero.
-Unfortunately, so is the oscillator strength, $f$, between the two states,[@klessingerExcitedStatesPhotochemistry1995] which is non-optimal for OLED materials that should emit light.
-Consequently, there is a trade-off between having a small HOMO-LUMO overlap to reduce the exchange interaction, while still maintaining a sufficient oscillator strength.
+Unfortunately, so is the oscillator strength, $f$, between the S~1~ excited state and the S~0~ ground state,[@klessingerExcitedStatesPhotochemistry1995] which is non-optimal for OLED materials that should emit light.[@gomez-bombarelliDesignEfficientMolecular2016] Consequently, there is a trade-off between having a small HOMO-LUMO overlap to reduce the exchange interaction, while still maintaining a sufficient oscillator strength.
 
@@
 We used four separate datasets for this study.
-The first is a set of azaphenalenes previously studied by some of us, comprising 256 substituted compounds, for which excitation energies were computed with (2nd-order) algebraic-diagrammatic construction, {{< var reference_level >}}/[Robert: Fill out], and geometries were optimized with B3LYP/[Robert: Fill out][@polliceOrganicMoleculesInverted2021] The second is a set substituted 138 azaazulenes with {{< var reference_level >}}/[Robert: Fill out] excitation energies and optimized with the B97-3c composite method.[Robert: Add ref to dataset if it exists.] The third is a set of 16 rationally designed scaffolds, which including substitutions amounts to {{< var global_n_total >}} unique compounds, optimized at the B97-3c level and with excitation energies at the {{< var reference_level >}}/[Robert: Fill out] level.[Robert: Add ref to dataset if it exists.] The fourth is a set 315 non-alternant hydrocarbons including substitutions, divided into three subsets of size 76, 187 and 52, optimized at the $\omega$B97X-D/def2-TZVP level and with excitation energies at the CC2/aug-cc-pVTZ level.
-
-::: {.callout-important}
-Robert: I have 138 azazulenes with data in the csv file but 144 xyz files.
-:::
+The first is a set of azaphenalenes previously studied by some of us, comprising 256 substituted compounds, for which excitation energies were computed with (2nd-order) algebraic-diagrammatic construction, {{< var reference_level >}}/cc-pVDZ, and geometries were optimized with B3LYP/cc-pVDZ[@polliceOrganicMoleculesInverted2021] The second is a set substituted 138 azaazulenes with {{< var reference_level >}}/cc-pVDZ excitation energies and optimized with the B97-3c composite method.[@nigamArtificialDesignOrganic2023] The third is a set of 16 rationally designed scaffolds, which including substitutions amounts to {{< var global_n_total >}} unique compounds, optimized at the B97-3c level and with excitation energies at the {{< var reference_level >}}/cc-pVDZ level.[@polliceRationalDesignOrganic2023] The fourth is a set 315 non-alternant hydrocarbons including substitutions, divided into three subsets of size 76, 187 and 52, optimized at the $\omega$B97X-D/def2-TZVP level and with excitation energies at the CC2/aug-cc-pVTZ level.
 
@@
 
-## Qualitative understanding of the inverted gap
+## Orbital decomposition of the inverted gap
 
@@
 Three excitations, HOMO-3 → LUMO+2, HOMO-2 → LUMO+1 and HOMO-1 → LUMO+3 are identified as the main contributors to the spin polarization effect, and could potentially be tuned by substituent effects.
-
-![Orbitals energies and frontier orbitals of pentalene at the idealized $D_\mathrm{2h}$ geometry and bond lengths of 1.4 Å.]({{< var fig_pentalene_ideal_orbitals >}}.png){#fig-pentalene-ideal-orbitals}
 
@@
 
-![Orbitals energies and frontier orbitals of pentalene at the optimized $C_\mathrm{2h}$ geometry.]({{< var fig_pentalene_orbitals >}}.png){#fig-pentalene-orbitals}
-
 ![Excitations and their contribution to $\DeltaE[DSP]{SCF}=$ {{< var pentalene_dsp_scf >}} kcal/mol for pentalene at the optimized $C_\mathrm{2h}$ geometry.]({{< var fig_pentalene_excitations >}}.png){#fig-pentalene-excitations}
@@
 Can we capture the inverted gap, as first measured by Leupin and Wirz in 1980?[@leupinLowlyingElectronicallyExcited1980] The calculations reveal that the HOMO and LUMO are well separated with spatial overlap of {{< var azaphenalene_overlap >}} and a correspondingly low exchange interaction of only {{< var azaphenalene_2_exchange >}} kcal/mol ([@fig-azaphenalene-orbitals]).
-The larger azaphenalene has 30 excitations that could possibly contribute to the large $\DeltaE[DSP]{SCF}$ of {{< var azaphenalene_dsp_scf >}} kcal/mol, but the most import by far are those from the doubly degenerate HOMO-1 to the doubly degenerate LUMO+1 ([@fig-azaphenalene-excitations]).
+The larger azaphenalene has 30 excitations that could possibly contribute to the large $\DeltaE[DSP]{SCF}$ of {{< var azaphenalene_dsp_scf >}} kcal/mol, but the most important by far are those from the doubly degenerate HOMO-1 to the doubly degenerate LUMO+1 ([@fig-azaphenalene-excitations]).
 In total, $\DeltaEST[DSP]{SCF}=$ {{< var azaphenalene_t1_s1_dsp_scf >}} kcal/mol, which is slightly smaller at the CIS level with $\DeltaEST[DSP]{CIS}=$ {{< var azaphenalene_t1_s1_dsp_cis >}} kcal/mol.
@@
 For these two compounds, the calculated $\DeltaEST[DSP]{}$ are more negative at the SCF level compared to CIS.
-Compared to {{< var reference_level >}}, it would seem that CIS is preferable, but to elucidate to which method is better, we now turn to larger datasets.
-
-![Orbitals energies and frontier orbitals of azaphenalene.]({{< var fig_azaphenalene_orbitals >}}.png){#fig-azaphenalene-orbitals}
+Compared to {{< var reference_level >}}, it would seem that CIS is preferable, but to elucidate which method is better, we now turn to larger datasets.
 
@@
 Unfortunately, the oscillator strengths are not as well reproduced with an $R^2$ of {{< var local_azaphenalenes_r2_f >}}, although the Spearman $\rho$ of {{< var local_azaphenalenes_rho_f >}} indicates the values might be used to rank potential candidates ([@fig-local-azaphenalenes]c).
-The problems of obtaining accurate oscillator strengths at the PPP level of theory are well known in the literature, and the problem is especially exacerbated with low oscillator strengths as here.[@misurkinElectronicStructureConjugated1974] Crucially, the $\DeltaE{ST}$ shows good correlations, with $R^2=$ {{< var local_azaphenalenes_r2_cis >}} using $\DeltaEST[DSP]{CIS}$ ([@fig-local-azaphenalenes]f).
+The problem of obtaining accurate oscillator strengths at the PPP level of theory is well known in the literature, and is especially exacerbated with low oscillator strengths which are prevalent in this dataset.[@misurkinElectronicStructureConjugated1974] Crucially, the $\DeltaE{ST}$ shows good correlations, with $R^2=$ {{< var local_azaphenalenes_r2_cis >}} using $\DeltaEST[DSP]{CIS}$ ([@fig-local-azaphenalenes]f).
 This $R^2$ value is essentially unchanged from the gap at the CIS level without DSP ({{< var local_azaphenalenes_r2_cis_raw >}}, [@fig-local-azaphenalenes]d) and markedly better than for the gap at the SCF level ({{< var local_azaphenalenes_r2_scf >}}, [@fig-local-azaphenalenes]e), showing the importance of going beyond the SCF level to include at least some configuration interaction.
@@
 These metrics mean that we are able to capture a large proportion of the molecules with inverted gaps, while filtering out most non-inverted molecules.
+Below we will show how we can improve the results even more with a linear correction to $\DeltaEST[DSP]{CIS}$ based on results from a wider set of compounds.
 The results for these azaphenalenes indicate that the PPP protocol could be used to prescreen candidates of this compound class for $\DeltaE{ST}$, while further pruning with another method is likely needed for the oscillator strengths.
 
-![Excitation properties for azaphenalenes _vs._ the reference level.
+![Excitation properties for azaphenalenes against the reference level.
 (a) Singlet excitation energies.
@@
 For the oscillator strengths, the situation is rather catastrophic with $R^2$ of {{< var local_azulenes_r2_f >}} implying no correlation whatsoever ([@fig-local-azulenes]c).
-The situation for the $\DeltaE{ST}$ is also worse than for the azaphenalenes, with $R^2=$ {{< var local_azulenes_r2_cis >}} with CIS + DSP, a marked improvement on {{< var local_azulenes_r2_scf >}} with SCF + DSP ([@fig-local-azulenes]e,f).
+The situation for the $\DeltaE{ST}$ is also worse than for the azaphenalenes, with $R^2=$ {{< var local_azulenes_r2_cis >}} with CIS + DSP, a marked improvement over {{< var local_azulenes_r2_scf >}} with SCF + DSP ([@fig-local-azulenes]e,f).
 Unfortunately, the negative gaps are not recovered, leading to recalls of {{< var local_azulenes_recall_cis >}} as none of five inverted molecules could be identified ([@tbl-local-azulenes]).
@@
 
-![Excitation properties for azaazulenes _vs._ the reference level.
+![Excitation properties for azaazulenes against the reference level.
 (a) Singlet excitation energies.
@@
 
-Here we used a dataset previously generated by some of us using rational design rules, comprising 16 different molecular scaffolds, the details of which will be presented elsewhere.
-Out of {{< var global_n_total >}} compounds, only {{< var global_n_failed >}} failed to compute, which corresponds to a success rate of {{< var global_percent_success >}}%.
-To put this into context, ~25% of the CASSCF calculations and 9% of the CIS calculations failed in a recent HTVS workflow by Troisi, Padula and co-workers.[@omarIdentificationUnknownInverted2023] The total runtime using four cores on the M2 processor was {{< var time_cpu_hours >}} CPU hours, with the mean time per compounds at {{< var time_mean >}} seconds.
-The S~1~ and T~1~ excitation energies are fairly reproduced with ([@fig-global-all]a,b) $R^2$ of {{< var global_all_r2_singlet >}} and {{< var global_all_r2_triplet >}} vs {{< var reference_level >}}, respectively.
+In this section, we used a dataset previously generated by some of us using rational design rules, comprising 16 different molecular scaffolds, the details of which will be presented elsewhere.
+Out of {{< var global_n_total >}} compounds, only {{< var global_n_failed >}} cyclobutadienes failed to compute as they had rearranged in the DFT simulations and our topology detection algorithm interpreted them as having formed transannular single bond.
+This corresponds to a success rate of {{< var global_percent_success >}}%.
+To put this into context, ~25% of the CASSCF calculations and 9% of the CIS calculations failed in a recent HTVS workflow by Troisi, Padula and co-workers.[@omarIdentificationUnknownInverted2023] The total runtime using four cores on the M2 processor was {{< var time_cpu_hours >}} CPU hours, with a mean time per compounds of {{< var time_mean >}} seconds.
+The S~1~ and T~1~ excitation energies are fairly reproduced with ([@fig-global-all]a,b) $R^2$ of {{< var global_all_r2_singlet >}} and {{< var global_all_r2_triplet >}} against {{< var reference_level >}}, respectively.
 As seen in [@fig-global-all-gaps]a, the $\DeltaE{ST}$ at the SCF + DSP level shows a moderate $R^2$ of {{< var global_all_r2_scf >}}, which increases significantly to {{< var global_all_r2_cis >}} at the CIS + DSP level ([@fig-global-all-gaps]b).
@@
 The per-scaffold S~1~ and T~1~ excitation energies can be found in [@fig-global-scaffolds-singlet] and [@fig-global-scaffolds-triplet], respectively.
-Notable failures are seen for {{< var I >}} with $R^2$ of {{< var global_scaffolds_singlet_r2_I >}} and {{< var global_scaffolds_triplet_r2_I >}}, respectively, and {{< var XII >}} and {{< var XVI >}}, where some negative T~1~ excitation energies indicate restricted-unrestricted instabilities.
+Notable failures are seen for {{< var I >}} with $R^2$ of {{< var global_scaffolds_singlet_r2_I >}} and {{< var global_scaffolds_triplet_r2_I >}}, respectively.
+Additionally, {{< var XII >}} and {{< var XVI >}} show some negative T~1~ excitation energies, indicating restricted-unrestricted instabilities.
 A closer inspection of the dataset reveals that {{< var negative_all_nnegtriplet_rational_design >}} out of {{< var negative_all_ntotal_rational_design >}} calculated compounds show negative triplet excitation energies ({{< var negative_all_negtriplet_rational_design >}}%, see [@tbl-negative]).
@@
 
-![Singlet-triplet energy gaps for rationally design set _vs._ reference level for (a) SCF + DSP, (b) CIS + DSP and (c) CIS + DSP and linear correction.]({{< var fig_global_all_gaps >}}.pdf){#fig-global-all-gaps}
+![Singlet-triplet energy gaps for rational design set against reference level for (a) SCF + DSP, (b) CIS + DSP and (c) CIS + DSP and linear correction.]({{< var fig_global_all_gaps >}}.pdf){#fig-global-all-gaps}
 
@@
 
-![Singlet-triplet energy gaps for CIS + DSP and linear correction _vs._ reference level for the subsets of (a) non-alternants (b) non-alternants constrained to higher symmetry and (c) substituted non-alternants.
+![Singlet-triplet energy gaps for CIS + DSP and linear correction against reference level for the subsets of (a) non-alternants (b) non-alternants constrained to higher symmetry and (c) substituted non-alternants.
 Compounds contained in the rationally designed sets are marked with dashed circles and those with negative triplet excitation energies are marked with a semi-transparent red circle.]({{< var fig_garner_gaps >}}.pdf){#fig-garner-gaps}
@@
 Further limitations include (4) lack of solvation model and need for (5) already optimized geometries.
-We believe that at at least some of these limitations could be mitigated by a reparametrization specifically targeting inverted gaps, while geometries from the fast GFN-xTB method would be suitable.
+We believe that at least some of these limitations could be mitigated by a reparametrization specifically targeting inverted gaps, while geometries from the fast GFN-xTB method would be suitable.
 
@@
 Frøseth.
+This research was undertaken thanks in part to funding provided to the Acceleration Consortium of the University of Toronto from the Canada First Research Excellence Fund.
 
@@
 
-The code to perform all the calculations and to generate the PDF of the manuscript are found at <https://https://github.com/kjelljorner/ppp-invest>.
+The code to perform all the calculations and to generate the PDF of the manuscript using the Quarto publishing system are found at <https://https://github.com/kjelljorner/ppp-invest>.
 Snakemake[@molderSustainableDataAnalysis2021] was used as workflow manager, ensuring reproducible data generation.
@@
 
-The F1 score is the harmonic mean of the precision and sensitivity, and is defined as
+For definitions, see the Wikipedia article on Precision and Recall.[@PrecisionRecall2023] The F1 score is the harmonic mean of the precision and sensitivity, and is defined as
 
@@
 
-![Singlet and triplet excitation energies as well as oscillator strengths for the rationally designed dataset *vs.* the reference level.]({{< var fig_global_all >}}.pdf){#fig-global-all}
+![Orbital energies and frontier orbitals of pentalene at the idealized $D_\mathrm{2h}$ geometry and bond lengths of 1.4 Å.]({{< var fig_pentalene_ideal_orbitals >}}.png){#fig-pentalene-ideal-orbitals}
 
-![Singlet-triplet energy gaps split by scaffold for the rationally designed dataset *vs.* the reference level.]({{< var fig_global_scaffolds_gaps >}}.png){#fig-global-scaffolds-gaps}
+![Orbital energies and frontier orbitals of pentalene at the optimized $C_\mathrm{2h}$ geometry.]({{< var fig_pentalene_orbitals >}}.png){#fig-pentalene-orbitals}
 
-![Singlet excitation energies split by scaffold for the rationally designed dataset *vs.* the reference level.]({{< var fig_global_scaffolds_singlet >}}.png){#fig-global-scaffolds-singlet}
+![Orbital energies and frontier orbitals of azaphenalene.]({{< var fig_azaphenalene_orbitals >}}.png){#fig-azaphenalene-orbitals}
 
-![Triplet excitation energies split by scaffold for the rationally designed dataset *vs.* the reference level.]({{< var fig_global_scaffolds_triplet >}}.png){#fig-global-scaffolds-triplet}
+![Singlet and triplet excitation energies as well as oscillator strengths for the rationally designed dataset against the reference level.]({{< var fig_global_all >}}.pdf){#fig-global-all}
 
-![Oscillator strengths split by scaffold for the rationally designed dataset *vs.* the reference level.]({{< var fig_global_scaffolds_oscillator >}}.png){#fig-global-scaffolds-oscillator}
+![Singlet-triplet energy gaps split by scaffold for the rationally designed dataset against the reference level.]({{< var fig_global_scaffolds_gaps >}}.png){#fig-global-scaffolds-gaps}
 
-![Singlet-triplet energy gaps with CIS + DSP and linear correction for the (a) azaphenalenes and (b) azaazulenes *vs.* the reference level.]({{< var fig_local_corr >}}.pdf){#fig-local-corr}
+![Singlet excitation energies split by scaffold for the rationally designed dataset against the reference level.]({{< var fig_global_scaffolds_singlet >}}.png){#fig-global-scaffolds-singlet}
+
+![Triplet excitation energies split by scaffold for the rationally designed dataset against the reference level.]({{< var fig_global_scaffolds_triplet >}}.png){#fig-global-scaffolds-triplet}
+
+![Oscillator strengths split by scaffold for the rationally designed dataset against the reference level.]({{< var fig_global_scaffolds_oscillator >}}.png){#fig-global-scaffolds-oscillator}
+
+![Singlet-triplet energy gaps with CIS + DSP and linear correction for the (a) azaphenalenes and (b) azaazulenes against the reference level.]({{< var fig_local_corr >}}.pdf){#fig-local-corr}
 
@@
 
-![Singlet-triplet energy gaps with CIS + DSP *vs.* the reference level with points with negative triplet excitation energies marked for the (a) azaphenalenes and (b) azaazulenes (c) rationally designed datasets.]({{< var fig_negative_all >}}.pdf){#fig-negative-all}
+![Singlet-triplet energy gaps with CIS + DSP against the reference level with points with negative triplet excitation energies marked for the (a) azaphenalenes and (b) azaazulenes (c) rationally designed datasets.]({{< var fig_negative_all >}}.pdf){#fig-negative-all}
 
-![Singlet-triplet energy gaps with CIS + DSP *vs.* the reference level with points with negative triplet excitation energies marked split by scaffold for the rationally designed dataset.]({{< var fig_negative_scaffolds >}}.png){#fig-negative-scaffolds}
+![Singlet-triplet energy gaps with CIS + DSP against the reference level with points with negative triplet excitation energies marked split by scaffold for the rationally designed dataset.]({{< var fig_negative_scaffolds >}}.png){#fig-negative-scaffolds}
 

### 2023-08-06 workflow/notebooks/utils.py

@@ -23,6 +23,7 @@ from coulson.ppp import (
     calculate_exchange,
     homo_lumo_overlap,
 )
+from numpy.typing import ArrayLike
 from rdkit import Chem
 from rdkit.Chem import AllChem
 from sklearn.linear_model import LinearRegression
@@ -109,6 +110,24 @@ ALTERNANT: dict[str, str] = {
     "XVI": "non-alternant",
 }
 """Classification into alternant/non-alternant for the parent compounds of the rational design set."""
+
+
+def f1_score(y_true: ArrayLike, y_pred: ArrayLike):
+    """Return F1 score which does not break when tp and fp are zero.
+
+    Args:
+        y_true: True target values
+        y_pred: Predicted target values
+
+    Returns:
+        f1: F1 score
+    """
+    tn, fp, fn, tp = sklearn.metrics.confusion_matrix(
+        y_true, y_pred, labels=[False, True]
+    ).ravel()
+    f1 = 2 * tp / (2 * tp + fp + fn)
+
+    return f1
 
 
 def plot_zero_zero(
@@ -208,7 +227,7 @@ def plot_zero_zero(
             roc_auc = sklearn.metrics.roc_auc_score(y_true, y_pred)
         except ValueError:
             roc_auc = np.nan
-        f1 = sklearn.metrics.f1_score(y_true, y_pred, zero_division=np.nan)
+        f1 = f1_score(y_true, y_pred)
         accuracy = sklearn.metrics.accuracy_score(y_true, y_pred)
         recall = sklearn.metrics.recall_score(y_true, y_pred, zero_division=np.nan)
 

### 2023-09-22 manuscript/manuscript.qmd
paper update

@@
     orcid: 0000-0002-8277-4434
+    email: aspuru@utoronto.ca
+    corresponding: true    
     affiliations:
@@
 
-title: "Ultrafast computational screening of molecules with inverted singlet-triplet energy gaps using semi-empirical quantum chemistry"
+title: "Ultrafast computational screening of molecules with inverted singlet-triplet energy gaps using the Pariser-Parr-Pople semi-empirical quantum chemistry method"
 abstract: |
-  Molecules with an inverted energy gap between their first excited states of singlet and triplet multiplicity have promising applications in the next generation of organic light-emitting diode (OLED) materials.
+  Molecules with an inverted energy gap between their first singlet and triplet excited states have promising applications in the next generation of organic light-emitting diode (OLED) materials.
 Unfortunately, such molecules are rare and only a handful of examples are currently known.
-High-throughput virtual screening could assist in finding novel classes of molecules, but current efforts are hampered by the high computational cost of the required quantum-chemical methods.
-We present a method based on the semi-empirical Pariser-Parr-Pople theory augmented by perturbation theory and show that it reproduces inverted gaps at a fraction of the cost of the currently employed methods.
-Our study opens the doors for ultra-high-throughput virtual screening and inverse design to accelerate the discovery and development of the new generation of OLED materials.
+High-throughput virtual screening could assist in finding novel classes of these molecules, but current efforts are hampered by the high computational cost of the required quantum-chemical methods.
+We present a method based on the semi-empirical Pariser-Parr-Pople theory augmented by perturbation theory and show that it reproduces inverted gaps at a fraction of the cost of currently employed excited state calculations..
+Our study paves the way for ultra-high-throughput virtual screening and inverse design to accelerate the discovery and development of this new generation of OLED materials.
 bibliography: [references.bib] 
@@
       name: "Journal of Physical Chemistry A"
-      model: "3p"    
+      model: "1p"    
     number-sections: false
-    keep-tex: false            
+    keep-tex: true            
     include-in-header:
@@
 
-Organic light-emitting diodes (OLEDs) is a technology for generating light from electricity using organic molecules.[@yangRecentAdvancesOrganic2017] The first generation of OLEDs were based on fluorescent organic molecules and limited to an efficiency of 25% due to spin statistics.
+Organic light-emitting diodes (OLEDs) is a technology for generating light from electricity using organic molecules.[@yangRecentAdvancesOrganic2017] The first generation of OLEDs were based on fluorescent organic molecules with a maximum efficiency of 25% due to spin statistics.
 As the transition T~1~ $\rightarrow$ S~0~, from the lowest excited triplet state to the singlet ground state, is spin-forbidden, the OLED molecule can only emit efficiently from its first excited state of singlet multiplicity, S~1~.
@@
 However, as the T~1~ population is still substantial, problems with stability and lower than ideal quantum yields due to non-radiative decay persist.
-The logical step for the next generation of OLEDs is to emit directly from an S~1~ state that lies below the T~1~ state, potentially converting all of the electrically generated excitons into photons, _i.e._, 100% efficiency, while avoiding decay reactions from the T~1~ state.
+The logical step for the next generation of OLEDs is to emit directly from an S~1~ state that lies below the T~1~ state, potentially converting all of the electrically generated excitons into photons, _i.e._, 100% internal quantum efficiency, while avoiding decay reactions from the T~1~ state.
 
-Molecules that fulfill this criterion have been branded with the acronym INVErted Singlet Triplet energy gap (INVEST) and are exceedingly rare.[@liOrganicMoleculesInverted2022] They are said to violate Hund's rule of maximum multiplicity as applied to the S~1~ and T~1~ states, which describes that the electronic state with the highest spin (in this case the triplet) should be lowest in energy.
-While examples of molecules violating Hund's rule in the excited state have been known since the 1970s and 1980s,[@kollmarViolationHundRule1978; @leupinLowlyingElectronicallyExcited1980; @kosekiViolationHundMultiplicity1985;@toyotaViolationHundRule1988;@toyotaViolationHundMultiplicity1986] it is only recently that they have received considerable attention for application in OLEDs.[@desilvaInvertedSingletTriplet2019;@polliceOrganicMoleculesInverted2021;@aizawaDelayedFluorescenceInverted2022] As they are rare, recent efforts have used high-throughput virtual screening (HTVS) with computational chemistry to identify new compounds with inverted gaps, focusing on expanding the local chemical space around specific scaffolds[@polliceOrganicMoleculesInverted2021;@sanz-rodrigoNegativeSingletTriplet2021;@ricciSingletTripletExcited2021;@sobolewskiAreHeptazineBasedOrganic2021;@aizawaDelayedFluorescenceInverted2022;@ricciEstablishingDesignStrategies2022;@tuckovaOriginInvertedSinglet2022] or scanning larger (combinatorially or experimentally derived) datasets to identify novel scaffolds.[@terenceblaskovitsSymmetryInducedSinglet2023;@omarIdentificationUnknownInverted2023;@garnerDoublebondDelocalizationNonalternant2023] These HTVS efforts have been successful at identifying several new INVEST candidates, some of which have also been synthesized and tested experimentally.[@aizawaDelayedFluorescenceInverted2022]
+Molecules with an INVErted Singlet-Triplet energy gap (INVEST) are exceedingly rare.[@liOrganicMoleculesInverted2022] They are said to violate Hund's rule of maximum multiplicity as applied to the S~1~ and T~1~ states, which describes that the electronic state with the highest spin (in this case the triplet) should be lowest in energy.
+While a handful of examples of molecules violating Hund's rule in the excited state have been known since the 1970s and 1980s,[@kollmarViolationHundRule1978; @leupinLowlyingElectronicallyExcited1980; @kosekiViolationHundMultiplicity1985;@toyotaViolationHundRule1988;@toyotaViolationHundMultiplicity1986] it is only recently that they have received considerable attention for application in OLEDs.[@desilvaInvertedSingletTriplet2019;@polliceOrganicMoleculesInverted2021;@aizawaDelayedFluorescenceInverted2022] As they are rare, recent efforts have used high-throughput virtual screening (HTVS) with computational chemistry to identify new compounds with inverted gaps, focusing on expanding the local chemical space around specific scaffolds[@polliceOrganicMoleculesInverted2021;@sanz-rodrigoNegativeSingletTriplet2021;@ricciSingletTripletExcited2021;@sobolewskiAreHeptazineBasedOrganic2021;@aizawaDelayedFluorescenceInverted2022;@ricciEstablishingDesignStrategies2022;@tuckovaOriginInvertedSinglet2022] or scanning larger (combinatorially or experimentally derived) datasets to identify novel scaffolds.[@terenceblaskovitsSymmetryInducedSinglet2023;@omarIdentificationUnknownInverted2023;@garnerDoublebondDelocalizationNonalternant2023] These HTVS efforts have been successful at identifying several new INVEST candidates, some of which have also been synthesized and tested experimentally.[@aizawaDelayedFluorescenceInverted2022]
 
@@
 At the same time, it retains the conceptual clarity of Hückel theory and allows for straightforward interpretation of the inverted gap in terms of the well-known concept of dynamic spin polarization (DSP).
-We show that the method is capable of finding hits both in the local chemical space around known scaffolds, and of identifying hits in more diverse datasets.
+We show that the method is capable of finding promising hits both in the local chemical space around known scaffolds, and of identifying hits in more diverse datasets.
 While there are clear limitations to the method, we believe that it will open the doors for ultra-high-throughput virtual screening (UHTVS) of compound libraries several orders of magnitude larger than today.
@@
 Given $\mathbf{F}$, the corresponding $\mathbf{C}$ and $\mathbf{E}$ can then be determined either self-consistently (following Pople[@popleElectronInteractionUnsaturated1953]) or via configuration interaction (following Pariser and Parr[@pariserSemiEmpiricalTheory1953a; @pariserSemiEmpiricalTheory1953]), most often starting from a guess $\mathbf{C}$ from the corresponding Hückel model.
-The main effort is then to construct $\mathbf{F}$, which has the following matrix elements[@beveridgeParametrizationSemiempiricalPi1971]
+The main computational effort is then to construct $\mathbf{F}$, which has the following matrix elements[@beveridgeParametrizationSemiempiricalPi1971]
 $$
@@
 
-where $P_{rs}$ are elements of the density matrix $\mathbf{P}$ and $\gamma_{rr}$ and $\gamma_{rs}$ are parameters called the *one-center and two-center electron repulsion integrals*, respectively.
+where $P_{rs}$ are elements of the density matrix $\mathbf{P}$ and $\gamma_{rr}$ and $\gamma_{rs}$ are parameters called the *one-center and two-center electron repulsion integrals* for centers $r$ and $s$, respectively.
 The *core resonance integral* matrix elements of $\mathbf{H}$ are given by
@@
 
-where $r_{rs}$ is the distance between the two atoms $r$ and $s$, and
+where $r_{rs}$ is the distance between the two atom centers $r$ and $s$, and
 $$
@@
 
-The PPP method was implemented in the Python package [coulson]{.smallcaps}[@jornerCoulson] which is freely available on GitHub with a permissive MIT license.
+The PPP method was implemented in the Python package [coulson]{.smallcaps}[@jornerCoulson] which is freely available on GitHub with a permissive open-source MIT license.
 Further details on [coulson]{.smallcaps} will be reported elsewhere.
@@
 For the oscillator strengths, the situation is rather catastrophic with $R^2$ of {{< var local_azulenes_r2_f >}} implying no correlation whatsoever ([@fig-local-azulenes]c).
-The situation for the $\DeltaE{ST}$ is also worse than for the azaphenalenes, with $R^2=$ {{< var local_azulenes_r2_cis >}} with CIS + DSP, a marked improvement over {{< var local_azulenes_r2_scf >}} with SCF + DSP ([@fig-local-azulenes]e,f).
+The situation for the $\DeltaE{ST}$ is also worse than for the azaphenalenes, with $R^2=$ {{< var local_azulenes_r2_cis >}} with CIS + DSP, although a marked improvement over {{< var local_azulenes_r2_scf >}} with SCF + DSP ([@fig-local-azulenes]e,f).
 Unfortunately, the negative gaps are not recovered, leading to recalls of {{< var local_azulenes_recall_cis >}} as none of five inverted molecules could be identified ([@tbl-local-azulenes]).
-It could be speculated that the worse performance for the azaazulenes comes from the fact that they are non-alternants, for which the approximations are expected to be less applicable.
-Considering the contrasting performance for the azaphenalenes and azaazulenes, it seems clear that the PPP screening protocol will inevitably miss some hits belonging to particular compound classes.
+It could be speculated that the worse performance for the azaazulenes comes from the fact that they are non-alternant molecules, for which the approximations are expected to be less applicable.
+Considering the contrasting performance for the azaphenalenes and azaazulenes, it seems clear that the PPP  protocol will miss hits for some compound classes.
 To investigate the global performance of the model we turned to a more diverse set with 16 different compound classes.
@@
 The wide range in $R^2$ and recall further reinforces that the method struggles with some particular compounds classes despite its favorable global performance.
+Also, there seems to be no clear performance difference between alternant and non-alternant molecules, as we speculated above based on the results for azaphenalenes and azaazulenes.
 Despite these problems, the method is able to capture inverted molecules in {{< var global_scaffolds_n_captured >}} out of the 15 scaffolds where they occur.
@@
 The chief limitation of the method is that it only includes the $\pi$-electron system and therefore struggles with: (1) functional groups without a clear $\sigma$--$\pi$ separation (found in common functional groups such as sulfonyls) (2) inductive effects and (3) neglect of $n \rightarrow \pi^*$ transitions.
-Further limitations include (4) lack of solvation model and need for (5) already optimized geometries.
-We believe that at least some of these limitations could be mitigated by a reparametrization specifically targeting inverted gaps, while geometries from the fast GFN-xTB method would be suitable.
+Further limitations of the approach used in this study include (4) lack of solvation model and need for (5) already optimized geometries.
+We believe that at least some of these limitations could be mitigated by a reparametrization specifically targeting inverted gaps, while geometries from the fast GFN-xTB family of methods would be suitable.
 
@@
 
-::: {.callout-important}
-
-Please add acknowledgements.
-
-:::
-
 # Data availability
@@
 Snakemake[@molderSustainableDataAnalysis2021] was used as workflow manager, ensuring reproducible data generation.
-An archive of the workflow run used to generate this manuscript can be found on <https://doi.org/[doi]> together with all the output data generated.
+An archive of the workflow run used to generate this manuscript can be found on <https://doi.org/10.5281/zenodo.8289952> together with all the output data generated.
 
@@
 
-![Singlet and triplet excitation energies as well as oscillator strengths for the rationally designed dataset against the reference level.]({{< var fig_global_all >}}.pdf){#fig-global-all}
+![Singlet excitation energies, triplet excitation energies and oscillator strengths for the rationally designed dataset against the reference level.]({{< var fig_global_all >}}.pdf){#fig-global-all}
 

### 2023-09-22 workflow/notebooks/utils.py

@@ -40,10 +40,10 @@ AXIS_LABELS: dict[str, str] = {
     "homo_lumo_overlap": r"$O_\mathrm{HL}$",
     "oscillator_strength": r"$f_\mathrm{CIS}$",
     "oscillator_strength_ref": r"$f_\mathrm{REF}$",
-    "s1_cis": r"$\Delta E_\mathrm{S,CIS}} (eV)$",
-    "s1_ref": r"$\Delta E_\mathrm{S,REF}} (eV)$",
-    "t1_cis": r"$\Delta E_\mathrm{T,CIS}} (eV)$",
-    "t1_ref": r"$\Delta E_\mathrm{T,REF}} (eV)$",
+    "s1_cis": r"$\Delta E_\mathrm{S,CIS}}$ (eV)",
+    "s1_ref": r"$\Delta E_\mathrm{S,REF}}$ (eV)",
+    "t1_cis": r"$\Delta E_\mathrm{T,CIS}}$ (eV)",
+    "t1_ref": r"$\Delta E_\mathrm{T,REF}}$ (eV)",
     "t1_s1_cc2_aug_dz": r"$\Delta E_{\mathrm{ST,REF}}$ (eV)",
     "t1_s1_cc2_dz": r"$\Delta E_{\mathrm{ST,REF}}$ (eV)",
     "t1_s1_eom_aug_dz": r"$\Delta E_{\mathrm{ST,REF}}$ (eV)",
@@ -112,7 +112,7 @@ ALTERNANT: dict[str, str] = {
 """Classification into alternant/non-alternant for the parent compounds of the rational design set."""
 
 
-def f1_score(y_true: ArrayLike, y_pred: ArrayLike):
+def f1_score(y_true: ArrayLike, y_pred: ArrayLike) -> float:
     """Return F1 score which does not break when tp and fp are zero.
 
     Args:
@@ -125,7 +125,7 @@ def f1_score(y_true: ArrayLike, y_pred: ArrayLike):
     tn, fp, fn, tp = sklearn.metrics.confusion_matrix(
         y_true, y_pred, labels=[False, True]
     ).ravel()
-    f1 = 2 * tp / (2 * tp + fp + fn)
+    f1: float = 2 * tp / (2 * tp + fp + fn)
 
     return f1
 
@@ -140,9 +140,9 @@ def plot_zero_zero(
     zero_zero: bool = True,
     n_dec: int = 2,
     legend_loc: str | None = None,
-    **kwargs,
[…]

### 2023-09-22 manuscript/manuscript.qmd
Fix typo

@@
 
-The code to perform all the calculations and to generate the PDF of the manuscript using the Quarto publishing system are found at <https://https://github.com/kjelljorner/ppp-invest>.
+The code to perform all the calculations and to generate the PDF of the manuscript using the Quarto publishing system are found at <https://github.com/kjelljorner/ppp-invest>.
 Snakemake[@molderSustainableDataAnalysis2021] was used as workflow manager, ensuring reproducible data generation.

### 2023-09-24 manuscript/manuscript.qmd
Fix typo

@@
       - ref: toronto-cs    
-  - name: "Alán Aspuru Guzik"
+  - name: "Alán Aspuru-Guzik"
     orcid: 0000-0002-8277-4434

### 2024-01-25 manuscript/manuscript.qmd
Revisions

@@
     department: "Canadian Institute for Advanced Research (CIFAR)"
+    address: "661 University Ave."
     postal-code: "M5G 1M1"
@@
     department: "University of Toronto"
+    address: 700 University Ave.
+    postal-code: "M5G 1Z5"
     city: "Toronto"
@@
 
-title: "Ultrafast computational screening of molecules with inverted singlet-triplet energy gaps using the Pariser-Parr-Pople semi-empirical quantum chemistry method"
+title: "Ultrafast Computational Screening of Molecules with Inverted Singlet-triplet Energy Gaps Using the Pariser-Parr-Pople Semi-empirical Quantum Chemistry Method"
 abstract: |
@@
 High-throughput virtual screening could assist in finding novel classes of these molecules, but current efforts are hampered by the high computational cost of the required quantum-chemical methods.
-We present a method based on the semi-empirical Pariser-Parr-Pople theory augmented by perturbation theory and show that it reproduces inverted gaps at a fraction of the cost of currently employed excited state calculations..
+We present a method based on the semi-empirical Pariser-Parr-Pople theory augmented by perturbation theory and show that it reproduces inverted gaps at a fraction of the cost of currently employed excited state calculations.
 Our study paves the way for ultra-high-throughput virtual screening and inverse design to accelerate the discovery and development of this new generation of OLED materials.
+
 bibliography: [references.bib] 
@@
       name: "Journal of Physical Chemistry A"
-      model: "1p"    
+      model: "3p"    
     number-sections: false
@@
       - file: macros.tex    
+      - file: packages.tex
 csl: "acs.csl"   
@@
 tab_garner_gaps = ""
+tab_parameters = ""
+tab_opt = ""
 ```
@@
 
-Molecules with an INVErted Singlet-Triplet energy gap (INVEST) are exceedingly rare.[@liOrganicMoleculesInverted2022] They are said to violate Hund's rule of maximum multiplicity as applied to the S~1~ and T~1~ states, which describes that the electronic state with the highest spin (in this case the triplet) should be lowest in energy.
-While a handful of examples of molecules violating Hund's rule in the excited state have been known since the 1970s and 1980s,[@kollmarViolationHundRule1978; @leupinLowlyingElectronicallyExcited1980; @kosekiViolationHundMultiplicity1985;@toyotaViolationHundRule1988;@toyotaViolationHundMultiplicity1986] it is only recently that they have received considerable attention for application in OLEDs.[@desilvaInvertedSingletTriplet2019;@polliceOrganicMoleculesInverted2021;@aizawaDelayedFluorescenceInverted2022] As they are rare, recent efforts have used high-throughput virtual screening (HTVS) with computational chemistry to identify new compounds with inverted gaps, focusing on expanding the local chemical space around specific scaffolds[@polliceOrganicMoleculesInverted2021;@sanz-rodrigoNegativeSingletTriplet2021;@ricciSingletTripletExcited2021;@sobolewskiAreHeptazineBasedOrganic2021;@aizawaDelayedFluorescenceInverted2022;@ricciEstablishingDesignStrategies2022;@tuckovaOriginInvertedSinglet2022] or scanning larger (combinatorially or experimentally derived) datasets to identify novel scaffolds.[@terenceblaskovitsSymmetryInducedSinglet2023;@omarIdentificationUnknownInverted2023;@garnerDoublebondDelocalizationNonalternant2023] These HTVS efforts have been successful at identifying several new INVEST candidates, some of which have also been synthesized and tested experimentally.[@aizawaDelayedFluorescenceInverted2022]
+Molecules with an INVErted Singlet-Triplet energy gap (INVEST) are exceedingly rare.[@liOrganicMoleculesInverted2022] They \deleted{are said to} violate Hund's rule of maximum multiplicity as applied to the S~1~ and T~1~ states \added{of molecules}, which describes that the electronic state with the highest spin (in this case the triplet) should be lowest in energy.
+While a handful of examples of molecules violating Hund's rule in the excited state have been known since the 1970s and 1980s,[@kollmarViolationHundRule1978; @leupinLowlyingElectronicallyExcited1980; @kosekiViolationHundMultiplicity1985;@toyotaViolationHundRule1988;@toyotaViolationHundMultiplicity1986] it is only recently that they have received considerable attention for application in OLEDs.[@desilvaInvertedSingletTriplet2019;@polliceOrganicMoleculesInverted2021;@aizawaDelayedFluorescenceInverted2022] As they are rare, recent efforts have used high-throughput virtual screening (HTVS) with computational chemistry to identify new compounds with inverted gaps, focusing on expanding the local chemical space around specific scaffolds[@polliceOrganicMoleculesInverted2021;@sanz-rodrigoNegativeSingletTriplet2021;@ricciSingletTripletExcited2021;@sobolewskiAreHeptazineBasedOrganic2021;@aizawaDelayedFluorescenceInverted2022;@ricciEstablishingDesignStrategies2022;@tuckovaOriginInvertedSinglet2022] or scanning larger (combinatorially or experimentally derived) datasets to identify novel scaffolds.[@terenceblaskovitsSymmetryInducedSinglet2023;@omarIdentificationUnknownInverted2023a;@garnerDoublebondDelocalizationNonalternant2023a] These HTVS efforts have been successful at identifying several new INVEST candidates, some of which have also been synthesized and tested experimentally.[@aizawaDelayedFluorescenceInverted2022]
 
 While this early success of HTVS is highly encouraging, it is hampered by the cost of the computational methods.
-Excitation energies are routinely calculated by time-dependent density functional theory (TD-DFT), which has become a workhorse of computational photophysics over the last decades.[@casidaProgressTimeDependentDensityFunctional2012] Unfortunately, it has been shown that TD-DFT fails to capture the inverted gap of INVEST compounds, as the method only considers single electronic excitations.[@desilvaInvertedSingletTriplet2019] Inclusion of at least double excitations are necessary to reproduce the inverted gaps, corresponding to methods such as double hybrid TD-DFT or excited state coupled-cluster methods such as second-order approximate coupled cluster singles and doubles (CC2) or equation of motion coupled cluster with single and double excitations (EOM-CCSD).
+Excitation energies are routinely calculated by time-dependent density functional theory (TD-DFT), which has become a workhorse of computational photophysics over the last decades.[@casidaProgressTimeDependentDensityFunctional2012] Unfortunately, it has been shown that TD-DFT fails to capture the inverted gap of INVEST compounds, as the method only considers single electronic excitations.[@desilvaInvertedSingletTriplet2019] Inclusion of at least double excitations are necessary to reproduce the inverted gaps, corresponding to methods such as double hybrid TD-DFT \added{(via perturbation theory)}[@grimmeDoublehybridDensityFunctional2007] or excited state coupled-cluster methods such as second-order approximate coupled cluster singles and doubles (CC2) or equation of motion coupled cluster with single and double excitations (EOM-CCSD).
 In line with the results of benchmarking studies,[@ricciSingletTripletExcited2021;@sancho-garciaViolationHundRule2022] recent HTVS studies have used methods such as the complete active space self-consistent field (CASSCF) and CC2 for preliminary screening, while confirming inverted gaps with more expensive methods such as multistate complete active space second-order perturbation theory (MS-CASPT2) and EOM-CCSD.
-These methods are not only expensive, but in the case of the CAS methods, also require the choice of an orbital active space, something which is not routinely automatized.
+These methods are not only expensive \added{(compared to TD-DFT)}, but in the case of the CAS methods, also require the choice of an orbital active space, something which is not routinely automatized \added{(although advances have been made)}.[@steinAutomatedSelectionActive2016;@sayfutyarovaAutomatedConstructionMolecular2017]
 
@@
 The Pariser-Parr-Pople (PPP) method is an extension of the semi-empirical Hückel molecular orbital theory for $\pi$-electron systems that includes electron correlation.[@pariserSemiEmpiricalTheory1953a;@pariserSemiEmpiricalTheory1953; @popleElectronInteractionUnsaturated1953] It originated in the 1950s, with new parametrizations being developed mainly in the 1960s and 1970s, and was used in the dye industry at least until the 1980s before the advent of more accurate ab initio methods.[@griffithsPracticalAspectsColour1982] The PPP method retains the conceptual clarity of Hückel theory, while at the same time including the electron correlation that is necessary to capture the inverted gap.
+\added{For another recent application of PPP to INVEST compounds, see the work by Painelli and co-workers.}[@bedogniShiningLightInverted2023]
 
@@
 according to recursion formulas given by Mulliken.[@mullikenFormulasNumericalTables1949] Thus, the four parameters $U_{r}$, $\gamma_{rr}$, $\gamma_{rs}$ and $\beta_{rs}$ are expressed completely in terms of the corresponding valence state ionization potentials $\mathrm{IP}_v$ and electron affinities $\mathrm{EA}_v$ as well as the completely empirical parameter $C$.
+\added{We have used the set of valence state ionization potentials and electron affinities from Beveridge and Hinze} ([@tbl-parameters]).[@beveridgeParametrizationSemiempiricalPi1971]
 
@@
 \beta_{rs}^\prime = \beta_{rs} \cos{\theta}
-$${#eq-beta-rs}
+$${#eq-beta-rs-prime}
 
@@
 Singlet-triplet inversion occurs also for the ground state, where several violations of Hund’s rule are well known, for example for bond-equalized cyclobutadiene at the *D*~4h~ geometry.
-Borden and Davidson explained this effect in terms of dynamic spin polarization (DSP),[@bordenTheoreticalStudiesDiradicals1981] in which the electrons of a pair of disjoint non-bonding molecular orbitals (NBMOs, [@fig-dsp]a) experience stronger electron correlation effects in the singlet state as compared to the triplet state ([@fig-dsp]b).
-For a pedagogic introduction to the topic, the reader is referred to an excellent article by Karafiloglou.[@karafiloglouDoubleDynamicSpin1989] As outlined by Kollmar and Staemmler,[@kollmarViolationHundRule1978] as well as Malrieu and co-workers,[@benamorSpinPolarizationElectronic2020] an alternative view of the same phenomenon is through admixture of singly-excited configurations into the electronic wavefunction of the ground state ([@fig-dsp]c).
-The DSP effect for the S~1~ excited state can correspondingly be described by admixture of doubly excited configurations,[@kollmarViolationHundRule1978] as they are singly excited with respect to the S~1~ state, and doubly excited with respect to the S~0~ ground state ([@fig-dsp]d).
+Borden and Davidson explained this effect in terms of dynamic spin polarization (DSP),[@bordenTheoreticalStudiesDiradicals1981] in which the electrons of a pair of disjoint non-bonding molecular orbitals (NBMOs, [@fig-dsp]a) experience stronger electron correlation effects in the singlet state as compared to the triplet state ([@fig-dsp]b).[@albrightOrbitalInteractionsChemistry2013] For a pedagogic introduction to the topic, the reader is referred to an excellent article by Karafiloglou.[@karafiloglouDoubleDynamicSpin1989] As outlined by Kollmar and Staemmler,[@kollmarViolationHundRule1978] as well as Malrieu and co-workers,[@benamorSpinPolarizationElectronic2020; @davidImprovedEvaluationSpinpolarization2020] an alternative view of \replaced{spin polarization}{the same phenomenon} is through admixture of \deleted{singly-}excited configurations into the electronic wavefunction of the ground state ([@fig-dsp]c).[@chipmanSpinPolarizationModel1992;@engelsRecentDevelopmentsConfiguration1996;@coulaudAnalysisPhysicalContributions2012] \added{This is shown in }[@fig-dsp]\added{c for the case of static spin polarization in the allyl radical.} The DSP effect for the S~1~ excited state can correspondingly be described by admixture of doubly excited configurations,[@kollmarViolationHundRule1978] \deleted{as they are} singly excited with respect to the S~1~ state, and doubly excited with respect to the S~0~ ground state ([@fig-dsp]d).
 
-As first shown by Kollmar and Staemmler,[@kollmarViolationHundRule1978] and recently re-emphasized by Pernal and co-workers,[@drwalRoleSpinPolarization2023] the $\DeltaE{ST}$ can be approximated by a combination of two terms, the first one being the exchange interaction $2K$ (given by @eq-exchange) and a correction term $\DeltaE[DSP]{}$ due to the DSP, which can be approximated with perturbation theory:[@kollmarViolationHundRule1978]
+As first shown by Kollmar and Staemmler,[@kollmarViolationHundRule1978] and recently re-emphasized by Pernal and co-workers,[@drwalRoleSpinPolarization2023a] the $\DeltaE{ST}$ can be approximated by a combination of two terms, the first one being the exchange interaction $2K$ (given by @eq-exchange) and a correction term $\DeltaE[DSP]{}$ due to the DSP, which can be approximated with perturbation theory:[@kollmarViolationHundRule1978]
 
@@
 (d) In the same way as cyclobutadiene is stabilized by DSP in the singlet ground state, molecules with inverted singlet-triplet gaps in the excited state are also stabilized by DSP that can be described by admixture of excited configurations.
-For singly excited states, this means addition of doubly excited states.](../data/figures/dynamic_spin_polarization.pdf){#fig-dsp}
+For singly excited states, this means addition of doubly excited states.
+Panel b has been adapted from ref.
+[@karafiloglouDoubleDynamicSpin1989].](../data/figures/dynamic_spin_polarization.pdf){#fig-dsp}
 
@@
 The PPP wavefunction was converged with the self-consistent field method using the naive variational principle.
-To speed up the calculations, we derived a spline interpolation of the overlap integrals,[@mullikenFormulasNumericalTables1949] which will be reported separately.
+\added{CIS calculations used the converged SCF wavefunction as reference and modelled vertical excitations based on the pre-computed geometries.} To \replaced{make}{speed up} the calculations \added{more robust}, we derived a spline interpolation of the overlap integrals,[@mullikenFormulasNumericalTables1949] which will be reported separately.
+\added{Evaluated on the azaphenalene dataset (vide infra), it shows good accuracy with $R^2=$} {{< var interpolation_all_r2_cis >}} \added{and RMSE =} {{< var interpolation_all_rmse_cis >}} eV \added{for $\DeltaEST[DSP]{CIS}$} ([@fig-interpolation]).
 All calculations have been performed on a MacBook Air laptop computer with an M2 processor.
 We further integrated [coulson]{.smallcaps} with [PySCF]{.smallcaps}[@sunRecentDevelopmentsPySCF2020] to provide alternative algorithms for the SCF convergence and CIS and found the results consistent within numerical accuracy.
-Oscillator strengths were calculated with the dipole length approximation.[@jorgensenTimedependentHartreeFockCalculations1970;@klessingerExcitedStatesPhotochemistry1995;@damourGroundExcitedStateDipole2023] Some compounds exhibit negative triplet excitation energies and in some case even negative singlet excitation energies, which indicates a restricted-unrestricted instability of the ground state wavefunction.
+\added{To give an indication of the computational cost, the SCF calculation for azaphenalene (13 heavy atoms, 14 electrons), takes} {{< var timing_mean_scf_only >}}$\pm${{< var timing_std_scf_only >}} \added{ms (average and standard deviation over 100 runs, respectively).
+On top of this, DSP takes} {{< var timing_mean_part_scf >}}$\pm${{< var timing_std_part_scf >}} \added{ms, and CIS + DSP takes} {{< var timing_mean_part_cis >}}$\pm${{< var timing_std_part_cis >}} \added{ms.} Oscillator strengths were calculated with the dipole length approximation.[@jorgensenTimedependentHartreeFockCalculations1970;@klessingerExcitedStatesPhotochemistry1995;@damourGroundExcitedStateDipole2023] Some compounds exhibit negative triplet excitation energies and in some case even negative singlet excitation energies, which indicates a restricted-unrestricted instability of the ground state wavefunction.
 We have here taken the pragmatic approach to ignore these issues, while we will highlight below some examples where it occurs and if it has any effect on the overall results.
@@
 We used four separate datasets for this study.
-The first is a set of azaphenalenes previously studied by some of us, comprising 256 substituted compounds, for which excitation energies were computed with (2nd-order) algebraic-diagrammatic construction, {{< var reference_level >}}/cc-pVDZ, and geometries were optimized with B3LYP/cc-pVDZ[@polliceOrganicMoleculesInverted2021] The second is a set substituted 138 azaazulenes with {{< var reference_level >}}/cc-pVDZ excitation energies and optimized with the B97-3c composite method.[@nigamArtificialDesignOrganic2023] The third is a set of 16 rationally designed scaffolds, which including substitutions amounts to {{< var global_n_total >}} unique compounds, optimized at the B97-3c level and with excitation energies at the {{< var reference_level >}}/cc-pVDZ level.[@polliceRationalDesignOrganic2023] The fourth is a set 315 non-alternant hydrocarbons including substitutions, divided into three subsets of size 76, 187 and 52, optimized at the $\omega$B97X-D/def2-TZVP level and with excitation energies at the CC2/aug-cc-pVTZ level.
+The first is a set of azaphenalenes previously studied by some of us, comprising 256 substituted compounds, for which excitation energies were computed with (2nd-order) algebraic-diagrammatic construction, {{< var reference_level >}}/cc-pVDZ, and geometries were optimized with B3LYP/cc-pVDZ\added{.}[@polliceOrganicMoleculesInverted2021] The second is a set \added{of 138} substituted \deleted{138} azaazulenes with {{< var reference_level >}}/cc-pVDZ excitation energies and optimized with the B97-3c composite method.[@nigamArtificialDesignOrganic2024] The third is a set of 16 rationally designed scaffolds, which including substitutions amounts to {{< var global_n_total >}} unique compounds, optimized at the B97-3c level and with excitation energies at the {{< var reference_level >}}/cc-pVDZ level.[@polliceRationalDesignOrganic2023] The fourth is a set 315 non-alternant hydrocarbons including substitutions, divided into three subsets of size 76, 187 and 52, optimized at the $\omega$B97X-D/def2-TZVP level and with excitation energies at the CC2/aug-cc-pVTZ level.
+\added{Below, we compare our PPP-based method to the reference levels at the DFT-optimized geometries given in the datasets.
+For the azaphenalene dataset, we also investigated the impact of geometry.
+An initial geometry of each molecule was generated with the EmbedMolecule function in RDKit,}[@RDKitOpensourceCheminformatics]\added{ and then optimized with MMFF94.}[@halgrenMerckMolecularForce1996]\added{ This geometry was then further refined with either the GFN-FF force field,}[@spicherRobustAtomisticModeling2020]\added{ the GFN2-xTB semi-empirical method,}[@bannwarthGFN2xTBAccurateBroadly2019] \added{or the ANI-1ccx machine learning potential}[@smithApproachingCoupledCluster2019]\added{ (as implemented in TorchANI}[@gaoTorchANIFreeOpen2020]\added{ and using ASE as the optimization backend}[@hjorthlarsenAtomicSimulationEnvironment2017]\added{).}
 
@@
 We first demonstrate that our method is capable of capturing the inverted $\DeltaE{ST}$ for some model compounds.
-As shown by Toyota and co-workers, pentalene at the ideal  $D_\mathrm{2h}$ geometry has an inverted gap, while relaxation to the minimum with $C_\mathrm{2h}$ symmetry leads to a normal gap.[@kosekiViolationHundMultiplicity1985] In our calculations, pentalene with equal bond lengths of 1.4 Å displays a small HOMO-LUMO overlap of {{< var pentalene_ideal_overlap >}}, leading to a small exchange interaction of only $2K=$ {{< var pentalene_ideal_2_exchange >}} kcal/mol (@fig-pentalene-ideal-orbitals).
-The dynamic spin polarization correction $\DeltaE[DSP]{SCF}$ of {{< var pentalene_ideal_dsp_scf >}} kcal/mol leads to a net predicted $\DeltaEST[DSP]{SCF}$ of {{< var pentalene_ideal_t1_s1_dsp_scf >}} kcal/mol.
-Adding additional correlation with CIS leads to a $\DeltaEST[DSP]{CIS}$ of {{< var pentalene_ideal_t1_s1_dsp_cis >}} kcal/mol.
+As shown by Toyota and co-workers, pentalene at the ideal  $D_\mathrm{2h}$ geometry has an inverted gap, while relaxation to the minimum with $C_\mathrm{2h}$ symmetry \added{(using B97-3c)} leads to a normal gap.[@kosekiViolationHundMultiplicity1985] In our calculations, pentalene with equal bond lengths of 1.4 Å displays a small HOMO-LUMO overlap of {{< var pentalene_ideal_overlap >}}, leading to a small exchange interaction of only $2K=$ {{< var pentalene_ideal_2_exchange >}} eV (@fig-pentalene-ideal-orbitals).
+The dynamic spin polarization correction $\DeltaE[DSP]{SCF}$ of {{< var pentalene_ideal_dsp_scf >}} eV leads to a net predicted $\DeltaEST[DSP]{SCF}$ of {{< var pentalene_ideal_t1_s1_dsp_scf >}} eV.
+Adding additional correlation with CIS leads to a $\DeltaEST[DSP]{CIS}$ of {{< var pentalene_ideal_t1_s1_dsp_cis >}} eV.
 The simple perturbation theory model allows us to gain additional insight into the contributions to $\DeltaE[DSP]{SCF}$ as they correspond to single excitations from doubly occupied orbitals below the HOMO to unoccupied orbitals above the LUMO (@fig-pentalene-ideal-excitations).
-Three excitations, HOMO-3 → LUMO+2, HOMO-2 → LUMO+1 and HOMO-1 → LUMO+3 are identified as the main contributors to the spin polarization effect, and could potentially be tuned by substituent effects.
+Three excitations, HOMO-3 → LUMO+2, HOMO-2 → LUMO+1 and HOMO-1 → LUMO+3 \added{(1 → 7, 2 → 6 and 3 → 8 in }[@fig-pentalene-ideal-excitations]\added{)} are identified as the main contributors to the spin polarization effect, and could potentially be tuned by substituent effects.
 
-![Excitations and their contribution to $\DeltaE[DSP]{SCF}=$ {{< var pentalene_ideal_dsp_scf >}} kcal/mol for pentalene at the $D_\mathrm{2h}$ geometry and bond lengths of 1.4 Å.]({{< var fig_pentalene_ideal_excitations >}}.png){#fig-pentalene-ideal-excitations}
+![Excitations and their contribution to $\DeltaE[DSP]{SCF}=$ {{< var pentalene_ideal_dsp_scf >}} eV for pentalene at the $D_\mathrm{2h}$ geometry and bond lengths of 1.4 Å.]({{< var fig_pentalene_ideal_excitations >}}.png){#fig-pentalene-ideal-excitations}
 
-Using the relaxed $C_\mathrm{2h}$ geometry leads to dramatic changes in the frontier orbitals, with an increased HOMO-LUMO overlap of {{< var pentalene_overlap >}} and a sizeable exchange interaction of {{< var pentalene_2_exchange >}} kcal/mol (@fig-pentalene-orbitals).
-Consequently, the calculated gap is now normal at $\DeltaEST[DSP]{SCF}=$ {{< var pentalene_t1_s1_dsp_scf >}} kcal/mol, aggravated by a diminished $\DeltaE[DSP]{SCF}$ of only {{< var pentalene_dsp_scf >}} kcal/mol.
-A similar result is obtained with CIS: $\DeltaEST[DSP]{CIS}=$ {{< var pentalene_t1_s1_dsp_cis >}} kcal/mol.
-The reference value calculated with {{< var reference_level >}} is {{< var pentalene_t1_s1_ref >}} kcal/mol.
-The excitations contributing to $\DeltaE[DSP]{SCF}$ at the $D_\mathrm{2h}$ geometry have been significantly diminished, only partially alleviated by the addition of a minor stabilization from HOMO-2 → LUMO+3 (@fig-pentalene-excitations).
+Using the relaxed $C_\mathrm{2h}$ geometry leads to dramatic changes in the frontier orbitals, with an increased HOMO-LUMO overlap of {{< var pentalene_overlap >}} and a sizeable exchange interaction of {{< var pentalene_2_exchange >}} eV (@fig-pentalene-orbitals).
+Consequently, the calculated gap is now normal at $\DeltaEST[DSP]{SCF}=$ {{< var pentalene_t1_s1_dsp_scf >}} eV, aggravated by a diminished $\DeltaE[DSP]{SCF}$ of only {{< var pentalene_dsp_scf >}} eV.
+A similar result is obtained with CIS: $\DeltaEST[DSP]{CIS}=$ {{< var pentalene_t1_s1_dsp_cis >}} eV.
+The reference value calculated with {{< var reference_level >}} is {{< var pentalene_t1_s1_ref >}} eV.
+The excitations contributing to $\DeltaE[DSP]{SCF}$ at the $D_\mathrm{2h}$ geometry have been significantly diminished, only partially alleviated by the addition of a minor stabilization from HOMO-2 → LUMO+3 (\added{2 → 8 in }[@fig-pentalene-excitations]).
 
-![Excitations and their contribution to $\DeltaE[DSP]{SCF}=$ {{< var pentalene_dsp_scf >}} kcal/mol for pentalene at the optimized $C_\mathrm{2h}$ geometry.]({{< var fig_pentalene_excitations >}}.png){#fig-pentalene-excitations}
+![Excitations and their contribution to $\DeltaE[DSP]{SCF}=$ {{< var pentalene_dsp_scf >}} eV for pentalene at the optimized $C_\mathrm{2h}$ geometry (optimized with B97-3c).]({{< var fig_pentalene_excitations >}}.png){#fig-pentalene-excitations}
 
 Azaphenalene is the prototypical molecule that started the renewed investigations into INVEST molecules,[@desilvaInvertedSingletTriplet2019;@sobolewskiAreHeptazineBasedOrganic2021] and has been the subject of numerous studies with high-level quantum chemical methods.
-Can we capture the inverted gap, as first measured by Leupin and Wirz in 1980?[@leupinLowlyingElectronicallyExcited1980] The calculations reveal that the HOMO and LUMO are well separated with spatial overlap of {{< var azaphenalene_overlap >}} and a correspondingly low exchange interaction of only {{< var azaphenalene_2_exchange >}} kcal/mol ([@fig-azaphenalene-orbitals]).
-The larger azaphenalene has 30 excitations that could possibly contribute to the large $\DeltaE[DSP]{SCF}$ of {{< var azaphenalene_dsp_scf >}} kcal/mol, but the most important by far are those from the doubly degenerate HOMO-1 to the doubly degenerate LUMO+1 ([@fig-azaphenalene-excitations]).
-In total, $\DeltaEST[DSP]{SCF}=$ {{< var azaphenalene_t1_s1_dsp_scf >}} kcal/mol, which is slightly smaller at the CIS level with $\DeltaEST[DSP]{CIS}=$ {{< var azaphenalene_t1_s1_dsp_cis >}} kcal/mol.
+Can we capture the inverted gap, as first measured by Leupin and Wirz in 1980?[@leupinLowlyingElectronicallyExcited1980] The calculations reveal that the HOMO and LUMO are well separated with spatial overlap of {{< var azaphenalene_overlap >}} and a correspondingly low exchange interaction of only {{< var azaphenalene_2_exchange >}} eV ([@fig-azaphenalene-orbitals]).
+The larger azaphenalene has 30 excitations that could possibly contribute to the large $\DeltaE[DSP]{SCF}$ of {{< var azaphenalene_dsp_scf >}} eV, but the most important by far are those from the doubly degenerate HOMO-1 to the doubly degenerate LUMO+1 (\added{5 → 10 and 6 → 9 in }[@fig-azaphenalene-excitations]).
+In total, $\DeltaEST[DSP]{SCF}=$ {{< var azaphenalene_t1_s1_dsp_scf >}} eV, which is slightly smaller at the CIS level with $\DeltaEST[DSP]{CIS}=$ {{< var azaphenalene_t1_s1_dsp_cis >}} eV.
 
@@
 
-![Excitations and their contribution to $\DeltaE[DSP]{SCF}=$ {{< var azaphenalene_dsp_scf >}} kcal/mol for azaphenalene]({{< var fig_azaphenalene_excitations >}}.png){#fig-azaphenalene-excitations}
+![Excitations and their contribution to $\DeltaE[DSP]{SCF}=$ {{< var azaphenalene_dsp_scf >}} eV for azaphenalene]({{< var fig_azaphenalene_excitations >}}.png){#fig-azaphenalene-excitations height=95%}
 
@@
 This $R^2$ value is essentially unchanged from the gap at the CIS level without DSP ({{< var local_azaphenalenes_r2_cis_raw >}}, [@fig-local-azaphenalenes]d) and markedly better than for the gap at the SCF level ({{< var local_azaphenalenes_r2_scf >}}, [@fig-local-azaphenalenes]e), showing the importance of going beyond the SCF level to include at least some configuration interaction.
-The results are of similar or better quality recently achieved by Pernal and co-workers for a different set of azaphenalenes using SCF + DSP with orbitals from DFT.[@drwalRoleSpinPolarization2023] To further analyze the results in terms of a binary classification into normal/inverted, we calculated the true positives (TP), true negatives (TN), false positives (FP) and false negatives (FN), as well as a range of common classification scores (@tbl-local-azaphenalenes).
+The results are of similar or better quality recently achieved by Pernal and co-workers for a different set of azaphenalenes using SCF + DSP with orbitals from DFT.[@drwalRoleSpinPolarization2023a] To further analyze the results in terms of a binary classification into normal/inverted, we calculated the true positives (TP), true negatives (TN), false positives (FP) and false negatives (FN), as well as a range of common classification scores (@tbl-local-azaphenalenes).
 For definitions, see [@eq-f1] to [@eq-roc-auc].
@@
 
+\small
 ```{python}
@@
 ```
+\normalsize
 
@@
 This corresponds to a success rate of {{< var global_percent_success >}}%.
-To put this into context, ~25% of the CASSCF calculations and 9% of the CIS calculations failed in a recent HTVS workflow by Troisi, Padula and co-workers.[@omarIdentificationUnknownInverted2023] The total runtime using four cores on the M2 processor was {{< var time_cpu_hours >}} CPU hours, with a mean time per compounds of {{< var time_mean >}} seconds.
+To put this into context, ~25% of the CASSCF calculations and 9% of the CIS calculations failed in a recent HTVS workflow by Troisi, Padula and co-workers.[@omarIdentificationUnknownInverted2023a] The total runtime using four cores on the M2 processor was {{< var time_cpu_hours >}} CPU hours, with a mean time per compounds of {{< var time_mean >}} seconds.
 The S~1~ and T~1~ excitation energies are fairly reproduced with ([@fig-global-all]a,b) $R^2$ of {{< var global_all_r2_singlet >}} and {{< var global_all_r2_triplet >}} against {{< var reference_level >}}, respectively.
@@
 
-A more detailed breakdown of the correlations into the different scaffolds shows that the highest $R^2$ of {{< var global_scaffolds_r2_max >}} is obtained for {{< var global_scaffolds_r2_max_name >}}, while the lowest is obtained for {{< var global_scaffolds_r2_min_name >}} with {{< var global_scaffolds_r2_min >}}.
+A more detailed breakdown of the correlations into the different scaffolds ([@fig-global-scaffolds-gaps]) shows that the highest $R^2$ of {{< var global_scaffolds_r2_max >}} is obtained for {{< var global_scaffolds_r2_max_name >}}, while the lowest is obtained for {{< var global_scaffolds_r2_min_name >}} with {{< var global_scaffolds_r2_min >}}.
 For recall, the corresponding range is {{< var global_scaffolds_recall_max >}} for {{< var global_scaffolds_recall_max_name >}} to {{< var global_scaffolds_recall_min >}} for {{< var global_scaffolds_recall_min_name >}}.
@@
 To further test the validity of the method, we calculated a series of compounds recently published by Garner _et.
-al_.[@garnerDoublebondDelocalizationNonalternant2023] The dataset comes in three parts: (1) non-alternants, (2) non-alternants with constrained high-symmetry geometries (we here follow the original terminology and call these "avoided symmetry") and (3) substituted non-alternants (here called "substituted").
+al_.[@garnerDoublebondDelocalizationNonalternant2023a] The dataset comes in three parts: (1) non-alternants, (2) non-alternants with constrained high-symmetry geometries (we here follow the original terminology and call these "avoided symmetry") and (3) substituted non-alternants (here called "substituted").
 Out of the substituted compounds, four could not be calculated as they contain four-coordinate P atoms for which our PPP model lacks parameters.
@@
 We also indicate in the plots which compounds show instabilities ([@fig-garner-gaps]).
-There seems to be no clear deterioration in the performance, although also the reference data might exhibit instabilities as there is no mention of any stability analysis in the original manuscript.[@garnerDoublebondDelocalizationNonalternant2023]
+There seems to be no clear deterioration in the performance, although also the reference data might exhibit instabilities as there is no mention of any stability analysis in the original manuscript.[@garnerDoublebondDelocalizationNonalternant2023a]
 
@@
 ```
+\normalsize
+
+## Effect of geometry
+
+\added{Even though our PPP protocol is very fast, this speed would not be beneficial if DFT-optimized geometries were required to accurately reproduce the energy gaps.
+In the benchmarking above, we used the same geometries as in the original datasets to allow for a comparison based on equal footing.
+For ultrafast screening, we would need geometries from force fields, semi-empirical methods or machine learning potentials that can be obtained on a similar time scale as the PPP results.
+We chose the azaphenalene dataset for a limited benchmark, and optimized the geometries with the MMFF94 (average runtime of }{{< var opt_time_mmff >}}\added{ ms) and GFN-FF (}{{< var opt_time_gfnff >}}\added{ ms) force field methods, the GFN2-xTB semi-empirical method (}{{< var opt_time_gfn2 >}}\added{ ms) and the ANI-1ccx machine learning potential (}{{< var opt_time_ani1ccx >}}\added{ ms, CPU-based).
+While the two force fields methods are insufficiently accurate, both GFN2-xTB and ANI-1ccx provide sufficiently good geometries for screening (}[@fig-opt-corr]\added{, }[@fig-opt-gap]\added{ and }[@tbl-opt]\added{).
+In particular, GFN2-xTB shows good correlation ($R^2=$ }{{< var opt_corr_r2_gfn2 >}}\added{) with the gaps at the DFT-optimized geometries.
+Compared to the reference }{{< var reference_level >}}\added{/cc-pVDZ level, the results with GFN2-xTB geometries are equally good to those from DFT geometries ($R^2$ of }{{< var opt_gap_r2_gfn2 >}}\added{ vs.
+}{{< var opt_gap_r2_dft >}}\added{ and recall of }{{< var opt_gap_recall_gfn2 >}}\added{ vs.
+}{{< var opt_gap_recall_dft >}}\added{, respectively).
+The optimization runtime of }{{< var opt_time_gfn2 >}}\added{ ms is slower, but of a similar magnitude compared to the runtime of the PPP protocol itself (}{{< var run_time_gfn2 >}}\added{ ms), and significantly faster than DFT optimization.
+Based on the good performance, it is likely that the GFN2-xTB geometries could be used in virtual screening campaigns.}
 
@@
 Further limitations of the approach used in this study include (4) lack of solvation model and need for (5) already optimized geometries.
-We believe that at least some of these limitations could be mitigated by a reparametrization specifically targeting inverted gaps, while geometries from the fast GFN-xTB family of methods would be suitable.
+We believe that at least some of these limitations could be mitigated by a reparametrization specifically targeting inverted gaps, while \added{our preliminary tests indicate that} geometries from the fast GFN-xTB family of methods would be suitable.
+\added{Alternatively, these limitations could be overcome by applying the perturbation theory description of dynamic spin polarization}[@kollmarViolationHundRule1978]\added{ to the more costly but potentially more accurate all-electron semi-empirical methods, such as those of the OMx family.}[@dralSemiempiricalQuantumChemicalOrthogonalizationCorrected2016]
 
@@
 The method has great potential to accelerate the discovery of the next generation of OLED materials based on INVEST.
+
+# Supporting information
+
+\added{Definition of classification metrics, semi-empirical parameters, additional figures and tables.}
 
@@
 Snakemake[@molderSustainableDataAnalysis2021] was used as workflow manager, ensuring reproducible data generation.
-An archive of the workflow run used to generate this manuscript can be found on <https://doi.org/10.5281/zenodo.8289952> together with all the output data generated.
+An archive of the workflow run used to generate this manuscript can be found on <https://doi.org/10.5281/zenodo.10569815> together with all the output data generated.
+
+{{< pagebreak >}}
+
+# TOC graphic
+![TOC graphic.](../data/figures/TOC.pdf){#fig-toc}
+
+{{< pagebreak >}}
 
@@
 
-\newpage
+{{< pagebreak >}}
+
 \appendix
@@
 
-For definitions, see the Wikipedia article on Precision and Recall.[@PrecisionRecall2023] The F1 score is the harmonic mean of the precision and sensitivity, and is defined as
+For definitions, \replaced{see the review by Tharwat.}{see the Wikipedia article on Precision and Recall.}^S1^ The F1 score is the harmonic mean of the precision and sensitivity, and is defined as
 
@@
 The recall, or true positive rate (TPR), is defined as
+
 $$
@@
 The fallout, or false positive rate (FPR), is defined as
+
 $$
 \frac{\mathrm{FP}}{\mathrm{FP} + \mathrm{TN}}
-$$ {#eq-recall}
+$$ {#eq-fpr}
 
@@
 
+\small
+```{python}
+# | label: tbl-parameters
+# | tbl-cap: Valence state ionization potentials and electron affinities for the elements in the dataset.
+For elements that can contribute more than one $\pi$-electron, X1 is the element X with one $\pi$-electron, X2 with two $\pi$-electrons.
+# | echo: false
+import pandas as pd
+from IPython.display import display, Markdown
+
+with open(tab_parameters) as f:
+    md = f.read()
+display(Markdown(md))
+```
+\normalsize
+
+![Comparison of singlet-triplet energy gaps using non-interpolated and interpolated overlap integrals  for (a) SCF + DSP (b) CIS + DSP.]({{< var fig_interpolation >}}.pdf){#fig-interpolation}
+
 ![Orbital energies and frontier orbitals of pentalene at the idealized $D_\mathrm{2h}$ geometry and bond lengths of 1.4 Å.]({{< var fig_pentalene_ideal_orbitals >}}.png){#fig-pentalene-ideal-orbitals}
 
-![Orbital energies and frontier orbitals of pentalene at the optimized $C_\mathrm{2h}$ geometry.]({{< var fig_pentalene_orbitals >}}.png){#fig-pentalene-orbitals}
+![Orbital energies and frontier orbitals of pentalene at the optimized $C_\mathrm{2h}$ geometry (optimized with B97-3c).]({{< var fig_pentalene_orbitals >}}.png){#fig-pentalene-orbitals}
 
@@
 
+![Singlet-triplet energy gaps with CIS + DSP for azaphenalenes at different geometries compared to DFT.]({{< var fig_opt_corr >}}.pdf){#fig-opt-corr}
+
+![Singlet-triplet energy gaps with CIS + DSP against the reference level for azaphenalenes at different geometries.]({{< var fig_opt_gap >}}.pdf){#fig-opt-gap}
+
+\small
+```{python}
+# | label: tbl-opt
+# | tbl-cap: Metrics for azaphenalenes at different geometries.
+# | echo: false
+import pandas as pd
+from IPython.display import display, Markdown
+
+with open(tab_opt) as f:
+    md = f.read()
+display(Markdown(md))
+```
+\normalsize
+
+\small
 ```{python}
@@
 ```
+\normalsize
+
+{{< pagebreak >}}
+
+# References
+
+(S1) Tharwat, A. Classification Assessment Methods. *Appl. Comput. Inform.* **2021**, *17* (1), 168–192. https://doi.org/10.1016/j.aci.2018.08.003.

### 2024-01-25 workflow/notebooks/calculate.py

@@ -12,17 +12,19 @@ from coulson.ppp import (
 )
 
 
-def calculate_compound(filename: str | PathLike) -> dict[str, list[int]]:
+def calculate_compound(
+    filename: str | PathLike, overlap_method: str = "interpolation"
+) -> dict[str, list[int]]:
     """Calculate excitation roperties of molecule."""
     # Process filename with RDKit mol detection from XYZ
-    start_time = time.time()
+    start_time = time.perf_counter()
     mol = mol_from_xyz(filename)
     input_data, mask = process_rdkit_mol(mol)
     if not any(mask) is True:
         raise ValueError("Molecule does not have pi system.")
 
     # Perform SCF calculation
-    ppp = PPPCalculator(input_data)
+    ppp = PPPCalculator(input_data, overlap_method=overlap_method)
     ppp.scf(max_iter=500)
 
     # Calcualte exchange integral
@@ -45,7 +47,7 @@ def calculate_compound(filename: str | PathLike) -> dict[str, list[int]]:
     dsp_cis, _ = calculate_dsp(
         ppp, ci=True, energy_s_1=energy_s1_cis, energy_t_1=energy_t1_cis
     )
-    end_time = time.time()
+    end_time = time.perf_counter()
     run_time = end_time - start_time
 
     results = {

### 2024-01-25 workflow/notebooks/utils.py

@@ -14,7 +14,7 @@ import PIL
 import scipy.stats
 import sklearn.metrics
 import skunk
-from coulson.data import HARTREE_TO_EV, HARTREE_TO_KCAL
+from coulson.data import HARTREE_TO_EV
 from coulson.draw import draw_mol, draw_orbital_energies
 from coulson.interface import process_rdkit_mol
 from coulson.ppp import (
@@ -32,7 +32,16 @@ AXIS_LABELS: dict[str, str] = {
     "t1_s1_ref": r"$\Delta E_{\mathrm{ST,REF}}$ (eV)",
     "t1_s1_ppp": r"$\Delta E_{\mathrm{ST,CIS}}$ (eV)",
     "t1_s1_dsp_scf": r"$\Delta E_{\mathrm{ST,SCF}}^{\mathrm{DSP}}$ (eV)",
+    "t1_s1_dsp_scf_interpolation": r"$\Delta E_{\mathrm{ST,SCF}}^{\mathrm{DSP}}$ (eV) interp.",
+    "t1_s1_dsp_scf_no_interpolation": r"$\Delta E_{\mathrm{ST,SCF}}^{\mathrm{DSP}}$ (eV) no interp.",
     "t1_s1_dsp_cis": r"$\Delta E_{\mathrm{ST,CIS}}^{\mathrm{DSP}}$ (eV)",
+    "t1_s1_dsp_cis_dft": r"$\Delta E_{\mathrm{ST,CIS,DFT}}^{\mathrm{DSP}}$ (eV)",
+    "t1_s1_dsp_cis_mmff": r"$\Delta E_{\mathrm{ST,CIS,MMFF}}^{\mathrm{DSP}}$ (eV)",
+    "t1_s1_dsp_cis_gfnff": r"$\Delta E_{\mathrm{ST,CIS,GFNFF}}^{\mathrm{DSP}}$ (eV)",
+    "t1_s1_dsp_cis_gfn2": r"$\Delta E_{\mathrm{ST,CIS,GFN2xTB}}^{\mathrm{DSP}}$ (eV)",
+    "t1_s1_dsp_cis_ani1ccx": r"$\Delta E_{\mathrm{ST,CIS,ANI1ccx}}^{\mathrm{DSP}}$ (eV)",
+    "t1_s1_dsp_cis_interpolation": r"$\Delta E_{\mathrm{ST,CIS}}^{\mathrm{DSP}}$ (eV) interp.",
[…]
